### 2020-12-10 README.Rmd
added README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fasano.franceschini.test

<!-- badges: start -->
<!-- badges: end -->

The fasano.franceschini.test package is an R implementation of the 2-D Kolmogorov-Smirnov (KS) two-sample test as defined by Fasano and Franceschini (1987). This is a variant of the 2-D two-sample KS test as originally defined by Peacock (1983).

## Installation

You can install the released version of fasano.franceschini.test from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("fasano.franceschini.test")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nesscoder/fasano.franceschini.test")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(fasano.franceschini.test)
## basic example code

sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 3),
                          y = rnorm(n = 100,mean = 0, sd = 1))
sample2Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1),
                          y = rnorm(n = 100,mean = 0, sd = 3))

fasano.franceschini.test(sample1Data,sample2Data)
```

### 2020-12-10 R/fasano_franceschini_test.R

@@ -40,7 +40,7 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
     x_marg <- c(S1[, 1], S2[, 1])
     y_marg <- c(S1[, 2], S2[, 2])
 
-    d <- mclapply(X = 1:nBootstrap, mc.cores = cores, FUN = function(i) {
+    d <- parallel::mclapply(X = 1:nBootstrap, mc.cores = cores, FUN = function(i) {
       S1_resample <- data.frame(
         x = sample(x = x_marg, size = n1, replace = T),
         y = sample(x = y_marg, size = n1, replace = T)

### 2020-12-10 R/interal_functions_FasFran.R

@@ -74,7 +74,7 @@ get_d_stat <- function(originSamples, S1, S2, cores = 1) {
 
   d <- 0
   n <- dim(originSamples)[1]
-  dList <- mclapply(X = 1:n, mc.cores = cores, FUN = function(j){
+  dList <- parallel::mclapply(X = 1:n, mc.cores = cores, FUN = function(j){
     quadct_S1 <- quadCount(originSamples[j, 1], originSamples[j, 2], S1[, 1], S1[, 2])
     quadct_S2 <- quadCount(originSamples[j, 1], originSamples[j, 2], S2[, 1], S2[, 2])
     d <- max(

### 2020-12-10 R/fasano_franceschini_test.R
updated documentation and readme file

@@ -1,11 +1,12 @@
 #' Fasano Franceschini Test
 #'
 #' Computes the 2-D Kolmogorov-Smirnov two-sample test as described by Fasano and Franceschini (1987).
+#'
 #' Code adapted from Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688
 #'
 #' @param S1 a [n by 2] `data.frame` of x and y coordinates of sample 1
 #' @param S2 a [n by 2] `data.frame` of x and y coordinates of sample 2
-#' @param nBootstrap `numeric` number of bootstrapped samples to be generated for computing the empirical p-value (note this procedure is slow and computationally expensive on the order of nBootStrap*O(n^2). Default is set to 0.
+#' @param nBootstrap a `numeric` defining the number of bootstrapped samples to be generated for computing the empirical p-value (note this procedure is slow and computationally expensive on the order of nBootStrap*O(n^2). Default is set to 0.
 #' If nBootstrap is 0, the Fasano Franceschini distributional approximation is used for defining the p-value. See Fasano and Franceschini test (1987) for details.
 #' @param cores a `numeric` defining the number of cores to use of processing
 #'
@@ -14,9 +15,10 @@
 #'
 #' @references{
 #' \itemize{
-#' \item{Fasano, G., Franceschini, A. (1987)<DOI:10.1093/mnras/225.1.155>. A multidimensional version of the Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society 225:155-170.}}
-#' \item{Peacock J.A. (1983) <DOI:10.1093/mnras/202.3.615>. Two-dimensional goodness-of-fit testing in astronomy. Monthly Notices of the Royal Astronomical Society 202:615-627.}}
-#' \item{Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688}}
+#' \item{Fasano, G., Franceschini, A. (1987)<DOI:10.1093/mnras/225.1.155>. A multidimensional version of the Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society 225:155-170.}
[…]

### 2020-12-10 R/interal_functions_FasFran.R

@@ -90,7 +90,8 @@ get_d_stat <- function(originSamples, S1, S2, cores = 1) {
 
 #' KS probability
 #'
-#' p-value of getting the specified 2-D KS stat
+#' p-value of getting the specified 2-D KS stat.
+#'
 #' Code adapted from Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688
 #'
 #'

### 2020-12-10 README.Rmd
updated documentation and readme file
updated Fasano Franceschini R Console Output

@@
 
-# fasano.franceschini.test
+# Fasano Franceschini Test
 
@@
 
-This is a basic example which shows you how to solve a common problem:
+#### Underlying Distributions Are Different
 
-```{r example}
+```{r}
 library(fasano.franceschini.test)
-## basic example code
 
+#set seed for reproducible example
+set.seed(123)
+
+#create 2-D samples with the different underlying distributions
 sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 3),
@@
 ```
+
+#### Underlying Distributions Are The Same
+
+```{r}
+#set seed for reproducible example
+set.seed(123)
+
+#create 2-D samples with the same underlying distributions
+sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1),
+                          y = rnorm(n = 100,mean = 0, sd = 1))
+sample2Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1),
+                          y = rnorm(n = 100,mean = 0, sd = 1))
+
+fasano.franceschini.test(sample1Data,sample2Data)
+```

### 2020-12-10 R/fasano_franceschini_test.R

@@ -8,10 +8,9 @@
 #' @param S2 a [n by 2] `data.frame` of x and y coordinates of sample 2
 #' @param nBootstrap a `numeric` defining the number of bootstrapped samples to be generated for computing the empirical p-value (note this procedure is slow and computationally expensive on the order of nBootStrap*O(n^2). Default is set to 0.
 #' If nBootstrap is 0, the Fasano Franceschini distributional approximation is used for defining the p-value. See Fasano and Franceschini test (1987) for details.
-#' @param cores a `numeric` defining the number of cores to use of processing
+#' @param cores a `numeric` defining the number of cores to use for processing
 #'
 #' @return the 2-D ks statistic and p-value
-#' @export
 #'
 #' @references{
 #' \itemize{
@@ -22,6 +21,29 @@
 #' }
 #'
 #' @examples
+#' #Underlying distributions are different
+#' #set seed for reproducible example
+#' set.seed(123)
+#'
+#' #create 2-D samples with different underlying distributions
+#' sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 3), y = rnorm(n = 100,mean = 0, sd = 1))
+#' sample2Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1), y = rnorm(n = 100,mean = 0, sd = 3))
+#'
+#' fasano.franceschini.test(sample1Data,sample2Data)
+#'
+#'
+#' #Underlying distributions are the same
+#' #set seed for reproducible example
+#' set.seed(123)
+#'
+#' #create 2-D samples with the same underlying distributions
+#' sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1), y = rnorm(n = 100,mean = 0, sd = 1))
+#' sample2Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1), y = rnorm(n = 100,mean = 0, sd = 1))
+#'
+#' fasano.franceschini.test(sample1Data,sample2Data)
+#'
+#' @export
+
 fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
   start <- Sys.time()
   # determine number of samples in each data set
@@ -57,8 +79,13 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
     # count the number of bootstrapped d stats that are larger than the observed
[…]

### 2020-12-10 R/interal_functions_FasFran.R

@@ -13,14 +13,8 @@
 #' @return a `vector` of frequencies of the number of points in each of the four quadrants defined by the origin point
 #' @export
 #'
-#' @references{
-#' \itemize{
-#'    \item{Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688}}
-#' }
 #'
 #'
-#'
-#' @examples
 quadCount <- function(x_origin, y_origin, x, y) {
   n <- length(x)
   nq1 <- 0
@@ -69,7 +63,6 @@ quadCount <- function(x_origin, y_origin, x, y) {
 #' @return a `numeric` defining the D stat with the largest difference between the quad frequencies, after checking each point as the origin
 #' @export
 #'
-#' @examples
 get_d_stat <- function(originSamples, S1, S2, cores = 1) {
 
   d <- 0
@@ -100,7 +93,6 @@ get_d_stat <- function(originSamples, S1, S2, cores = 1) {
 #' @return a `numeric` defining the p-value of observing the given 2-D KS stat
 #' @export
 #'
-#' @examples
 ksCDF <- function(lambda) {
   eps1 <- 0.001
   eps2 <- 1E-8

### 2020-12-10 R/fasano_franceschini_test.R
updated documentation

@@ -52,9 +52,9 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
 
   # is there a better way to do the S1 from outer to inner function?
   # determine max difference assuming first sample as origins
-  d1 <- get_d_stat(originSamples = S1, S1 = S1, S2 = S2, cores = cores)
+  d1 <- getDstat(originSamples = S1, S1 = S1, S2 = S2, cores = cores)
   # determine max difference assuming second sample as origins
-  d2 <- get_d_stat(originSamples = S2, S1 = S1, S2 = S2, cores = cores)
+  d2 <- getDstat(originSamples = S2, S1 = S1, S2 = S2, cores = cores)
 
   # average KS stat
   D <- (d1 + d2) / 2
@@ -73,7 +73,7 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
         x = sample(x = x_marg, size = n2, replace = T),
         y = sample(x = y_marg, size = n2, replace = T)
       )
-      get_d_stat(originSamples = S1_resample, S1 = S1_resample, S2 = S2_resample, cores = 1)
+      getDstat(originSamples = S1_resample, S1 = S1_resample, S2 = S2_resample, cores = 1)
     })
 
     # count the number of bootstrapped d stats that are larger than the observed

### 2020-12-10 R/interal_functions_FasFran.R

@@ -63,7 +63,7 @@ quadCount <- function(x_origin, y_origin, x, y) {
 #' @return a `numeric` defining the D stat with the largest difference between the quad frequencies, after checking each point as the origin
 #' @export
 #'
-get_d_stat <- function(originSamples, S1, S2, cores = 1) {
+getDstat <- function(originSamples, S1, S2, cores = 1) {
 
   d <- 0
   n <- dim(originSamples)[1]

### 2020-12-15 README.Rmd
updated readme and package vignette

@@
 
-The fasano.franceschini.test package is an R implementation of the 2-D Kolmogorov-Smirnov (KS) two-sample test as defined by Fasano and Franceschini (1987).
+The fasano.franceschini.test package is an R implementation of the 2-D Kolmogorov-Smirnov (**KS**) two-sample test as defined by Fasano and Franceschini (1987).
 This is a variant of the 2-D two-sample KS test as originally defined by Peacock (1983).
@@
 #create 2-D samples with the different underlying distributions
-sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 3),
-                          y = rnorm(n = 100,mean = 0, sd = 1))
-sample2Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1),
-                          y = rnorm(n = 100,mean = 0, sd = 3))
+sample1Data <- data.frame(
+  x = rnorm(n = 100, mean = 0, sd = 3),
+  y = rnorm(n = 100, mean = 0, sd = 1)
+)
+sample2Data <- data.frame(
+  x = rnorm(n = 100, mean = 0, sd = 1),
+  y = rnorm(n = 100, mean = 0, sd = 3)
+)
 
@@
 #create 2-D samples with the same underlying distributions
-sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1),
-                          y = rnorm(n = 100,mean = 0, sd = 1))
-sample2Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1),
-                          y = rnorm(n = 100,mean = 0, sd = 1))
+sample1Data <- data.frame(
+  x = rnorm(n = 100, mean = 0, sd = 1),
+  y = rnorm(n = 100, mean = 0, sd = 1)
+)
+sample2Data <- data.frame(
+  x = rnorm(n = 100, mean = 0, sd = 1),
+  y = rnorm(n = 100, mean = 0, sd = 1)
+)
 

### 2020-12-15 R/fasano_franceschini_test.R

@@ -78,10 +78,9 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
 
     # count the number of bootstrapped d stats that are larger than the observed
     pval <- sum(unlist(d) > D) / nBootstrap
-    ksStat <- D
 
     new("Fasano Franceschini Test",
-        list(ksStat = ksStat,
+        list(D = D,
              pval = pval,
              time = Sys.time() - start,
              S1 = deparse(substitute(S1)),
@@ -95,10 +94,10 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
 
     # compute KS stat
     n <- n1 * n2 / (n1 + n2)
-    ksStat <- sqrt(n) * D / (1 + sqrt(1 - rr) * (0.25 - 0.75 / sqrt(n)))
-    pval <- ksCDF(ksStat)
+    lambda <- sqrt(n) * D / (1 + sqrt(1 - rr) * (0.25 - 0.75 / sqrt(n)))
+    pval <- ksCDF(lambda)
     new("Fasano Franceschini Test",
-        list(ksStat = ksStat,
+        list(D = D,
              pval = pval,
              time = Sys.time() - start,
              S1 = deparse(substitute(S1)),
@@ -111,10 +110,10 @@ setClass( "Fasano Franceschini Test", representation("list"))
 
 setMethod("show", "Fasano Franceschini Test", function(object) {
   cat("\n")
-  cat("\t2-D Two-sample Kolmogorov-Smirnov Test\n")
+  cat("\t\t2-D Two-sample Kolmogorov-Smirnov Test\n")
   cat("\n")
   cat("Fasano Franceschini Test (1987)\n")
   cat("Data: ", object$S1, "and", object$S2,"\n")
-  cat("D-stat = ", object$ksStat,", p-value = ", object$pval,"\n")
+  cat("D-stat = ", object$D,", p-value = ", object$pval,"\n")
   cat("Run Time (s) = ", object$time,"\n")
 })

### 2020-12-16 journalMS/article.tex
initial commit of package manuscript

\documentclass[codesnippet]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Elan Ness-Cohn\\Northwestern University
   \And Rosemary Braun\\Northwestern University}
\Plainauthor{Elan Ness-Cohn, Rosemary Braun}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{Fasano Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in \proglang{R}}
\Plaintitle{Fasano Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in R}
\Shorttitle{2-Dimensional Kolmogorov--Smirnov in \proglang{R}}

%% - \Abstract{} almost as usual
\Abstract{
  The \pkg{fasano.franceschini.test} package is an  \proglang{R} implementation of the 2-D Kolmogorov-Smirnov (KS) two--sample test as defined by Fasano and Franceschini \citep{Fasano1987}. This is a variant of the 2-D two-sample KS test as originally defined by Peacock \citep{Peacock1983}, previously implemented in \proglang{R} in the \pkg(Peacock.test) package \citep{Xiao2016}. The \pkg{fasano.franceschini.test} package provides a two fold improvement over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): \textbf{(i)} The Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$. \textbf{(ii)} The package implements a parrellized bootstrapping procedure for improved significance testing.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{Fasano--Franceschini test, 2-D Kolmogorov--Smirnov test, Multivariate, KS test, \proglang{R}}
\Plainkeywords{Fasan--Franceschini test, 2-D Kolmogorov--Smirnov test, Multivariate, KS test,  R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Elan Ness-Cohn\\
  Department of Molecular Biosciences,\\
  Biostatistics Division of The Department of Preventive Medicine,\\
  \emph{and}\\
  NSF-Simons Center for Quantitative Biology\\
  Northwestern University\\
  Evanston, IL 60208\\
  E-mail: \email{elan.ness-cohn@northwestern.edu}\\
  URL: \url{https://sites.northwestern.edu/elannesscohn/}\\

Rosemary Braun\\
  Department of Molecular Biosciences,\\
  Biostatistics Division of The Department of Preventive Medicine,\\
  NSF-Simons Center for Quantitative Biology,\\
  Department of Engineering Sciences and Applied Mathematic,\\
  Department of Physics and Astronomy,\\
  \emph{and}\\
  Northwestern Instutute on Complex Systems\\
  Northwestern University\\
  Evanston, IL 60208\\
  E-mail: \email{rbraun@northwestern.edu}\\
  URL: \url{https://sites.northwestern.edu/braunlab/}\\

}
\begin{document}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section[Introduction]{Introduction} \label{sec:intro}

% R -> Peacock test (computationally inificent)
% Python <- Implementation
% C <- Implementation

The Kolmogorov--Smirnov (KS) is a non--parametric method for determining whether a sample is consistent with a given probability distribution. First derived by Kolmogorov and Smirnov \citep{KStest}, the was developed in 1-D. There is still a desire to be able to test 2-D and higher distributions against eachother. As a result, the first 2-D KS test was articulated by Peacock \citep{Peacock1983}.

The \pkg{fasano.franceschini.test} package is an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987}. This is a variant of the 2-D two-sample KS test as originally defined by Peacock \citep{Peacock1983}. In \proglang{R} \citep{R}, the only current implementation of the 2-D two-sample KS test is available in the \pkg{Peacock.test} package via the \fct{peacock2} function. The manuscript that this document is based on \citep{Fasano1987} then introduced hurdle. A \proglang{C} implementation of the Fasano Franceschini test is available in \citep{numericalRecipes}; however does not offer a bootstrapping procedure nor has been documentation for easy of use in \proglang{R} community.

% \begin{leftbar}
% The introduction is in principle ``as usual''. However, it should usually embed
% both the implemented \emph{methods} and the \emph{software} into the respective
% relevant literature. For the latter both competing and complementary software
% should be discussed (within the same software environment and beyond), bringing
% out relative (dis)advantages. All software mentioned should be properly
% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
% \textsc{Bib}{\TeX}.)
%
% For writing about software JSS requires authors to use the markup
% \verb|\proglang{}| (programming languages and large programmable systems),
% \verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
% arguments, etc.). If there is such markup in (sub)section titles (as above), a
% plain text version has to be provided in the {\LaTeX} command as well. Below we
% also illustrate how abbrevations should be introduced and citation commands can
% be employed. See the {\LaTeX} code for more details.
% \end{leftbar}
%
% Modeling count variables is a common task in economics and the social sciences.
% The classical Poisson regression model for count data is often of limited use in
% these disciplines because empirical count data sets typically exhibit
% overdispersion and/or an excess number of zeros. The former issue can be
% addressed by extending  the plain Poisson regression model in various
% directions: e.g., using sandwich covariances or estimating an additional
% dispersion parameter (in a so-called quasi-Poisson model). Another more formal
% way is to use a negative binomial (NB) regression. All of these models belong to
% the family of generalized linear models (GLMs). However, although these models
% typically can capture overdispersion rather well, they are in many applications
% not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
% increased interest in zero-augmented models that address this issue by a second
% model component capturing zero counts. An overview of count data models in
% econometrics, including  hurdle and zero-inflated models, is provided in
% \cite{Cameron+Trivedi:2013}.
%
% In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
% \fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
% \citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
% diagnostics and inference. The manuscript that this document is based on
% \citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
% count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
% package \citep{Jackman:2015}. Of course, much more software could be discussed
% here, including (but not limited to) generalized additive models for count data
% as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006},
% \pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

\section{Models and software} \label{sec:models}

\subsection{1-D Kolmogorov--Smirnov Test}

The Kolmogorov--Smirnov (KS) test is a non--parametric method for determining where a sample is consistent with a given probability distribution. While a complete discussion of the 1-D KS test can be found elsewhere, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.

The 1-D KS test can be divided into to variants of the same statistical procedure:

\begin{leftbar}
\textbf{The One-Sample Test}
\begin{itemize}
\item Compares the data to an analytical distribution.
\end{itemize}
\textbf{The Two-Sample Test}
\begin{itemize}
\item Compares two data samples to determine if they are drawn from the same underlying probability distribution.
\end{itemize}
\end{leftbar}

In both cases, the KS test works as follows:

The Kolmogorov-Smimov statistic (D) is the defined by the maximum difference between the cumulative density functions of the data and model (one--sample), or between the two data sets (two--sample). In the following example, the orange and blue curve in the left panel represents the underlying distribution (pdf) for sample1 and sample2, respectively (\textbf{Figure \ref{fig:kstest1D}}). The eCDF is then defined by data drawn from these underlying distribution, and the D--stat is the computed as the largest between difference between the eCDF (black dotted line) in the right panel.

\begin{figure}[t!]
\centering
\includegraphics{pdfvsCDF}
\caption{\label{fig:kstest1D} Frequency distribution for number of days absent
from school.}
\end{figure}

Kolmogorov and Smimov work defined the cumulative distribution for a random D-stat as:

\begin{equation} \label{eq:1}
Q_{KS}(\lambda) = 2 \sum_{j=1}^{\infty} -1^{j-1}e^{-2j^2\lambda^2}
\end{equation}

and thus allows for $p$-value hypothesis testing to determine if there is a significance difference between the cumulative density functions of the data and model (one-sample), or between the two data sets (two-sample).

\textbf{For the 1-Sample case}, the $p$-value probability of a given D-stat is defined by

\begin{equation} \label{eq:2}
Prob(D > observed) = Q_{KS} ( D\sqrt{N})
\end{equation}

where $N$ is the number of data points sampled. A $p$-value $< 0.05$, indicates a significant difference between the samples.

\textbf{In the 2-Sample case}, the $p$-value probability of a given D-stat is defined by
\begin{equation} \label{eq:3}
Prob(D > observed) = Q_{KS} ( D\sqrt{\frac{n_1n_2}{n_1+n_2}})
\end{equation}

where $n1$ and $n2$ is the number of data points in the first and second samples respectively. A $p$-value $< 0.05$, indicates a significant difference between the samples.

\subsection{Higher Dimensional KS Test}

Similar to the 1-D KS test, the obvious solution to extend the theory into higher dimensional would be to define the maximal cumulative difference between the two 2-D distributions. This however, presents an issue as the commutative distribution function is not well defined in more than one dimension\citep{Peacock1983}.

\subsection{Peacock Test (1983) and Fasano Franceschini Test (1987)}
The Peacock test was the original generalization of the 1-D KS test to higher dimensions. Peacock solved the higher dimensionality issue by using the total intergrated probability -- i.e. fraction of data -- in each of the four quadrants around all possible points in the data. For example, for an given $n$ points in a two-dimensional space, the cumulative distribution functions is calculated in the $4n^2$ quadrants of the plane defined by all pairs ($X_i$ ,$Y_j$), $X_i$ and $Y_j$ being coordinates of any pairs of points in the given sample. By ranging over all possible pairs of data points and quadrants, the 2-D D-stat is defined by the maximal difference of the integrated probabilities between samples.

The slight variation defined by Fasano and Franceschini was to only consider quadrants centered in each point of the given samples to compute the cumulative distribution functions. e.g, rather than looking over all ($X_i$ ,$Y_j$; where $i,j={1,n}$), only use the detected points ($X_i$ ,$Y_i$; where $i={1,n}$). Thus for an given $n$ points in a two-dimensional space, those $n$ points define $4n$ quadrants.

The 2-D two sample Fasano and Franceschini test is illustrated in the following example to compare the 2-D distributions of orange vs blue points (\textbf{Figure \ref{fig:kstest2D}}).

\begin{figure}[t!]
\centering
\includegraphics{fftestOutput}
\caption{\label{fig:kstest2D} Frequency distribution for number of days absent
from school.}
\end{figure}

The algorithm loops through each point in the orange sample in turn to define the origin of 4 quadrants (grey dotted line). The fraction of orange and blue points in each quadrant is computed (corners of plot) and the quadrant with the maximal difference is designated the current maximum for the specified origin. By ranging over all pairs of data points and quadrants, the 2-D D-stat is defined by the maximal difference of the integrated probabilities between samples. In this case using the orange point as the origin, the maximal difference is $D = 0.35$.

This process is repeated using the blue points as the origins to compute the maximal D-stat. Finally, both maximal D-stats using the orange and blue points as origins are then averaged to compute the overall D-stat for hypothesis testing.

\subsection{Defining the Fasano and Franceschini Null Distribution}

Using Monte Carlo simulation, Fasano and Franceschini approximated the distribution of D for significance testing -- as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.

\textbf{For the 1-Sample case}

\begin{equation} \label{eq:4}
Prob(D > observed) = Q_{KS} ( \frac{D\sqrt{N}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{N})})
\end{equation}

\textbf{The 2-Sample cas}

Uses the same formula as above, but with the slight variation where

\begin{equation} \label{eq:5}
N = \frac{n_1n_2}{n_1+n_2}
\end{equation}

In both cases, $r$ is defined by:

\begin{equation} \label{eq:6}
r = \frac{\sum_{i}^{}(x_i-\bar{x})(y_i-\bar{y})}{\sqrt{\sum_{i}^{}(x_i-\bar{x})^2}\sqrt{\sum_{i}^{}(y_i-\bar{y})^2}}
\end{equation}

% The basic Poisson regression model for count data is a special case of the GLM
% framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
% response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
% $y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
% $\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
% log link and a linear predictor
% %
% \begin{equation} \label{eq:mean}
% \log(\mu_i) \quad = \quad x_i^\top \beta,
% \end{equation}
% %
% where the regression coefficients $\beta$ are estimated by maximum likelihood
% (ML) using the iterative weighted least squares (IWLS) algorithm.
%
% \begin{leftbar}
% Note that around the \verb|{equation}| above there should be no spaces (avoided
% in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
% not a new paragraph started.
% \end{leftbar}
%
% \proglang{R} provides a very flexible implementation of the general GLM
% framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
% \pkg{stats} package. Its most important arguments are
% \begin{Code}
% glm(formula, data, subset, na.action, weights, offset,
%   family = gaussian, start = NULL, control = glm.control(...),
%   model = TRUE, y = TRUE, x = FALSE, ...)
% \end{Code}
% where \code{formula} plus \code{data} is the now standard way of specifying
% regression relationships in \proglang{R}/\proglang{S} introduced in
% \cite{Chambers+Hastie:1992}. The remaining arguments in the first line
% (\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
% standard  for setting up formula-based regression models in
% \proglang{R}/\proglang{S}. The arguments in the second line control aspects
% specific to GLMs while the arguments in the last line specify which components
% are returned in the fitted model object (of class \class{glm} which inherits
% from \class{lm}). For further arguments to \fct{glm} (including alternative
% specifications of starting values) see \code{?glm}. For estimating a Poisson
% model \code{family = poisson} has to be specified.
%
% \begin{leftbar}
% As the synopsis above is a code listing that is not meant to be executed,
% one can use either the dedicated \verb|{Code}| environment or a simple
% \verb|{verbatim}| environment for this. Again, spaces before and after should be
% avoided.
%
% Finally, there might be a reference to a \verb|{table}| such as
% Table~\ref{tab:overview}. Usually, these are placed at the top of the page
% (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
% column headers and captions in sentence style, and if possible avoiding vertical
% lines.
% \end{leftbar}
%
% \begin{table}[t!]
% \centering
% \begin{tabular}{lllp{7.4cm}}
% \hline
% Type           & Distribution & Method   & Description \\ \hline
% GLM            & Poisson      & ML       & Poisson regression: classical GLM,
%                                            estimated by maximum likelihood (ML) \\
%                &              & Quasi    & ``Quasi-Poisson regression'':
%                                            same mean function, estimated by
%                                            quasi-ML (QML) or equivalently
%                                            generalized estimating equations (GEE),
%                                            inference adjustment via estimated
%                                            dispersion parameter \\
%                &              & Adjusted & ``Adjusted Poisson regression'':
%                                            same mean function, estimated by
%                                            QML/GEE, inference adjustment via
%                                            sandwich covariances\\
%                & NB           & ML       & NB regression: extended GLM,
%                                            estimated by ML including additional
%                                            shape parameter \\ \hline
% Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
%                                            hurdle Poisson \\
%                & NB           & ML       & Zero-inflated NB (ZINB),
%                                            hurdle NB \\ \hline
% \end{tabular}
% \caption{\label{tab:overview} Overview of various count regression models. The
% table is usually placed at the top of the page (\texttt{[t!]}), centered
% (\texttt{centering}), has a caption below the table, column headers and captions
% are in sentence style, and if possible vertical lines should be avoided.}
% \end{table}

%% -- Illustrations ------------------------------------------------------------

%% - Virtually all JSS manuscripts list source code along with the generated
%%   output. The style files provide dedicated environments for this.
%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%%   or knitr using the render_sweave() hook - are used (without the need to
%%   load Sweave.sty).
%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
%% - The code input should use "the usual" command prompt in the respective
%%   software system.
%% - For R code, the prompt "R> " should be used with "+  " as the
%%   continuation prompt.
%% - Comments within the code chunks should be avoided - these should be made
%%   within the regular LaTeX text.

\section{Illustrations} \label{sec:illustrations}

\subsection{Fasano Franceschini Test Usage}

In their paper, Fasano and Franceschini use Monte Carlo simulation to approximate the distribution of D as a function of $D$, the sample size $N$, and the coefficient of correlation $r$. The resultant distributional approximation -- as described above in equation \ref{eq:4} -- can be used to test the significance levels for the two-dimensional K-S test.

Note that the above formulas are accurate enough when $N \geq 20$, and when the indicated probability (significance level) is less than (more significant than) 0.20 or so. When the indicated probability is $> 0.20$, its value may not be accurate, but the implication that the data and model (or two data sets) are not significantly different is certainly correct.

\begin{CodeChunk}
\begin{CodeInput}
R> #set seed for reproducible example
R> set.seed(123)
R>
R> #create 2-D samples with the same underlying distributions
R> sample1Data <- data.frame(
R>  x = rnorm(n = 100, mean = 0, sd = 1),
R>  y = rnorm(n = 100, mean = 0, sd = 1)
R> )
R> sample2Data <- data.frame(
R>  x = rnorm(n = 100, mean = 0, sd = 1),
R>  y = rnorm(n = 100, mean = 0, sd = 1)
R> )
R>
R> fasano.franceschini.test(sample1Data,sample2Data)
\end{CodeInput}
\begin{CodeOutput}
      2-D Two-sample Kolmogorov-Smirnov Test

 Fasano Franceschini Test (1987)
 Data:  sample1Data and sample2Data
 D-stat =  0.14 , p-value =  0.4420642
 Run Time (s) =  0.04288507
\end{CodeOutput}
\end{CodeChunk}

\subsection{Bootstrap Version of the Fasano Franceschini Test}

Some arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.

To combat these accretions, one can bootstrap the significance levels for the particular multidimensional statistic directly from the particular data set under study. As Fasano and Franceschini's paper was originally released in 1987, this approach was unfeasible at scale. However, as we are currently in the age of high-performance computers -- a null distribution of D-stats can be bootstrapped from the data to test significance.

The \pkg{fasano.franceschini.test} R package, implements a parallelized bootstrapping procedure. The marginal distribution from 2-dimensional data set is resampled with replacements to generate randomized 2-dimensional data set \code{nBootStrap} times. The frequency count by quadrant is performed for each bootstrapped resampling as described above to compute the D-stats. The observed D-stat is then compared to the bootstrapped D-stats to compute a $p$-value. See \fct{fasano.franceschini.test} for source code details and implementation.

Note, bootstrapping may significantly increase run time performance. To improve run time, adjust the \code{cores} parameter to specify the number of parallel operations to run. See the R \pkg{parrellel} package and the \fct{mclapply} function for further details.

\begin{CodeChunk}
\begin{CodeInput}
R> #set seed for reproducible example
R> set.seed(123)
R>
R> #create 2-D samples with the same underlying distributions
R> sample1Data <- data.frame(
R>  x = rnorm(n = 100, mean = 0, sd = 1),
R>  y = rnorm(n = 100, mean = 0, sd = 1)
R> )
R> sample2Data <- data.frame(
R>  x = rnorm(n = 100, mean = 0, sd = 1),
R>  y = rnorm(n = 100, mean = 0, sd = 1)
R> )
R>
R> fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data,
R>                          nBootstrap = 1000, cores = 6)
\end{CodeInput}
\begin{CodeOutput}
      2-D Two-sample Kolmogorov-Smirnov Test

 Fasano Franceschini Test (1987)
 Data:  sample1Data and sample2Data
 D-stat =  0.14 , p-value =  0.672
 Run Time (s) =  2.694002
\end{CodeOutput}
\end{CodeChunk}

\subsection{Computational Efficency}
The main distinction between the Peacock and Fasano Franceschini test comes down to compute time. For a benchmarking of the algorithms see \citep{Lopes2007}

In Summary:
\begin{itemize}
\item The Fasano Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as dataset sizes scale.
\item The Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
\end{itemize}

See the \pkg{Peacock.test} package for \proglang{R} implementation of the Peacock test.

% For a simple illustration of basic Poisson and NB count regression the
% \code{quine} data from the \pkg{MASS} package is used. This provides the number
% of \code{Days} that children were absent from school in Australia in a
% particular year, along with several covariates that can be employed as regressors.
% The data can be loaded by
% %
% \begin{CodeChunk}
% \begin{CodeInput}
% R> data("quine", package = "MASS")
% \end{CodeInput}
% \end{CodeChunk}
% %
% and a basic frequency distribution of the response variable is displayed in
% Figure~\ref{fig:quine}.
%
% \begin{leftbar}
% For code input and output, the style files provide dedicated environments.
% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
% hook. Please make sure that all code is properly spaced, e.g., using
% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
% use ``the usual'' command prompt in the respective software system. For
% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
% the continuation prompt. Generally, comments within the code chunks should be
% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
% before and after code input/output should be avoided (see above).
% \end{leftbar}
%
% \begin{figure}[t!]
% \centering
% \includegraphics{article-visualization}
% \caption{\label{fig:quine} Frequency distribution for number of days absent
% from school.}
% \end{figure}
%
% As a first model for the \code{quine} data, we fit the basic Poisson regression
% model. (Note that JSS prefers when the second line of code is indented by two
% spaces.)
% %
% \begin{CodeChunk}
% \begin{CodeInput}
% R> m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
% +    family = poisson)
% \end{CodeInput}
% \end{CodeChunk}
% %
% To account for potential overdispersion we also consider a negative binomial
% GLM.
% %
% \begin{CodeChunk}
% \begin{CodeInput}
% R> library("MASS")
% R> m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
% \end{CodeInput}
% \end{CodeChunk}
% %
% In a comparison with the BIC the latter model is clearly preferred.
% %
% \begin{CodeChunk}
% \begin{CodeInput}
% R> BIC(m_pois, m_nbin)
% \end{CodeInput}
% \begin{CodeOutput}
%        df      BIC
% m_pois 18 2046.851
% m_nbin 19 1157.235
% \end{CodeOutput}
% \end{CodeChunk}
% %
% Hence, the full summary of that model is shown below.
% %
% \begin{CodeChunk}
% \begin{CodeInput}
% R> summary(m_nbin)
% \end{CodeInput}
% \begin{CodeOutput}
% Call:
% glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
%     init.theta = 1.60364105, link = log)
%
% Deviance Residuals:
%     Min       1Q   Median       3Q      Max
% -3.0857  -0.8306  -0.2620   0.4282   2.0898
%
% Coefficients: (1 not defined because of singularities)
%             Estimate Std. Error z value Pr(>|z|)
% (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
% EthN        -0.24591    0.39135  -0.628  0.52977
% SexM        -0.77181    0.38021  -2.030  0.04236 *
% AgeF1       -0.02546    0.41615  -0.061  0.95121
% AgeF2       -0.54884    0.54393  -1.009  0.31296
% AgeF3       -0.25735    0.40558  -0.635  0.52574
% LrnSL        0.38919    0.48421   0.804  0.42153
% EthN:SexM    0.36240    0.29430   1.231  0.21818
% EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876
% EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 **
% EthN:AgeF3   0.04721    0.44883   0.105  0.91622
% EthN:LrnSL   0.06847    0.34040   0.201  0.84059
% SexM:AgeF1   0.02257    0.47360   0.048  0.96198
% SexM:AgeF2   1.55330    0.51325   3.026  0.00247 **
% SexM:AgeF3   1.25227    0.45539   2.750  0.00596 **
% SexM:LrnSL   0.07187    0.40805   0.176  0.86019
% AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870
% AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363
% AgeF3:LrnSL       NA         NA      NA       NA
% ---
% Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
%
% (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)
%
%     Null deviance: 235.23  on 145  degrees of freedom
% Residual deviance: 167.53  on 128  degrees of freedom
% AIC: 1100.5
%
% Number of Fisher Scoring iterations: 1
%
%
%               Theta:  1.604
%           Std. Err.:  0.214
%
%  2 x log-likelihood:  -1062.546
% \end{CodeOutput}
% \end{CodeChunk}

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}

The \pkg{fasano.franceschini.test} package is an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987}.
Improvements:
  (i) Fast algorith
  (ii) Prallelized Bootstrapping Procedure for Improved statisitcal testing

  See \url{https://nesscoder.github.io/fasano.franceschini.test/} for the complete package documentation and source code.

%% -- Optional special unnumbered sections -------------------------------------

\section*{Computational details}

% \begin{leftbar}
% If necessary or useful, information about certain computational details
% such as version numbers, operating systems, or compilers could be included
% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
% maps, tables, \dots) that are not cited in the main text can be credited here.
% \end{leftbar}

The results in this paper were obtained using
\proglang{R}~3.6.3 with the
\pkg{fasano.franceschini.test}~1.0.0 package. \proglang{R} itself
and all packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

Research reported in this publication was supported by the NSF-Simons Center for Quantitative Biology at Northwestern University, an NSF-Simons MathBioSys Research Center. This work was supported by a grant from the Simons Foundation/SFARI (597491-RWC) and the National Science Foundation (1764421). The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Science Foundation and Simons Foundation.

E.N.C developed the \pkg{fasano.franceschini.test} package and produced the tutorials/documentation; E.N.C.\,and R.B.\,wrote the paper.

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}
\newpage

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

% \begin{appendix}
%
% \section{More technical details} \label{app:technical}
%
% \begin{leftbar}
% Appendices can be included after the bibliography (with a page break). Each
% section within the appendix should have a proper section title (rather than
% just \emph{Appendix}).
%
% For more technical style details, please check out JSS's style FAQ at
% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
% which includes the following topics:
% \begin{itemize}
%   \item Title vs.\ sentence case.
%   \item Graphics formatting.
%   \item Naming conventions.
%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
%   \item Trouble shooting.
%   \item Many other potentially helpful details\dots
% \end{itemize}
% \end{leftbar}
%
%
% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
%
% \begin{leftbar}
% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
% formats of author-year citations and allow to include additional details (e.g.,
% pages, chapters, \dots) in brackets. In case you are not familiar with these
% commands see the JSS style FAQ for details.
%
% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
% when acquiring the entries automatically from mixed online sources. However,
% it is important that informations are complete and presented in a consistent
% style to avoid confusions. JSS requires the following format.
% \begin{itemize}
%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
%     be used in the references.
%   \item Titles should be in title case.
%   \item Journal titles should not be abbreviated and in title case.
%   \item DOIs should be included where available.
%   \item Software should be properly cited as well. For \proglang{R} packages
%     \code{citation("pkgname")} typically provides a good starting point.
% \end{itemize}
% \end{leftbar}
%
% \end{appendix}

%% -----------------------------------------------------------------------------

\end{document}

### 2020-12-16 journalMS/article.R
added script for journal figures

@@ -1,33 +1,132 @@
 ## Package and options
-library("MASS")
-options(prompt = "R> ", continue = "+  ", width = 70,
-  useFancyQuotes = FALSE)
+library("fasano.franceschini.test")
+library("parallel")
 
+#-------------------------------------#
+#    fasano.franceschini.test Usage
+#-------------------------------------#
+# set seed for reproducible example
+set.seed(123)
 
-## Data
-data("quine", package = "MASS")
+# create 2-D samples with the same underlying distributions
+sample1Data <- data.frame(
+  x = rnorm(n = 100, mean = 0, sd = 1),
+  y = rnorm(n = 100, mean = 0, sd = 1)
+)
+sample2Data <- data.frame(
+  x = rnorm(n = 100, mean = 0, sd = 1),
+  y = rnorm(n = 100, mean = 0, sd = 1)
+)
 
+fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data)
 
-## Visualization
-par(mar = c(4, 4, 1, 1))
-plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
-axis(2)
-axis(1, at = 0:16 * 5, labels = FALSE)
-axis(1, at = 0:8 * 10)
-box()
+#-------------------------------------#
+# fasano.franceschini.test Bootstrap
+#-------------------------------------#
+# set seed for reproducible example
+set.seed(123)
 
+# create 2-D samples with the same underlying distributions
+sample1Data <- data.frame(
+  x = rnorm(n = 100, mean = 0, sd = 1),
+  y = rnorm(n = 100, mean = 0, sd = 1)
+)
+sample2Data <- data.frame(
+  x = rnorm(n = 100, mean = 0, sd = 1),
+  y = rnorm(n = 100, mean = 0, sd = 1)
+)
 
-## Poisson model
-m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-  family = poisson)
-summary(m_pois)
+fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data, nBootstrap = 1000, cores = 6)
 
+#-------------------------------------#
+#  Visualization - 2-D KS test
[…]

### 2020-12-17 journalMS/article.tex
journal article revisions and updated bib

@@
 The Kolmogorov--Smirnov (KS) is a non--parametric method for determining whether a sample is consistent with a given probability distribution.
-First derived by Kolmogorov and Smirnov \citep{KStest}, the was developed in 1-D.
+First derived by Kolmogorov and Smirnov \citep{Kolmogorov1933,Smirnov1948}, the was developed in 1-D.
 There is still a desire to be able to test 2-D and higher distributions against eachother.
@@
 The Kolmogorov--Smirnov (KS) test is a non--parametric method for determining where a sample is consistent with a given probability distribution.
-While a complete discussion of the 1-D KS test can be found elsewhere, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
+While a comprehensive discussion and intracies of the 1-D KS test can be found elsewhere\citep{Stephens1992}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
 

### 2020-12-17 journalMS/article.tex
updated references to KS test

--- 2020-12-16 journalMS/article.tex
@@
 The Kolmogorov--Smirnov (KS) is a non--parametric method for determining whether a sample is consistent with a given probability distribution.
-First derived by Kolmogorov and Smirnov \citep{KStest}, the was developed in 1-D.
+First derived by Kolmogorov and Smirnov in a series of papers~\citep{Kolmogorov1933,Kolmogorov1933a,Smirnov1936,Smirnov1937,Smirnov1939,Smirnov1944,Smirnov1948}, the was developed in 1-D.
 There is still a desire to be able to test 2-D and higher distributions against eachother.
@@
 The Kolmogorov--Smirnov (KS) test is a non--parametric method for determining where a sample is consistent with a given probability distribution.
-While a complete discussion of the 1-D KS test can be found elsewhere, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
+While a comprehensive discussion pertaining the intracies of the 1-D KS test can be found elsewhere~\citep{Stephens1992}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
 

### 2020-12-19 journalMS/article.tex
updated KS equations

@@
 The Kolmogorov--Smirnov (KS) test is a non--parametric method for determining where a sample is consistent with a given probability distribution.
-While a comprehensive discussion pertaining the intracies of the 1-D KS test can be found elsewhere~\citep{Stephens1992}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
+While a comprehensive discussion pertaining the intracies of the 1-D KS test can be found elsewhere~\citep{Stephens1992a}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
 
@@
 \begin{equation} \label{eq:1}
-Q_{KS}(\lambda) = 2 \sum_{j=1}^{\infty} -1^{j-1}e^{-2j^2\lambda^2}
+\Phi(\lambda) = 2 \sum_{k=1}^{\infty} -1^{k-1}e^{-2k^2\lambda^2}
 \end{equation}
@@
 \begin{equation} \label{eq:2}
-Prob(D > observed) = Q_{KS} ( D\sqrt{N})
+Prob(D > observed) = \Phi ( D\sqrt{N})
 \end{equation}
@@
 \begin{equation} \label{eq:3}
-Prob(D > observed) = Q_{KS} ( D\sqrt{\frac{n_1n_2}{n_1+n_2}})
+Prob(D > observed) = \Phi ( D\sqrt{\frac{n_1n_2}{n_1+n_2}})
 \end{equation}
@@
 \begin{equation} \label{eq:4}
-Prob(D > observed) = Q_{KS} ( \frac{D\sqrt{N}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{N})})
+Prob(D > observed) = \Phi ( \frac{D\sqrt{N}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{N})})
 \end{equation}

### 2020-12-19 journalMS/article.tex
updated references to 2-D KS equations

@@
 
-Kolmogorov and Smimov work defined the cumulative distribution for a random D-stat as:
+Kolmogorov and Smimov work defined the cumulative distribution for a random D-stat, which was later refined by Kendall and Stuart~\citep{Kendall1946}:
 
@@
 Similar to the 1-D KS test, the obvious solution to extend the theory into higher dimensional would be to define the maximal cumulative difference between the two 2-D distributions.
-This however, presents an issue as the commutative distribution function is not well defined in more than one dimension\citep{Peacock1983}.
+This however, presents an issue as the commutative distribution function is not well defined in more than one dimension.
+In N dimensional space there are $2^{N}-1$ independant ways of defining the cumlative distribution function\citep{Peacock1983}.
+
+ADD FIGURE for showing the 4 ways (3 independent ways) of defining the cummaltive distribution in 2-D.
+$P(x < X, y < Y) = 1 - P(x > X, y > Y); P(x < X, y > Y); P(x > X, y < Y)$
 

### 2020-12-23 journalMS/article.tex
updated ff algorithm figure
added figure for issue with 2d CDF

@@
 This however, presents an issue as the commutative distribution function is not well defined in more than one dimension.
+In 2-D, there are 4 ways (3 independent ways) of defining the cummaltive distribution (\textbf{Figure \ref{fig:kstest2Dissue}}).
 In N dimensional space there are $2^{N}-1$ independant ways of defining the cumlative distribution function\citep{Peacock1983}.
 
-ADD FIGURE for showing the 4 ways (3 independent ways) of defining the cummaltive distribution in 2-D.
-$P(x < X, y < Y) = 1 - P(x > X, y > Y); P(x < X, y > Y); P(x > X, y < Y)$
+\begin{figure}[t!]
+\centering
+\includegraphics{CDF2Dissue}
+\caption{\label{fig:kstest2Dissue} 4 ways (3 independent ways) of defining the cummaltive distribution for a given point in 2-D.
+$P(x < X, y < Y) = 1 - P(x > X, y > Y); P(x < X, y > Y); P(x > X, y < Y)$}
+\end{figure}
 
@@
 By ranging over all pairs of data points and quadrants, the 2-D D-stat is defined by the maximal difference of the integrated probabilities between samples.
-In this case using the orange point as the origin, the maximal difference is $D = 0.35$.
+In this case using the orange point as the origin, the maximal difference is $D = 0.60$.
 

### 2020-12-23 journalMS/article.R

@@ -41,24 +41,26 @@ fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data, nBootstrap = 1000, 
 #-------------------------------------#
 #  Visualization - 2-D KS test
 #-------------------------------------#
-png(file = paste0("~/Desktop/fftestOutput.png"), units = "in", res = 300, width = 14, height = 14)
-par(mfrow = c(4, 5))
-set.seed(10)
+png(file = paste0("~/Desktop/fftestOutput.png"),units = "in",res = 300,width = 14,height = 7)
+par(mfrow = c(2,5),oma = c(0, 0, 3, 0))
+set.seed(12)
+
 spread <- 2
-N <- 20
+N <- 10
+cexSize <- 1.5
 S1col <- "orange"
 S2col <- "blue"
 
 S1 <- data.frame(
-  x = rnorm(n = N, mean = 0, sd = 1),
+  x = rnorm(n = N, mean = 1, sd = 1),
   y = rnorm(n = N, mean = 0, sd = 1)
 )
 S2 <- data.frame(
-  x = rnorm(n = N, mean = 0, sd = 1),
+  x = rnorm(n = N, mean = -1, sd = 1),
   y = rnorm(n = N, mean = 0, sd = 1)
 )
 
-S1 <- S1[with(S1, order(S1[, 1], S1[, 2], decreasing = F)), ]
+S1 <- S1[with(S1, order(S1[, 1], S1[, 2],decreasing = F)), ]
 originSamples <- S1
 d <- rep(0, nrow(S1))
 for (j in 1:nrow(S1)) {
@@ -70,19 +72,23 @@ for (j in 1:nrow(S1)) {
     abs(quadct_S1[3] - quadct_S2[3]),
     abs(quadct_S1[4] - quadct_S2[4])
   )
-  idx <- which.max(c(
-    abs(quadct_S1[1] - quadct_S2[1]),
-    abs(quadct_S1[2] - quadct_S2[2]),
-    abs(quadct_S1[3] - quadct_S2[3]),
-    abs(quadct_S1[4] - quadct_S2[4])
-  ))
-  par(pty = "s")
[…]

### 2020-12-24 journalMS/article.tex
updated journal article

@@
 
-Using Monte Carlo simulation, Fasano and Franceschini approximated the distribution of D for significance testing -- as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
+Using Monte Carlo simulation, Fasano and Franceschini created a series of critical value look up tables of values of D -- as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
+Teukolsky and Press later defined an approximate fit to this data in equation \ref{eq:4} for quick computaion \cite{numericalRecipes}.
 
@@
 In their paper, Fasano and Franceschini use Monte Carlo simulation to approximate the distribution of D as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
-The resultant distributional approximation -- as described above in equation \ref{eq:4} -- can be used to test the significance levels for the two-dimensional K-S test.
+The resultant distributional approximation fit by Teukolsky and Press-- as described above in equation \ref{eq:4} -- can be used to test the significance levels for the two-dimensional K-S test.
 
-Note that the above formulas are accurate enough when $N \geq 20$, and when the indicated probability (significance level) is less than (more significant than) 0.20 or so.
-When the indicated probability is $> 0.20$, its value may not be accurate, but the implication that the data and model (or two data sets) are not significantly different is certainly correct.
+Note that the above formulas are accurate when $N \geq 20$, and when the indicated probability (significance level) is less than (more significant than) $\sim 0.20$.
+When the indicated probability is $> 0.20$, its value may not be accurate, but the implication that the two data sets are not significantly different is certainly correct.
 

### 2021-02-13 journalMS/article.tex
updated manuscript and figure captions

@@
 The \pkg{fasano.franceschini.test} package provides a two fold improvement over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): \textbf{(i)} The Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$.
-\textbf{(ii)} The package implements a parrellized bootstrapping procedure for improved significance testing.
+\textbf{(ii)} The package implements a  bootstrapping procedure for improved significance testing.
 }
@@
   \emph{and}\\
-  Northwestern Instutute on Complex Systems\\
+  Northwestern Institute on Complex Systems\\
   Northwestern University\\
@@
 
-The Kolmogorov--Smirnov (KS) is a non--parametric method for determining whether a sample is consistent with a given probability distribution.
-First derived by Kolmogorov and Smirnov in a series of papers~\citep{Kolmogorov1933,Kolmogorov1933a,Smirnov1936,Smirnov1937,Smirnov1939,Smirnov1944,Smirnov1948}, the was developed in 1-D.
-There is still a desire to be able to test 2-D and higher distributions against eachother.
-As a result, the first 2-D KS test was articulated by Peacock \citep{Peacock1983}.
+The Kolmogorov--Smirnov (KS) is a non--parametric, univariate statistical test designed to assess whether a set of data is consistent with a given probability distribution.
+First derived by Kolmogorov and Smirnov in a series of papers~\citep{Kolmogorov1933,Kolmogorov1933a,Smirnov1936,Smirnov1937,Smirnov1939,Smirnov1944,Smirnov1948}, the one-sample KS test defines the distribution of the quantity $D_{KS}$ -- the maximal absolute difference between the empirical cumulative distribution function of a set of values and a reference probability distribution.
+Kolmogorov and Smirnov's key insight was proving the distribution of the quantity $D_{KS}$ to be distribution free -- meaning the distribtuion does not depend on the underlying cumulative distribution functions being tested.
+Thus, the test can effectively be used to compare any univaraite empirical data distribution to any continious univariate reference distribution.
+A similar two-sample KS test could futher be used to compare any two univariate empirical data distributions against eachother to determine if they are drawn from the same underlying univariate distribution.
 
-The \pkg{fasano.franceschini.test} package is an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987}.
-This is a variant of the 2-D two-sample KS test as originally defined by Peacock \citep{Peacock1983}.
-In \proglang{R} \citep{R}, the only current implementation of the 2-D two-sample KS test is available in the \pkg{Peacock.test} package via the \fct{peacock2} function.
-The manuscript that this document is based on \citep{Fasano1987} then introduced hurdle.
-A \proglang{C} implementation of the Fasano Franceschini test is available in \citep{numericalRecipes}; however does not offer a bootstrapping procedure nor has been documentation for easy of use in \proglang{R} community.
+The versatility of the univaraite KS test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines(Cite uses in a broad range of field).
+Nonetheless, a generalization of the KS test as proposed by Kolmogorov and Smirnov does not naturally extend when working with data in more than one dimension.
+Fortunetly, a proposed solution to the dimensionality issue was articulated by Peacock \citep{Peacock1983} and later extended by Fasano and Franceschini \citep{Fasano1987} for an improved 2-D KS test.
+
+Currently, only the Peacock implementation of the 2-D two-sample KS test is available in \proglang{R} \citep{R} with the \pkg{Peacock.test} package via the \fct{peacock2} function; and has been shown to be markedly slower than the Fasano and Franceschini algorithm\citep{Lopes2007}.
+A \proglang{C} implementation of the Fasano Franceschini test is available in \citep{numericalRecipes}; however, arguments have been made to the validity of the implementation of the test not being distribution-free\citep{Babu2006}.
+Furthermore, the \proglang{C} implementation statistical testing is based on a fit to Monte Carlo Simulation and is only valid if the indicated probability (significance level) is less than (more significant than) $\sim 0.20$, which may present potential issue in some analysis.
+
+Here we present the \pkg{fasano.franceschini.test} package as an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987} The \pkg{fasano.franceschini.test} package provides a two fold improvement over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): \textbf{(i)} The Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$.
+\textbf{(ii)} The package implements a bootstrapping procedure for improved significance testing and mitigates the concerns of the distribution-free nature of the test brought up by\citep{Babu2006}.
 
@@
 
-The 1-D KS test can be divided into to variants of the same statistical procedure:
+As descibed previously, the 1-D KS test can be divided into two variants of the same statistical procedure:
 
@@
 \includegraphics{pdfvsCDF}
-\caption{\label{fig:kstest1D} Frequency distribution for number of days absent
-from school.}
+\caption{\label{fig:kstest1D} \textbf{LEFT:} Probability density function (PDF) of two normal distribution: orange sample1 $\mathcal{N}(\mu = 0,\,\sigma^{2} = 1)$; blue sample2 $\mathcal{N}(\mu = 5,\,\sigma^{2} = 1)$.
+\textbf{RIGHT:} Emperical cummlative density fuctions (eCDF) of the two PDFs in the left panel.
+Black dotted line represents the maximal absolute difference between the eCDFs ($D_{KS}$).
+}
 \end{figure}
 
-Kolmogorov and Smimov work defined the cumulative distribution for a random D-stat, which was later refined by Kendall and Stuart~\citep{Kendall1946}:
+Kolmogorov and Smimov work mathmatically defined the distribution for D to be distribution free, and further derived a table of citical values for hypthesis testing.
+Kolmogorov and Smimov work was later refined by Kendall and Stuart ~\citep{Kendall1946} to define the probability distribution for large values of n  ($n \geq 80$), having the asymptotic expression:
 
@@
 
-The 2-D two sample Fasano and Franceschini test is illustrated in the following example to compare the 2-D distributions of orange vs blue points (\textbf{Figure \ref{fig:kstest2D}}).
+The 2-D two sample Fasano and Franceschini test is illustrated in the following example to compare the 2-D distributions of orange versus blue points (\textbf{Figure \ref{fig:kstest2D}}).
 
@@
 \includegraphics{fftestOutput}
-\caption{\label{fig:kstest2D} Frequency distribution for number of days absent
-from school.}
+\caption{\label{fig:kstest2D} Calculate fraction of points in each quadrant (empirical cdf) Calculate maximal difference (D) between samples eCDF.
+}
 \end{figure}
@@
 
-\textbf{The 2-Sample cas}
+\textbf{The 2-Sample case}
 
@@
 In their paper, Fasano and Franceschini use Monte Carlo simulation to approximate the distribution of D as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
-The resultant distributional approximation fit by Teukolsky and Press-- as described above in equation \ref{eq:4} -- can be used to test the significance levels for the two-dimensional K-S test.
+The resultant distributional approximation fit by Teukolsky and Press -- as described above in equation \ref{eq:4} -- can be used to test the significance levels for the two-dimensional K-S test.
 

### 2021-02-16 journalMS/article.tex
updated manuscript and figure captions

@@
   NSF-Simons Center for Quantitative Biology,\\
-  Department of Engineering Sciences and Applied Mathematic,\\
+  Department of Engineering Sciences and Applied Mathematics,\\
   Department of Physics and Astronomy,\\
@@
 First derived by Kolmogorov and Smirnov in a series of papers~\citep{Kolmogorov1933,Kolmogorov1933a,Smirnov1936,Smirnov1937,Smirnov1939,Smirnov1944,Smirnov1948}, the one-sample KS test defines the distribution of the quantity $D_{KS}$ -- the maximal absolute difference between the empirical cumulative distribution function of a set of values and a reference probability distribution.
-Kolmogorov and Smirnov's key insight was proving the distribution of the quantity $D_{KS}$ to be distribution free -- meaning the distribtuion does not depend on the underlying cumulative distribution functions being tested.
-Thus, the test can effectively be used to compare any univaraite empirical data distribution to any continious univariate reference distribution.
-A similar two-sample KS test could futher be used to compare any two univariate empirical data distributions against eachother to determine if they are drawn from the same underlying univariate distribution.
+Kolmogorov and Smirnov's key insight was proving the distribution of the quantity $D_{KS}$ to be distribution free -- meaning the distribution does not depend on the underlying cumulative distribution functions being tested.
+Thus, the test can effectively be used to compare any univariate empirical data distribution to any continuous univariate reference distribution.
+A similar two-sample KS test could further be used to compare any two univariate empirical data distributions against each other to determine if they are drawn from the same underlying univariate distribution.
 
-The versatility of the univaraite KS test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines(Cite uses in a broad range of field).
+The versatility of the univariate KS test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines(Cite uses in a broad range of field).
 Nonetheless, a generalization of the KS test as proposed by Kolmogorov and Smirnov does not naturally extend when working with data in more than one dimension.
-Fortunetly, a proposed solution to the dimensionality issue was articulated by Peacock \citep{Peacock1983} and later extended by Fasano and Franceschini \citep{Fasano1987} for an improved 2-D KS test.
+Fortunately, a solution to the dimensionality issue was articulated by Peacock \citep{Peacock1983} and later extended by Fasano and Franceschini \citep{Fasano1987}.
 
@@
 A \proglang{C} implementation of the Fasano Franceschini test is available in \citep{numericalRecipes}; however, arguments have been made to the validity of the implementation of the test not being distribution-free\citep{Babu2006}.
-Furthermore, the \proglang{C} implementation statistical testing is based on a fit to Monte Carlo Simulation and is only valid if the indicated probability (significance level) is less than (more significant than) $\sim 0.20$, which may present potential issue in some analysis.
+Furthermore, the \proglang{C} implementation statistical testing is based on a fit to Monte Carlo Simulation and is only valid if the indicated probability (significance level) is less than (more significant than) $\sim 0.20$, which may present potential issues in some analysis.
 
 Here we present the \pkg{fasano.franceschini.test} package as an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987} The \pkg{fasano.franceschini.test} package provides a two fold improvement over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): \textbf{(i)} The Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$.
-\textbf{(ii)} The package implements a bootstrapping procedure for improved significance testing and mitigates the concerns of the distribution-free nature of the test brought up by\citep{Babu2006}.
+\textbf{(ii)} The package implements a bootstrapping procedure for improved significance testing and mitigates the concerns over the distribution-free nature of the test brought up by\citep{Babu2006}.
 
@@
 The Kolmogorov--Smirnov (KS) test is a non--parametric method for determining where a sample is consistent with a given probability distribution.
-While a comprehensive discussion pertaining the intracies of the 1-D KS test can be found elsewhere~\citep{Stephens1992a}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
+While a comprehensive discussion pertaining the intricacies of the 1-D KS test can be found elsewhere~\citep{Stephens1992a}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
 
-As descibed previously, the 1-D KS test can be divided into two variants of the same statistical procedure:
+As described previously, the 1-D KS test can be divided into two variants of the same statistical procedure:
 
@@
 
-The Kolmogorov-Smimov statistic (D) is the defined by the maximum difference between the cumulative density functions of the data and model (one--sample), or between the two data sets (two--sample).
+The Kolmogorov-Smirnov statistic (D) is the defined by the maximum absolute difference between the cumulative density functions of the data and model (one--sample), or between the two data sets (two--sample).
 In the following example, the orange and blue curve in the left panel represents the underlying distribution (pdf) for sample1 and sample2, respectively (\textbf{Figure \ref{fig:kstest1D}}).
@@
 \caption{\label{fig:kstest1D} \textbf{LEFT:} Probability density function (PDF) of two normal distribution: orange sample1 $\mathcal{N}(\mu = 0,\,\sigma^{2} = 1)$; blue sample2 $\mathcal{N}(\mu = 5,\,\sigma^{2} = 1)$.
-\textbf{RIGHT:} Emperical cummlative density fuctions (eCDF) of the two PDFs in the left panel.
+\textbf{RIGHT:} Empirical cumulative density functions (eCDF) of the two PDFs in the left panel.
 Black dotted line represents the maximal absolute difference between the eCDFs ($D_{KS}$).
@@
 
-Kolmogorov and Smimov work mathmatically defined the distribution for D to be distribution free, and further derived a table of citical values for hypthesis testing.
-Kolmogorov and Smimov work was later refined by Kendall and Stuart ~\citep{Kendall1946} to define the probability distribution for large values of n  ($n \geq 80$), having the asymptotic expression:
+Kolmogorov and Smirnov work mathematically defined the distribution for D to be distribution free, and further derived a table of critical values for hypothesis testing.
+Kolmogorov and Smirnov work was later refined by Kendall and Stuart ~\citep{Kendall1946} to define the integrated probability distribution for large values of n  ($n \geq 80$), having the asymptotic expression:
 
@@
 
-and thus allows for $p$-value hypothesis testing to determine if there is a significance difference between the cumulative density functions of the data and model (one-sample), or between the two data sets (two-sample).
+Given the maximum absolute difference (D) and \textbf{Equation~\ref{eq:1}}, one can  compute a $p$-value for  hypothesis testing to determine if there is a significance difference between the cumulative density functions of the data and model (one-sample), or between the two data sets (two-sample).
 
@@
 
-where $n1$ and $n2$ is the number of data points in the first and second samples respectively. A $p$-value $< 0.05$, indicates a significant difference between the samples.
+where $n_1$ and $n_2$ is the number of data points in the first and second samples respectively. A $p$-value $< 0.05$, indicates a significant difference between the samples.
 
@@
 This however, presents an issue as the commutative distribution function is not well defined in more than one dimension.
-In 2-D, there are 4 ways (3 independent ways) of defining the cummaltive distribution (\textbf{Figure \ref{fig:kstest2Dissue}}).
-In N dimensional space there are $2^{N}-1$ independant ways of defining the cumlative distribution function\citep{Peacock1983}.
+In 2-D, there are 4 ways (3 independent ways) of defining the cumulative distribution, since the direction in which we define choose to order the points in the x and y are arbitrary (\textbf{Figure \ref{fig:kstest2Dissue}}).
+In N dimensional space there are $2^{N}-1$ independent ways of defining the cumulative distribution function~\citep{Peacock1983}.
 
@@
 \includegraphics{CDF2Dissue}
-\caption{\label{fig:kstest2Dissue} 4 ways (3 independent ways) of defining the cummaltive distribution for a given point in 2-D.
-$P(x < X, y < Y) = 1 - P(x > X, y > Y); P(x < X, y > Y); P(x > X, y < Y)$}
+\caption{\label{fig:kstest2Dissue} 4 ways (3 independent ways) of defining the cumulative distribution for a given point in 2-D.
+$P(x < X, y < Y) + P(x < X, y > Y) + P(x > X, y < Y) = 1 - P(x > X, y > Y)$}
 \end{figure}
@@
 \subsection{Peacock Test (1983) and Fasano Franceschini Test (1987)}
-The Peacock test was the original generalization of the 1-D KS test to higher dimensions.
-Peacock solved the higher dimensionality issue by using the total intergrated probability -- i.e.
+Peacock solved the higher dimensionality issue by defining the 2-D test statistic as the largest difference between the empirical and theoretical commutative distributions, after taking all possible ordering combinations into account.
+Peacock test thus computes the total integrated probability -- i.e.
 fraction of data -- in each of the four quadrants around all possible points in the data.
-For example, for an given $n$ points in a two-dimensional space, the cumulative distribution functions is calculated in the $4n^2$ quadrants of the plane defined by all pairs ($X_i$ ,$Y_j$), $X_i$ and $Y_j$ being coordinates of any pairs of points in the given sample.
+For example, for a given $n$ points in a two-dimensional space, the cumulative distribution functions is calculated in the $4n^2$ quadrants of the plane defined by all pairs ($X_i$ ,$Y_j$), $X_i$ and $Y_j$ being coordinates of any pairs of points in the given sample.
 By ranging over all possible pairs of data points and quadrants, the 2-D D-stat is defined by the maximal difference of the integrated probabilities between samples.
@@
 The slight variation defined by Fasano and Franceschini was to only consider quadrants centered in each point of the given samples to compute the cumulative distribution functions.
-e.g, rather than looking over all ($X_i$ ,$Y_j$; where $i,j={1,n}$), only use the detected points ($X_i$ ,$Y_i$; where $i={1,n}$).
-Thus for an given $n$ points in a two-dimensional space, those $n$ points define $4n$ quadrants.
+e.g, rather than looking over all ($X_i$ ,$Y_j$; where $i,j={1,n}$), Fasano and Franceschini only used the detected points ($X_i$ ,$Y_i$; where $i={1,n}$).
+Thus for an given $n$ points in a two-dimensional space, those $n$ points define $4n$ -- rather than $4n^2$ -- quadrants.
 
@@
 \includegraphics{fftestOutput}
-\caption{\label{fig:kstest2D} Calculate fraction of points in each quadrant (empirical cdf) Calculate maximal difference (D) between samples eCDF.
+\caption{\label{fig:kstest2D} Illustration of the Fasano Franceschini algorithmic search for the maximal difference (D) between sample 2-D eCDFs.
+Looping through each point in the sampled data to define a unique origin (grey dotted line), the fraction of orange and blue points in each quadrants are computed (plot corners).
+For each origin, the quadrant which maximizes the absolute difference in the integrated probabilities is indicated.
+The origin which maximizes the overall absolute difference in the integrated probabilities between samples is highlighted by the orange box.
 }
@@
 The algorithm loops through each point in the orange sample in turn to define the origin of 4 quadrants (grey dotted line).
-The fraction of orange and blue points in each quadrant is computed (corners of plot) and the quadrant with the maximal difference is designated the current maximum for the specified origin.
+The fraction of orange and blue points in each quadrant is computed (corners of plot) and the quadrant with the maximal difference is designated with the current maximum for the specified origin.
 By ranging over all pairs of data points and quadrants, the 2-D D-stat is defined by the maximal difference of the integrated probabilities between samples.
@@
 Using Monte Carlo simulation, Fasano and Franceschini created a series of critical value look up tables of values of D -- as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
-Teukolsky and Press later defined an approximate fit to this data in equation \ref{eq:4} for quick computaion \cite{numericalRecipes}.
+Teukolsky and Press later defined an approximate fit to this data in equation \ref{eq:4} for quick computation \cite{numericalRecipes}.
 
@@
 In their paper, Fasano and Franceschini use Monte Carlo simulation to approximate the distribution of D as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
-The resultant distributional approximation fit by Teukolsky and Press -- as described above in equation \ref{eq:4} -- can be used to test the significance levels for the two-dimensional K-S test.
-
-Note that the above formulas are accurate when $N \geq 20$, and when the indicated probability (significance level) is less than (more significant than) $\sim 0.20$.
-When the indicated probability is $> 0.20$, its value may not be accurate, but the implication that the two data sets are not significantly different is certainly correct.
+Unlike the 1-D KS test, the 2-D KS test is not rigorously true that the distribution of D is independent of the shape of the 2-D distribution.
+Fasano and Franceschini reasoned that the coefficient of correlation between the two-dimensional distributions is what governs the resultant distribution of D.
+For example, if the data sets are perfectly correlated ($r = 1$), the 2-D plane flattens to a single line and thus a 1-D KS test could be used.
+If the data sets are perfectly uncorrelated ($r = 0$), the 2-D distribution is independent in the x and y directions and thus one could use a 1-D KS test on the marginal x or y distribution of the 2-D plane.
+Results from Monte Carlo simulation support these assumptions -- showing that the distribution of D is nearly identical for varying distributions, assuming they have the same coefficient of correlation r ~\citep{Fasano1987}.
+The resultant distributional approximation fit by Teukolsky and Press -- as described above in equation \ref{eq:4} -- can be used to test the significance levels for the 2-D K-S test using the following code.
 
@@
 
-Some arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
+As a caveat, to the fit produced by Teukolsky and Press are accurate when $N \geq 20$, and the $p$-value is less than (more significant than) $\sim 0.20$.
+For $p$-value that are $> 0.20$, the distributional approximation may not be accurate.
+While the assertion the two data sets are \textit{not} significantly different is surely correct; analysis looking to compare $p$-value across experiments conditions -- such as if one was to do a cross study concordant analysis comparing rank order $p$-value (Cite bmal1 science Article or TimeTrial) -- would not be possible with the Teukolsky and Press approximate fit.
+Moreover, additional arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
 
@@
 The \pkg{fasano.franceschini.test} R package, implements a parallelized bootstrapping procedure.
-The marginal distribution from 2-dimensional data set is resampled with replacements to generate randomized 2-dimensional data set \code{nBootStrap} times.
+The marginal distribution from 2-dimensional data set is resampled with replacements to generate randomized 2-dimensional data sets \code{nBootStrap} times.
 The frequency count by quadrant is performed for each bootstrapped resampling as described above to compute the D-stats.
-The observed D-stat is then compared to the bootstrapped D-stats to compute a $p$-value.
-See \fct{fasano.franceschini.test} for source code details and implementation.
-
-Note, bootstrapping may significantly increase run time performance.
-To improve run time, adjust the \code{cores} parameter to specify the number of parallel operations to run.
-See the R \pkg{parrellel} package and the \fct{mclapply} function for further details.
+The observed D-stat is then compared to the distribution of bootstrapped D-stats to compute a $p$-value.
+The bootstrapped version of the Fasano Franceschini test can be implemented with the following lines of code -- see \fct{fasano.franceschini.test} for further source code details and implementation.
 
@@
 
+Note, bootstrapping may significantly increase run time performance.
+To improve run time, adjust the \code{cores} parameter to specify the number of parallel operations to run.
+See the R \pkg{parrellel} package and the \fct{mclapply} function for further details.
+
 \subsection{Computational Efficency}
@@
 \begin{itemize}
-\item The Fasano Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as dataset sizes scale.
+\item The Fasano Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale.
 \item The Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
@@
 Improvements:
-  (i) Fast algorith
-  (ii) Prallelized Bootstrapping Procedure for Improved statisitcal testing
+  (i) Fast algorithm
+  (ii) Prallelized Bootstrapping Procedure for Improved statistical testing
 

### 2021-02-16 journalMS/article.tex
updated D_KS and D_FF consistency throughout paper

@@
 \usepackage{thumbpdf,lmodern}
+\usepackage{amssymb}
 
@@
 
-The Kolmogorov-Smirnov statistic (D) is the defined by the maximum absolute difference between the cumulative density functions of the data and model (one--sample), or between the two data sets (two--sample).
+The Kolmogorov-Smirnov statistic ($D_{KS}$) is the defined by the maximum absolute difference between the cumulative density functions of the data and model (one--sample), or between the two data sets (two--sample).
 In the following example, the orange and blue curve in the left panel represents the underlying distribution (pdf) for sample1 and sample2, respectively (\textbf{Figure \ref{fig:kstest1D}}).
@@
 
-Kolmogorov and Smirnov work mathematically defined the distribution for D to be distribution free, and further derived a table of critical values for hypothesis testing.
+Kolmogorov and Smirnov work mathematically defined the distribution for $D_{KS}$ to be distribution free, and further derived a table of critical values for hypothesis testing.
 Kolmogorov and Smirnov work was later refined by Kendall and Stuart ~\citep{Kendall1946} to define the integrated probability distribution for large values of n  ($n \geq 80$), having the asymptotic expression:
@@
 
-Given the maximum absolute difference (D) and \textbf{Equation~\ref{eq:1}}, one can  compute a $p$-value for  hypothesis testing to determine if there is a significance difference between the cumulative density functions of the data and model (one-sample), or between the two data sets (two-sample).
+Given the maximum absolute difference ($D_{KS}$) and \textbf{Equation~\ref{eq:1}}, one can  compute a $p$-value for  hypothesis testing to determine if there is a significance difference between the cumulative density functions of the data and model (one-sample), or between the two data sets (two-sample).
 
-\textbf{For the 1-Sample case}, the $p$-value probability of a given D-stat is defined by
+\textbf{For the 1-Sample case}, the $p$-value probability of a given D is defined by
 
@@
 
-\textbf{In the 2-Sample case}, the $p$-value probability of a given D-stat is defined by
+\textbf{In the 2-Sample case}, the $p$-value probability of a given D is defined by
 \begin{equation} \label{eq:3}
@@
 \includegraphics{fftestOutput}
-\caption{\label{fig:kstest2D} Illustration of the Fasano Franceschini algorithmic search for the maximal difference (D) between sample 2-D eCDFs.
+\caption{\label{fig:kstest2D} Illustration of the Fasano Franceschini algorithmic search for the maximal difference ($D_{FF}$) between sample 2-D eCDFs.
 Looping through each point in the sampled data to define a unique origin (grey dotted line), the fraction of orange and blue points in each quadrants are computed (plot corners).
@@
 The fraction of orange and blue points in each quadrant is computed (corners of plot) and the quadrant with the maximal difference is designated with the current maximum for the specified origin.
-By ranging over all pairs of data points and quadrants, the 2-D D-stat is defined by the maximal difference of the integrated probabilities between samples.
+By ranging over all pairs of data points and quadrants, the 2-D $D_{FF}$ is defined by the maximal difference of the integrated probabilities between samples.
 In this case using the orange point as the origin, the maximal difference is $D = 0.60$.
 
-This process is repeated using the blue points as the origins to compute the maximal D-stat.
-Finally, both maximal D-stats using the orange and blue points as origins are then averaged to compute the overall D-stat for hypothesis testing.
+This process is repeated using the blue points as the origins to compute the maximal $D_{FF}$.
+Finally, both maximal D-stats using the orange and blue points as origins are then averaged to compute the overall $D_{FF}$ for hypothesis testing.
 
@@
 
-Using Monte Carlo simulation, Fasano and Franceschini created a series of critical value look up tables of values of D -- as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
+Using Monte Carlo simulation, Fasano and Franceschini created a series of critical value look up tables of values of $D_{FF}$ -- as a function of $D_{FF}$, the sample size $N$, and the coefficient of correlation $r$.
 Teukolsky and Press later defined an approximate fit to this data in equation \ref{eq:4} for quick computation \cite{numericalRecipes}.
@@
 In their paper, Fasano and Franceschini use Monte Carlo simulation to approximate the distribution of D as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
-Unlike the 1-D KS test, the 2-D KS test is not rigorously true that the distribution of D is independent of the shape of the 2-D distribution.
-Fasano and Franceschini reasoned that the coefficient of correlation between the two-dimensional distributions is what governs the resultant distribution of D.
+Unlike the 1-D KS test, the 2-D KS test is not rigorously true that the distribution of $D_{FF}$ is independent of the shape of the 2-D distribution.
+Fasano and Franceschini reasoned that the coefficient of correlation between the two-dimensional distributions is what governs the resultant distribution of $D_{FF}$.
 For example, if the data sets are perfectly correlated ($r = 1$), the 2-D plane flattens to a single line and thus a 1-D KS test could be used.
@@
 
-As a caveat, to the fit produced by Teukolsky and Press are accurate when $N \geq 20$, and the $p$-value is less than (more significant than) $\sim 0.20$.
+As a caveat, to the fit produced by Teukolsky and Press are accurate when $N \gtrsim 20$, and the $p$-value is less than (more significant than) $\sim 0.20$.
 For $p$-value that are $> 0.20$, the distributional approximation may not be accurate.
-While the assertion the two data sets are \textit{not} significantly different is surely correct; analysis looking to compare $p$-value across experiments conditions -- such as if one was to do a cross study concordant analysis comparing rank order $p$-value (Cite bmal1 science Article or TimeTrial) -- would not be possible with the Teukolsky and Press approximate fit.
+While the assertion the two data sets are \textit{not} significantly different is surely correct; analysis looking to compare $p$-value across experiments conditions -- such as if one was to do a cross study concordant analysis comparing rank order $p$-value \cite{Ness}
+%(Cite bmal1 science Article or TimeTrial)
+-- would not be possible with the Teukolsky and Press approximate fit.
 Moreover, additional arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
@@
 As Fasano and Franceschini's paper was originally released in 1987, this approach was unfeasible at scale.
-However, as we are currently in the age of high-performance computers -- a null distribution of D-stats can be bootstrapped from the data to test significance.
+However, as we are currently in the age of high-performance computers -- a null distribution of $D_{FF}$ can be bootstrapped from the data to test significance.
 
@@
 The marginal distribution from 2-dimensional data set is resampled with replacements to generate randomized 2-dimensional data sets \code{nBootStrap} times.
-The frequency count by quadrant is performed for each bootstrapped resampling as described above to compute the D-stats.
-The observed D-stat is then compared to the distribution of bootstrapped D-stats to compute a $p$-value.
+The frequency count by quadrant is performed for each bootstrapped resampling as described above to compute the $D_{FF}$.
+The observed $D_{FF}$ is then compared to the distribution of bootstrapped $D_{FF}$ to compute a $p$-value.
 The bootstrapped version of the Fasano Franceschini test can be implemented with the following lines of code -- see \fct{fasano.franceschini.test} for further source code details and implementation.

### 2021-02-16 journalMS/article.tex
updated reference list for additional citations

@@
 For $p$-value that are $> 0.20$, the distributional approximation may not be accurate.
-While the assertion the two data sets are \textit{not} significantly different is surely correct; analysis looking to compare $p$-value across experiments conditions -- such as if one was to do a cross study concordant analysis comparing rank order $p$-value \cite{Ness}
+While the assertion the two data sets are \textit{not} significantly different is surely correct; analysis looking to compare $p$-value across experiments conditions -- such as if one was to do a cross study concordant analysis comparing rank order $p$-value \citep{Ness-Cohn2020}
 %(Cite bmal1 science Article or TimeTrial)

### 2021-03-05 journalMS/article.tex
RB mods and notes in progress

@@
 \newcommand{\fct}[1]{\code{#1()}}
+
+\newcommand{\RBnote}[1]{\textcolor{red}{#1}}
 
@@
 \Abstract{
+\RBnote{Are citations allowed in the abstract (per journal style)?}
 The \pkg{fasano.franceschini.test} package is an  \proglang{R} implementation of the 2-D Kolmogorov-Smirnov (KS) two--sample test as defined by Fasano and Franceschini \citep{Fasano1987}.
-This is a variant of the 2-D two-sample KS test as originally defined by Peacock \citep{Peacock1983}, previously implemented in \proglang{R} in the \pkg(Peacock.test) package \citep{Xiao2016}.
-The \pkg{fasano.franceschini.test} package provides a two fold improvement over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): \textbf{(i)} The Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$.
-\textbf{(ii)} The package implements a  bootstrapping procedure for improved significance testing.
+This is a variant of the 2-D two-sample KS test as originally defined by Peacock \citep{Peacock1983}, previously implemented in \proglang{R} in the \pkg{Peacock.test} package \citep{Xiao2016}.
+The \pkg{fasano.franceschini.test} package provides two improvements over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): (i) the Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$; and (ii) the package implements a  bootstrapping procedure for improved significance testing.
+\RBnote{Summarize what the paper will cover.}
 }
@@
   Elan Ness-Cohn\\
-  Department of Molecular Biosciences,\\
-  Biostatistics Division of The Department of Preventive Medicine,\\
+  Department of Molecular Biosciences\\
   \emph{and}\\
@@
   Department of Molecular Biosciences,\\
-  Biostatistics Division of The Department of Preventive Medicine,\\
-  NSF-Simons Center for Quantitative Biology,\\
   Department of Engineering Sciences and Applied Mathematics,\\
   Department of Physics and Astronomy,\\
+  Northwestern Institute on Complex Systems,\\
   \emph{and}\\
-  Northwestern Institute on Complex Systems\\
+  NSF-Simons Center for Quantitative Biology\\
   Northwestern University\\
@@
 
+\RBnote{Please use nonbreaking spaces (tilde rather than space) before all cite commands}
+
 The Kolmogorov--Smirnov (KS) is a non--parametric, univariate statistical test designed to assess whether a set of data is consistent with a given probability distribution.
-First derived by Kolmogorov and Smirnov in a series of papers~\citep{Kolmogorov1933,Kolmogorov1933a,Smirnov1936,Smirnov1937,Smirnov1939,Smirnov1944,Smirnov1948}, the one-sample KS test defines the distribution of the quantity $D_{KS}$ -- the maximal absolute difference between the empirical cumulative distribution function of a set of values and a reference probability distribution.
-Kolmogorov and Smirnov's key insight was proving the distribution of the quantity $D_{KS}$ to be distribution free -- meaning the distribution does not depend on the underlying cumulative distribution functions being tested.
+First derived by Kolmogorov and Smirnov in a series of papers~\citep{Kolmogorov1933,Kolmogorov1933a,Smirnov1936,Smirnov1937,Smirnov1939,Smirnov1944,Smirnov1948}, the one-sample KS test defines the distribution of the quantity $D_{KS}$, the maximal absolute difference between the empirical cumulative distribution function (CDF) of a set of values and a reference probability distribution.
+Kolmogorov and Smirnov's key insight was proving the distribution of $D_{KS}$ was independent of the CDFs being tested.
 Thus, the test can effectively be used to compare any univariate empirical data distribution to any continuous univariate reference distribution.
-A similar two-sample KS test could further be used to compare any two univariate empirical data distributions against each other to determine if they are drawn from the same underlying univariate distribution.
+A two-sample KS test could further be used to compare any two univariate empirical data distributions against each other to determine if they are drawn from the same underlying univariate distribution.
 
-The versatility of the univariate KS test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines(Cite uses in a broad range of field).
+The versatility of the univariate KS test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines \RBnote{highlighting your note: (Cite uses in a broad range of field)}.
 Nonetheless, a generalization of the KS test as proposed by Kolmogorov and Smirnov does not naturally extend when working with data in more than one dimension.
-Fortunately, a solution to the dimensionality issue was articulated by Peacock \citep{Peacock1983} and later extended by Fasano and Franceschini \citep{Fasano1987}.
+Fortunately, a solution to the dimensionality issue was articulated by Peacock~\citep{Peacock1983} and later extended by Fasano and Franceschini~\citep{Fasano1987}.
 
-Currently, only the Peacock implementation of the 2-D two-sample KS test is available in \proglang{R} \citep{R} with the \pkg{Peacock.test} package via the \fct{peacock2} function; and has been shown to be markedly slower than the Fasano and Franceschini algorithm\citep{Lopes2007}.
+Currently, only the Peacock implementation of the 2-D two-sample KS test is available in \proglang{R}~\citep{R} with the \pkg{Peacock.test} package via the \fct{peacock2} function, but this has been shown to be markedly slower than the Fasano and Franceschini algorithm\citep{Lopes2007}.
 A \proglang{C} implementation of the Fasano Franceschini test is available in \citep{numericalRecipes}; however, arguments have been made to the validity of the implementation of the test not being distribution-free\citep{Babu2006}.
-Furthermore, the \proglang{C} implementation statistical testing is based on a fit to Monte Carlo Simulation and is only valid if the indicated probability (significance level) is less than (more significant than) $\sim 0.20$, which may present potential issues in some analysis.
+Furthermore, the \proglang{C} implementation statistical testing is based on a fit to Monte Carlo Simulation and is only valid for significance levels $\alpha \lessapprox 0.20$, which may present potential issues in some analysis.
 
-Here we present the \pkg{fasano.franceschini.test} package as an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987} The \pkg{fasano.franceschini.test} package provides a two fold improvement over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): \textbf{(i)} The Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$.
-\textbf{(ii)} The package implements a bootstrapping procedure for improved significance testing and mitigates the concerns over the distribution-free nature of the test brought up by\citep{Babu2006}.
+Here we present the \pkg{fasano.franceschini.test} package as an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987} The \pkg{fasano.franceschini.test} package provides two improvements over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): (i) the Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$; and (ii) the package implements a bootstrapping procedure for improved significance testing and mitigates the concerns over the distribution-free nature of the test brought up by~\citep{Babu2006}.
 
@@
 
-The Kolmogorov--Smirnov (KS) test is a non--parametric method for determining where a sample is consistent with a given probability distribution.
-While a comprehensive discussion pertaining the intricacies of the 1-D KS test can be found elsewhere~\citep{Stephens1992a}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
+\RBnote{This can be shortened substantially.  I have done so, but please look closely.  In general, avoid writing textbook material in papers.}
 
-As described previously, the 1-D KS test can be divided into two variants of the same statistical procedure:
-
-\begin{leftbar}
-\textbf{The One-Sample Test}
-\begin{itemize}
-\item Compares the data to an analytical distribution.
-\end{itemize}
-\textbf{The Two-Sample Test}
-\begin{itemize}
-\item Compares two data samples to determine if they are drawn from the same underlying probability distribution.
-\end{itemize}
-\end{leftbar}
-
-In both cases, the KS test works as follows:
-
-The Kolmogorov-Smirnov statistic ($D_{KS}$) is the defined by the maximum absolute difference between the cumulative density functions of the data and model (one--sample), or between the two data sets (two--sample).
+The Kolmogorov--Smirnov (KS) test is a non--parametric method for determining whether a sample is consistent with a given probability distribution.
+%While a comprehensive discussion pertaining the intricacies of the 1-D KS test can be found elsewhere~\citep{Stephens1992a}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
+%
+%As described previously, the 1-D KS test can be divided into two variants of the same statistical procedure:
+%
+%\begin{leftbar}
+%\textbf{The One-Sample Test}
+%\begin{itemize}
+%\item Compares the data to analytical distribution.
+%\end{itemize}
+%\textbf{The Two-Sample Test}
+%\begin{itemize}
+%\item Compares two data samples to determine if they are drawn from the same underlying probability distribution.
+%\end{itemize}
+%\end{leftbar}
+%
+%In both cases, the KS test works as follows:
+%
+In one dimension, 
+the Kolmogorov-Smirnov statistic ($D_{KS}$) is the defined by the maximum absolute difference between the cumulative density functions of the data and model (one--sample), or between the two data sets (two--sample),
+as illustrated in (\textbf{Figure~\ref{fig:kstest1D}}).
+%.
 In the following example, the orange and blue curve in the left panel represents the underlying distribution (pdf) for sample1 and sample2, respectively (\textbf{Figure \ref{fig:kstest1D}}).
@@
 
-Kolmogorov and Smirnov work mathematically defined the distribution for $D_{KS}$ to be distribution free, and further derived a table of critical values for hypothesis testing.
+%Kolmogorov and Smirnov work mathematically defined the distribution for $D_{KS}$ to be distribution free, and further derived a table of critical values for hypothesis testing.
 Kolmogorov and Smirnov work was later refined by Kendall and Stuart ~\citep{Kendall1946} to define the integrated probability distribution for large values of n  ($n \geq 80$), having the asymptotic expression:
 
+In the large--sample limit ($n \geq 80$), it can be shown~\citep{Kendall1946} that $D_{KS}$ converges in distribution to 
 \begin{equation} \label{eq:1}
-\Phi(\lambda) = 2 \sum_{k=1}^{\infty} -1^{k-1}e^{-2k^2\lambda^2}
+\Phi(\lambda) = 2 \sum_{k=1}^{\infty} -1^{k-1}e^{-2k^2\lambda^2} \,.
 \end{equation}
-
-Given the maximum absolute difference ($D_{KS}$) and \textbf{Equation~\ref{eq:1}}, one can  compute a $p$-value for  hypothesis testing to determine if there is a significance difference between the cumulative density functions of the data and model (one-sample), or between the two data sets (two-sample).
-
-\textbf{For the 1-Sample case}, the $p$-value probability of a given D is defined by
-
+%Given the maximum absolute difference ($D_{KS}$) and \textbf{Equation~\ref{eq:1}}, one can  compute a $p$-value for  hypothesis testing to determine if there is a significance difference between the cumulative density functions of the data and model (one-sample), or between the two data sets (two-sample).
+%
+%\textbf{For the 1-Sample case}, the $p$-value probability of a given D is defined by
+In the one-sample case, the $p$-value is given by 
 \begin{equation} \label{eq:2}
-Prob(D > observed) = \Phi ( D\sqrt{N})
+\mathbb{P}(D > observed) = \Phi ( D\sqrt{n})\,;
 \end{equation}
-
-where $N$ is the number of data points sampled. A $p$-value $< 0.05$, indicates a significant difference between the samples.
-
-\textbf{In the 2-Sample case}, the $p$-value probability of a given D is defined by
+in the two-sample case, the $p$-value is given by
+%where $N$ is the number of data points sampled. A $p$-value $< 0.05$, indicates a significant difference between the samples.
+%
+%\textbf{In the 2-Sample case}, the $p$-value probability of a given D is defined by
 \begin{equation} \label{eq:3}
-Prob(D > observed) = \Phi ( D\sqrt{\frac{n_1n_2}{n_1+n_2}})
+\mathbb{P}(D > observed) = \Phi ( D\sqrt{\frac{n_1n_2}{n_1+n_2}})\,.
 \end{equation}
-
-where $n_1$ and $n_2$ is the number of data points in the first and second samples respectively. A $p$-value $< 0.05$, indicates a significant difference between the samples.
+%
+%where $n_1$ and $n_2$ is the number of data points in the first and second samples respectively. A $p$-value $< 0.05$, indicates a significant difference between the samples.
 
@@
 
-Similar to the 1-D KS test, the obvious solution to extend the theory into higher dimensional would be to define the maximal cumulative difference between the two 2-D distributions.
-This however, presents an issue as the commutative distribution function is not well defined in more than one dimension.
+%Similar to the 1-D KS test, the obvious solution to extend the theory into higher dimensional would be to define the maximal cumulative difference between the two 2-D distributions.
+This however, presents an issue as the cumulative distribution function is not well defined in more than one dimension.
 In 2-D, there are 4 ways (3 independent ways) of defining the cumulative distribution, since the direction in which we define choose to order the points in the x and y are arbitrary (\textbf{Figure \ref{fig:kstest2Dissue}}).
 In N dimensional space there are $2^{N}-1$ independent ways of defining the cumulative distribution function~\citep{Peacock1983}.
+Extending the above to two or higher dimension is complicated by the fact that CDFs are not well-defined in more than one dimension.
+In 2-D, there are 4 ways (3 independent ways) of defining the cumulative distribution, since the direction in which we order the $x$ and $y$ points are arbitrary (\textbf{Figure \ref{fig:kstest2Dissue}}); more generally, in $P$-dimensional space there are $2^{P}-1$ independent ways of defining the cumulative distribution function~\citep{Peacock1983}.
 
@@
 \includegraphics{CDF2Dissue}
-\caption{\label{fig:kstest2Dissue} 4 ways (3 independent ways) of defining the cumulative distribution for a given point in 2-D.
-$P(x < X, y < Y) + P(x < X, y > Y) + P(x > X, y < Y) = 1 - P(x > X, y > Y)$}
+\caption{\label{fig:kstest2Dissue} Four ways (3 independent) of defining the cumulative distribution for a given point in 2-D.
+%$P(x < X, y < Y) + P(x < X, y > Y) + P(x > X, y < Y) = 1 - P(x > X, y > Y)$
+}
 \end{figure}
@@
 \subsection{Peacock Test (1983) and Fasano Franceschini Test (1987)}
-Peacock solved the higher dimensionality issue by defining the 2-D test statistic as the largest difference between the empirical and theoretical commutative distributions, after taking all possible ordering combinations into account.
-Peacock test thus computes the total integrated probability -- i.e.
-fraction of data -- in each of the four quadrants around all possible points in the data.
+Peacock solved the higher dimensionality issue by defining the 2-D test statistic as the largest difference between the empirical and theoretical cumulative distributions, after taking all possible ordering combinations into account.
+Peacock's test thus computes the total integrated probability -- i.e.\,fraction of data -- in each of the four quadrants around all possible points in the data.
 For example, for a given $n$ points in a two-dimensional space, the cumulative distribution functions is calculated in the $4n^2$ quadrants of the plane defined by all pairs ($X_i$ ,$Y_j$), $X_i$ and $Y_j$ being coordinates of any pairs of points in the given sample.
-By ranging over all possible pairs of data points and quadrants, the 2-D D-stat is defined by the maximal difference of the integrated probabilities between samples.
+By ranging over all possible pairs of data points and quadrants, the 2-dimensional $D$ statistic is defined by the maximal difference of the integrated probabilities between samples.
 
-The slight variation defined by Fasano and Franceschini was to only consider quadrants centered in each point of the given samples to compute the cumulative distribution functions.
-e.g, rather than looking over all ($X_i$ ,$Y_j$; where $i,j={1,n}$), Fasano and Franceschini only used the detected points ($X_i$ ,$Y_i$; where $i={1,n}$).
-Thus for an given $n$ points in a two-dimensional space, those $n$ points define $4n$ -- rather than $4n^2$ -- quadrants.
+The variation defined by Fasano and Franceschini was to only consider quadrants centered in each point of the given samples to compute the cumulative distribution functions.
+Rather than looking over all $X_i, Y_j: i,j={1,n}$, Fasano and Franceschini only used the detected points $X_i, Y_i : i={1,n}$.
+Thus for any given $n$ points in a two-dimensional space, those $n$ points define $4n$ -- rather than $4n^2$ -- quadrants.
 
 The 2-D two sample Fasano and Franceschini test is illustrated in the following example to compare the 2-D distributions of orange versus blue points (\textbf{Figure \ref{fig:kstest2D}}).
+
+\RBnote{Something is amiss with the way the points are counted!}
 
@@
 The origin which maximizes the overall absolute difference in the integrated probabilities between samples is highlighted by the orange box.
+\RBnote{Something is amiss with the way the points are counted!}
 }
@@
 The fraction of orange and blue points in each quadrant is computed (corners of plot) and the quadrant with the maximal difference is designated with the current maximum for the specified origin.
-By ranging over all pairs of data points and quadrants, the 2-D $D_{FF}$ is defined by the maximal difference of the integrated probabilities between samples.
-In this case using the orange point as the origin, the maximal difference is $D = 0.60$.
+By iterating over all pairs of data points and quadrants, the $D_{FF}$ is defined by the maximal difference of the integrated probabilities between samples.
+In this case using the orange point as the origin, the maximal difference is $D_{FF} = 0.60$.
+\RBnote{How are tied values counted? You have strict inequalities in the figure\dots}
 
+\RBnote{Since $D_{FF}$ from one sample isn't the final $D_{FF}$, you should note it appropriately (eg, $D_{FF,1}$ and $D_{FF,2}$).}
 This process is repeated using the blue points as the origins to compute the maximal $D_{FF}$.
-Finally, both maximal D-stats using the orange and blue points as origins are then averaged to compute the overall $D_{FF}$ for hypothesis testing.
+Finally, both maximal $D_{FF}$ statistics using the orange and blue points as origins are then averaged to compute the overall $D_{FF}$ for hypothesis testing.
 
-\subsection{Defining the Fasano and Franceschini Null Distribution}
+\subsection{Defining the Fasano and Franceschini null distribution}
 

### 2021-03-06 R/interal_functions_FasFran.R
updated test to handle ties

@@ -16,28 +16,31 @@
 #'
 #'
 quadCount <- function(x_origin, y_origin, x, y) {
-  n <- length(x)
-  nq1 <- 0
-  nq2 <- 0
-  nq3 <- 0
-  nq4 <- 0
 
-  # think about less than or equal to?
-  # i.e. ties
-  for (k in 1:n) {
-    if (y[k] >= y_origin) {
-      if (x[k] >= x_origin) {
-        nq1 <- nq1 + 1
-      } else {
-        nq2 <- nq2 + 1
-      }
-    } else {
-      if (x[k] >= x_origin) {
-        nq4 <- nq4 + 1
-      } else {
-        nq3 <- nq3 + 1
-      }
-    }
+  # sum number of points in each quadrant (counter clockwise from 1 to 4)
+  nq1 <- sum(y > y_origin & x > x_origin)
+  nq2 <- sum(y > y_origin & x < x_origin)
+  nq4 <- sum(y < y_origin & x > x_origin)
+  nq3 <- sum(y < y_origin & x < x_origin)
+
+  #origin point is never taken into account in the previous logic operations
+  #remove all points that fall on the origin
+  n <- length(x) - sum(y == y_origin & x == x_origin)
+
+  #if there are any ties that are NOT the origin
+  #i.e. a point is exactly on the line between two quadrants
+  #split the point count equally between the quadrants
+  if(any(xor(x == x_origin, y == y_origin))){
+    yties_12 <- sum(y > y_origin & x == x_origin)
+    yties_34 <- sum(y < y_origin & x == x_origin)
+
+    xties_14 <- sum(y == y_origin & x > x_origin)
+    xties_23 <- sum(y == y_origin & x < x_origin)
+
+    nq1 <- nq1 + 0.5*xties_14 + 0.5*yties_12
+    nq2 <- nq2 + 0.5*xties_23 + 0.5*yties_12
+    nq4 <- nq4 + 0.5*xties_14 + 0.5*yties_34
+    nq3 <- nq3 + 0.5*xties_23 + 0.5*yties_34
   }
 
   freqQuad1 <- nq1 / n
@@ -47,6 +50,7 @@ quadCount <- function(x_origin, y_origin, x, y) {
 
   return(c(freqQuad1, freqQuad2, freqQuad3, freqQuad4))
 }
+
 
 #' Get KS Stat
 #'
[…]

### 2021-03-22 journalMS/article.tex
RB edits

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Fasano Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in \proglang{R}}
-\Plaintitle{Fasano Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in R}
+\title{Fasano--Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in \proglang{R}}
+\Plaintitle{Fasano--Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in R}
 \Shorttitle{2-Dimensional Kolmogorov--Smirnov in \proglang{R}}
@@
 Currently, only the Peacock implementation of the 2-D two-sample KS test is available in \proglang{R}~\citep{R} with the \pkg{Peacock.test} package via the \fct{peacock2} function, but this has been shown to be markedly slower than the Fasano and Franceschini algorithm\citep{Lopes2007}.
-A \proglang{C} implementation of the Fasano Franceschini test is available in \citep{numericalRecipes}; however, arguments have been made to the validity of the implementation of the test not being distribution-free\citep{Babu2006}.
+A \proglang{C} implementation of the Fasano--Franceschini test is available in \citep{numericalRecipes}; however, arguments have been made to the validity of the implementation of the test not being distribution-free\citep{Babu2006}.
 Furthermore, the \proglang{C} implementation statistical testing is based on a fit to Monte Carlo Simulation and is only valid for significance levels $\alpha \lessapprox 0.20$, which may present potential issues in some analysis.
@@
 
-\subsection{Peacock Test (1983) and Fasano Franceschini Test (1987)}
+\subsection{Peacock Test (1983) and Fasano--Franceschini Test (1987)}
 Peacock solved the higher dimensionality issue by defining the 2-D test statistic as the largest difference between the empirical and theoretical cumulative distributions, after taking all possible ordering combinations into account.
@@
 
-\RBnote{Something is amiss with the way the points are counted!}
-
 \begin{figure}[t!]
@@
 \includegraphics{fftestOutput}
-\caption{\label{fig:kstest2D} Illustration of the Fasano Franceschini algorithmic search for the maximal difference ($D_{FF}$) between sample 2-D eCDFs.
+\caption{\label{fig:kstest2D} Illustration of the Fasano--Franceschini algorithmic search for the maximal difference ($D_{FF}$) between sample 2-D eCDFs.
 Looping through each point in the sampled data to define a unique origin (grey dotted line), the fraction of orange and blue points in each quadrants are computed (plot corners).
@@
 The origin which maximizes the overall absolute difference in the integrated probabilities between samples is highlighted by the orange box.
-\RBnote{Something is amiss with the way the points are counted!}
 }
@@
 By iterating over all pairs of data points and quadrants, the $D_{FF}$ is defined by the maximal difference of the integrated probabilities between samples.
-In this case using the orange point as the origin, the maximal difference is $D_{FF} = 0.60$.
-\RBnote{How are tied values counted? You have strict inequalities in the figure\dots}
+In this case using the orange point as the origin, the maximal difference is $D_{FF} = 0.52$.
+\RBnote{How are tied values counted? Describe that here.  Also discuss the treatment of the origin.}
 
@@
 
-Using Monte Carlo simulation, Fasano and Franceschini created a series of critical value look up tables of values of $D_{FF}$ -- as a function of $D_{FF}$, the sample size $N$, and the coefficient of correlation $r$.
-Teukolsky and Press later defined an approximate fit to this data in equation \ref{eq:4} for quick computation \cite{numericalRecipes}.
+Using Monte Carlo simulation, Fasano and Franceschini created a look-up table of critical values of $D_{FF}$ as a function of $D_{FF}$, the sample size $N$, and the coefficient of correlation $r$.
+Teukolsky and Press later defined an approximate fit to this data in equation \ref{eq:4} for quick computation~\cite{numericalRecipes}, as follows.
 
-\textbf{For the 1-Sample case}
-
+For the 1-Sample case,
 \begin{equation} \label{eq:4}
-Prob(D > observed) = \Phi ( \frac{D\sqrt{N}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{N})})
+Prob(D > observed) = \Phi ( \frac{D\sqrt{N}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{N})}) \, .
+\end{equation}
+The 2-Sample case uses the same formula as above, but with the slight variation where
+\begin{equation} \label{eq:5}
+N = \frac{n_1n_2}{n_1+n_2}\, .
+\end{equation}
+In both cases, $r$ is defined by
+\begin{equation} \label{eq:6}
+r = \frac{\sum_{i}^{}(x_i-\bar{x})(y_i-\bar{y})}{\sqrt{\sum_{i}^{}(x_i-\bar{x})^2}\sqrt{\sum_{i}^{}(y_i-\bar{y})^2}}\, .
 \end{equation}
 
-\textbf{The 2-Sample case}
-
-Uses the same formula as above, but with the slight variation where
-
-\begin{equation} \label{eq:5}
-N = \frac{n_1n_2}{n_1+n_2}
-\end{equation}
-
-In both cases, $r$ is defined by:
-
-\begin{equation} \label{eq:6}
-r = \frac{\sum_{i}^{}(x_i-\bar{x})(y_i-\bar{y})}{\sqrt{\sum_{i}^{}(x_i-\bar{x})^2}\sqrt{\sum_{i}^{}(y_i-\bar{y})^2}}
-\end{equation}
+\RBnote{In general, treat the equations as part of the sentences.  This means punctuating them -- see the examples above.}
+%use \verb|\,.| if the sentence ends with the equation, \verb|\,,| or \verb|\,;| if it's just the end of a clause.
+The \verb|\,| puts a little space before the punctuation to make it more readable; if you need more space, use \verb|\;| or \verb|$\quad$|.}
 
@@
 
-\subsection{Fasano Franceschini Test Usage}
+\subsection{Fasano--Franceschini Test Usage}
 
-In their paper, Fasano and Franceschini use Monte Carlo simulation to approximate the distribution of D as a function of $D$, the sample size $N$, and the coefficient of correlation $r$.
-Unlike the 1-D KS test, the 2-D KS test is not rigorously true that the distribution of $D_{FF}$ is independent of the shape of the 2-D distribution.
-Fasano and Franceschini reasoned that the coefficient of correlation between the two-dimensional distributions is what governs the resultant distribution of $D_{FF}$.
-For example, if the data sets are perfectly correlated ($r = 1$), the 2-D plane flattens to a single line and thus a 1-D KS test could be used.
-If the data sets are perfectly uncorrelated ($r = 0$), the 2-D distribution is independent in the x and y directions and thus one could use a 1-D KS test on the marginal x or y distribution of the 2-D plane.
-Results from Monte Carlo simulation support these assumptions -- showing that the distribution of D is nearly identical for varying distributions, assuming they have the same coefficient of correlation r ~\citep{Fasano1987}.
-The resultant distributional approximation fit by Teukolsky and Press -- as described above in equation \ref{eq:4} -- can be used to test the significance levels for the 2-D K-S test using the following code.
-
+In their paper, Fasano and Franceschini use Monte Carlo simulation to
+approximate the distribution of $D$ as a function of the sample
+size $N$ and the coefficient of correlation $r$. Unlike the 1-D KS
+test, the distribution of
+$D_{FF}$ is not independent of the shape of the 2-D distribution of the 
+underlying data. Fasano
+and Franceschini reasoned that the correlation between
+the two-dimensional distributions governs the resultant
+distribution of $D_{FF}$. For example, if the data sets are perfectly
+correlated ($r = 1$), the 2-D distribution lies along a single line and thus
+the 1-D KS test could be used. At the other extreme, if the data sets are perfectly
+uncorrelated ($r = 0$), the 2-D distribution is independent in the $x$
+and $y$ directions and thus one could use a 1-D KS test on the marginal
+distributions on the 2-D plane.
+%
+Results from Monte Carlo simulation support these assumptions, showing
+that the distribution of $D$ is nearly identical for varying
+distributions with the same correlation coefficient~\citep{Fasano1987}.
+The approximation 
+by Teukolsky and Press (Eq~\ref{eq:4}--\ref{eq:5})
+can be used to test the significance levels for the 2-D K-S test using
+the following code:
+%
 \begin{CodeChunk}
@@
 
- Fasano Franceschini Test (1987)
+ Fasano--Franceschini Test (1987)
  Data:  sample1Data and sample2Data
@@
 
-\subsection{Bootstrap Version of the Fasano Franceschini Test}
+\subsection{Bootstrap Version of the Fasano--Franceschini Test}
 
 As a caveat, to the fit produced by Teukolsky and Press are accurate when $N \gtrsim 20$, and the $p$-value is less than (more significant than) $\sim 0.20$.
-For $p$-value that are $> 0.20$, the distributional approximation may not be accurate.
-While the assertion the two data sets are \textit{not} significantly different is surely correct; analysis looking to compare $p$-value across experiments conditions -- such as if one was to do a cross study concordant analysis comparing rank order $p$-value \citep{Ness-Cohn2020}
-%(Cite bmal1 science Article or TimeTrial)
--- would not be possible with the Teukolsky and Press approximate fit.
-Moreover, additional arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
+For $p > 0.20$, the distributional approximation may not be accurate.
+While this inaccuracy still allows a simple rejection decision to be made, it is sometimes useful to accurately quantify $p$ (such as if one was to do a cross-ctudy concordance analysis comparing $p$ values between studies as in~\citep{Ness-Cohn2020}).
+%
+\RBnote{I'm not sure what this next sentence is getting at; elaborate or cut.} Moreover, additional arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
 
-To combat these accretions, one can bootstrap the significance levels for the particular multidimensional statistic directly from the particular data set under study.
+To address these limitations, one can bootstrap the significance levels for the particular multidimensional statistic directly from the particular data set under study.
 As Fasano and Franceschini's paper was originally released in 1987, this approach was unfeasible at scale.
-However, as we are currently in the age of high-performance computers -- a null distribution of $D_{FF}$ can be bootstrapped from the data to test significance.
+Today, modern computers can rapidly compute a bootstrapped null distribution of $D_{FF}$ from the data to test significance.
 
-The \pkg{fasano.franceschini.test} R package, implements a parallelized bootstrapping procedure.
-The marginal distribution from 2-dimensional data set is resampled with replacements to generate randomized 2-dimensional data sets \code{nBootStrap} times.
+The \pkg{fasano.franceschini.test} R package implements a parallelized bootstrapping procedure.
+The marginal distribution from 2-dimensional data set is resampled with replacement to generate randomized 2-dimensional data sets \code{nBootStrap} times.
 The frequency count by quadrant is performed for each bootstrapped resampling as described above to compute the $D_{FF}$.
 The observed $D_{FF}$ is then compared to the distribution of bootstrapped $D_{FF}$ to compute a $p$-value.
-The bootstrapped version of the Fasano Franceschini test can be implemented with the following lines of code -- see \fct{fasano.franceschini.test} for further source code details and implementation.
+The bootstrapped version of the Fasano--Franceschini test can be implemented as follows (see \fct{fasano.franceschini.test} for further source code details and implementation).
 
@@
 
-Note, bootstrapping may significantly increase run time performance.
-To improve run time, adjust the \code{cores} parameter to specify the number of parallel operations to run.
-See the R \pkg{parrellel} package and the \fct{mclapply} function for further details.
+Note, bootstrapping may significantly increase run time.
+To improve run time, one may adjust the \code{cores} parameter; see the R \pkg{parallel} package and the \fct{mclapply} function for further details.
 
 \subsection{Computational Efficency}
-The main distinction between the Peacock and Fasano Franceschini test comes down to compute time. For a benchmarking of the algorithms see \citep{Lopes2007}
-
-In Summary:
-\begin{itemize}
-\item The Fasano Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale.
-\item The Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
-\end{itemize}
-
-See the \pkg{Peacock.test} package for \proglang{R} implementation of the Peacock test.
+The main distinction between the Peacock and Fasano--Franceschini test comes down to compute time. For a benchmarking of the algorithms see \citep{Lopes2007}.  Briefly,
+%\begin{itemize}
+%\item The Fasano--Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale.
+%\item The Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
+%\end{itemize}
+the Fasano--Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale, while the Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
+%
+(See the \pkg{Peacock.test} package for \proglang{R} implementation of the Peacock test.)
 
@@
 The \pkg{fasano.franceschini.test} package is an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987}.
-Improvements:
-  (i) Fast algorithm
-  (ii) Prallelized Bootstrapping Procedure for Improved statistical testing
-
+It improves upon existing packages by implementing 
+  (i) a fast algorithm; and 
+  (ii) parallelized bootstrapping for improved statistical testing.
+\RBnote{CRAN?}
   See \url{https://nesscoder.github.io/fasano.franceschini.test/} for the complete package documentation and source code.
@@
 \url{https://CRAN.R-project.org/}.
+
+\RBnote{Any other package dependencies?}
 

### 2021-04-26 journalMS/article.tex
RB edits
paper revisions

@@
 \documentclass[codesnippet]{jss}
+%\documentclass[nojss]{jss}
 
@@
 \Abstract{
-\RBnote{Are citations allowed in the abstract (per journal style)?}
-The \pkg{fasano.franceschini.test} package is an  \proglang{R} implementation of the 2-D Kolmogorov-Smirnov (KS) two--sample test as defined by Fasano and Franceschini \citep{Fasano1987}.
-This is a variant of the 2-D two-sample KS test as originally defined by Peacock \citep{Peacock1983}, previously implemented in \proglang{R} in the \pkg{Peacock.test} package \citep{Xiao2016}.
-The \pkg{fasano.franceschini.test} package provides two improvements over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): (i) the Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$; and (ii) the package implements a  bootstrapping procedure for improved significance testing.
-\RBnote{Summarize what the paper will cover.}
+\RBnote{Are citations allowed in the abstract (per journal style)? - Yes, citation are found in the abstract both in the style guide template and also from checking published articles }
+The versatility of the univariate Kolmogorov-Smirnov (KS) test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines.
+Nonetheless, a generalization of the KS test as proposed by Kolmogorov and Smirnov does not naturally extend when working with data in more than one dimension.
+Here, we present the \pkg{fasano.franceschini.test} package, an  \proglang{R} implementation of the 2-D KS two--sample test as defined by Fasano and Franceschini~\citep{Fasano1987}.
+The \pkg{fasano.franceschini.test} package provides three improvements over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): (i) the Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$; (ii) the package implements a procedure for handling ties in the data; and (iii) the package implements a parallelized bootstrapping procedure for improved significance testing.
+Ultimately, the \pkg{fasano.franceschini.test} package presents a robust statistical test for analysing random samples defined in 2-dimensions.
+\RBnote{Summarize what the paper will cover. - ENC: Rewrote abstract}
 }
@@
   E-mail: \email{elan.ness-cohn@northwestern.edu}\\
-  URL: \url{https://sites.northwestern.edu/elannesscohn/}\\
+  URL: \url{https://www.nesscoder.com}\\
 
@@
 
-\RBnote{Please use nonbreaking spaces (tilde rather than space) before all cite commands}
+\RBnote{Please use nonbreaking spaces (tilde rather than space) before all cite commands - ENC: tilde has be added before each cite, removing all spaces}
 
@@
 
-The versatility of the univariate KS test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines \RBnote{highlighting your note: (Cite uses in a broad range of field)}.
+The versatility of the univariate KS test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines~\citep{Atasoy2017,Chiang2018,Hahne2018,Hargreaves2020,Wong2020,Kaczanowska2021} \RBnote{ENC: Cited a broad range of fields (neuropharmacology, climate change, covid-19, sustainable nutrition, robotic prostetics, cancer therapeutics)}.
 Nonetheless, a generalization of the KS test as proposed by Kolmogorov and Smirnov does not naturally extend when working with data in more than one dimension.
@@
 
-Currently, only the Peacock implementation of the 2-D two-sample KS test is available in \proglang{R}~\citep{R} with the \pkg{Peacock.test} package via the \fct{peacock2} function, but this has been shown to be markedly slower than the Fasano and Franceschini algorithm\citep{Lopes2007}.
-A \proglang{C} implementation of the Fasano--Franceschini test is available in \citep{numericalRecipes}; however, arguments have been made to the validity of the implementation of the test not being distribution-free\citep{Babu2006}.
+Currently, only the Peacock implementation of the 2-D two-sample KS test is available in \proglang{R}~\citep{R} with the \pkg{Peacock.test} package via the \fct{peacock2} function, but this has been shown to be markedly slower than the Fasano and Franceschini algorithm~\citep{Lopes2007}.
+A \proglang{C} implementation of the Fasano--Franceschini test is available in~\citep{numericalRecipes}; however, arguments have been made to the validity of the implementation of the test not being distribution-free~\citep{Babu2006}.
 Furthermore, the \proglang{C} implementation statistical testing is based on a fit to Monte Carlo Simulation and is only valid for significance levels $\alpha \lessapprox 0.20$, which may present potential issues in some analysis.
 
-Here we present the \pkg{fasano.franceschini.test} package as an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987} The \pkg{fasano.franceschini.test} package provides two improvements over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): (i) the Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$; and (ii) the package implements a bootstrapping procedure for improved significance testing and mitigates the concerns over the distribution-free nature of the test brought up by~\citep{Babu2006}.
+Here we present the \pkg{fasano.franceschini.test} package as an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini~\citep{Fasano1987} The \pkg{fasano.franceschini.test} package provides two improvements over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): (i) the Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$; and (ii) the package implements a bootstrapping procedure for improved significance testing and mitigates the concerns over the distribution-free nature of the test brought up by~\citep{Babu2006}.
 
@@
 %\textbf{For the 1-Sample case}, the $p$-value probability of a given D is defined by
-In the one-sample case, the $p$-value is given by 
+In the one-sample case, the $p$~value is given by
 \begin{equation} \label{eq:2}
@@
 \end{equation}
-in the two-sample case, the $p$-value is given by
+in the two-sample case, the $p$~value is given by
 %where $N$ is the number of data points sampled. A $p$-value $< 0.05$, indicates a significant difference between the samples.
@@
 \begin{equation} \label{eq:3}
-\mathbb{P}(D > observed) = \Phi ( D\sqrt{\frac{n_1n_2}{n_1+n_2}})\,.
+\mathbb{P}(D > observed) = \Phi \left( D\sqrt{\frac{n_1n_2}{n_1+n_2}} \right)\,.
 \end{equation}
@@
 %where $n_1$ and $n_2$ is the number of data points in the first and second samples respectively. A $p$-value $< 0.05$, indicates a significant difference between the samples.
+\RBnote{Note usage of left( and right) [alternatively, or amsmath big( ] to make bigger parens}
 
@@
 \includegraphics{fftestOutput}
-\caption{\label{fig:kstest2D} Illustration of the Fasano--Franceschini algorithmic search for the maximal difference ($D_{FF}$) between sample 2-D eCDFs.
+\caption{\label{fig:kstest2D} Illustration of the Fasano--Franceschini algorithmic search for the maximal difference ($D_{FF,1}$) between sample 2-D eCDFs.
 Looping through each point in the sampled data to define a unique origin (grey dotted line), the fraction of orange and blue points in each quadrants are computed (plot corners).
@@
 The fraction of orange and blue points in each quadrant is computed (corners of plot) and the quadrant with the maximal difference is designated with the current maximum for the specified origin.
-By iterating over all pairs of data points and quadrants, the $D_{FF}$ is defined by the maximal difference of the integrated probabilities between samples.
-In this case using the orange point as the origin, the maximal difference is $D_{FF} = 0.52$.
-\RBnote{How are tied values counted? Describe that here.  Also discuss the treatment of the origin.}
+By iterating over all pairs of data points and quadrants, the $D_{FF,1}$ is defined by the maximal difference of the integrated probabilities between samples.
+In this case using the orange point as the origin, the maximal difference is $D_{FF,1} = 0.52$.
+\RBnote{How are tied values counted? Describe that here.  Also discuss the treatment of the origin. - ENC: Discussion Added below}
 
-\RBnote{Since $D_{FF}$ from one sample isn't the final $D_{FF}$, you should note it appropriately (eg, $D_{FF,1}$ and $D_{FF,2}$).}
-This process is repeated using the blue points as the origins to compute the maximal $D_{FF}$.
-Finally, both maximal $D_{FF}$ statistics using the orange and blue points as origins are then averaged to compute the overall $D_{FF}$ for hypothesis testing.
+\RBnote{Since $D_{FF}$ from one sample isn't the final $D_{FF}$, you should note it appropriately (eg, $D_{FF,1}$ and $D_{FF,2}$). - ENC: Text has been updated accordingly}
+This process is repeated using the blue points as the origins to compute the maximal $D_{FF,2}$.
+Finally, both maximal $D_{FF,1}$ and $D_{FF,2}$ statistics using the orange and blue points as origins are averaged to compute the overall $D_{FF}$ for hypothesis testing.
+
+Note that in real data, it is possible for some points to be tied with the X and/or Y coordinates of the origin -- creating an ambuguity when computing the fraction of points in each quadrant.
+Since the Fransono Francheschini test attempts to define the maximal difference of the integrated probabilities, a natural solution would be to treat a point that is tied with the current X and/or Y coordinates of the origin as equally likely to have been drawn from any of the tied quadrants.
+In doing so, the integrated probability of the given point is exactly evenly distributed across the tied quadrants.
+For example, any data point sharing the same X or Y coordianted as the current origin is evenly distributed across the tied quadrants, with each of the two quadrants recieveing half a count.
+Any data point sharing the same X and Y coordiantes as the current origin  -- the origin included -- is evenly distributed across all quadrants, with all four quadrants recieveing a quarter count.
 
@@
 \begin{equation} \label{eq:4}
-Prob(D > observed) = \Phi ( \frac{D\sqrt{N}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{N})}) \, .
+\mathbb{P}(D > observed) = \Phi \left( \frac{D\sqrt{N}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{N})} \right) \, .
 \end{equation}
@@
  D-stat =  0.14 , p-value =  0.4420642
- Run Time (s) =  0.04288507
+ Run Time (s) =  0.06367922
 \end{CodeOutput}
@@
 For $p > 0.20$, the distributional approximation may not be accurate.
-While this inaccuracy still allows a simple rejection decision to be made, it is sometimes useful to accurately quantify $p$ (such as if one was to do a cross-ctudy concordance analysis comparing $p$ values between studies as in~\citep{Ness-Cohn2020}).
+While this inaccuracy still allows a simple rejection decision to be made, it is sometimes useful to accurately quantify $p$ (such as if one was to do a cross-ctudy concordance analysis comparing $p$~values between studies as in~\citep{Ness-Cohn2020}).
 %
-\RBnote{I'm not sure what this next sentence is getting at; elaborate or cut.} Moreover, additional arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
+\RBnote{I'm not sure what this next sentence is getting at; elaborate or cut. - ENC: removed the next line}
+%Moreover, additional arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
 
@@
 The frequency count by quadrant is performed for each bootstrapped resampling as described above to compute the $D_{FF}$.
-The observed $D_{FF}$ is then compared to the distribution of bootstrapped $D_{FF}$ to compute a $p$-value.
+The observed $D_{FF}$ is then compared to the distribution of bootstrapped $D_{FF}$ to compute a $p$~value.
 The bootstrapped version of the Fasano--Franceschini test can be implemented as follows (see \fct{fasano.franceschini.test} for further source code details and implementation).
@@
 R> fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data,
-R>                          nBootstrap = 1000, cores = 6)
+R>                          nBootstrap = 1000, cores = 1)
 \end{CodeInput}
@@
  Data:  sample1Data and sample2Data
- D-stat =  0.14 , p-value =  0.672
- Run Time (s) =  2.694002
+ D-stat =  0.14 , p-value =  0.62
+ Run Time (s) =  12.59964
 \end{CodeOutput}
@@
 \subsection{Computational Efficency}
-The main distinction between the Peacock and Fasano--Franceschini test comes down to compute time. For a benchmarking of the algorithms see \citep{Lopes2007}.  Briefly,
+The main distinction between the Peacock and Fasano--Franceschini test comes down to compute time. For a benchmarking of the algorithms see~\citep{Lopes2007}.  Briefly,
 %\begin{itemize}
@@
 
-The \pkg{fasano.franceschini.test} package is an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini \citep{Fasano1987}.
+The \pkg{fasano.franceschini.test} package is an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini~\citep{Fasano1987}.
 It improves upon existing packages by implementing
-  (i) a fast algorithm; and 
-  (ii) parallelized bootstrapping for improved statistical testing.
-\RBnote{CRAN?}
-  See \url{https://nesscoder.github.io/fasano.franceschini.test/} for the complete package documentation and source code.
+  (i) a fast algorithm;
+  (ii) a extension to the Fasano and Franceschini test for dealing with ties; and
+  (iii) parallelized bootstrapping for improved statistical testing.
+\RBnote{CRAN? - ENC: Ready to Submit to CRAN}
+  See the complete package documentation and source code via the Comprehensive
+\proglang{R} Archive Network (CRAN) at
+\url{https://CRAN.R-project.org/} or the package website at \url{https://nesscoder.github.io/fasano.franceschini.test/}.
 
@@
 The results in this paper were obtained using
-\proglang{R}~3.6.3 with the
+\proglang{R}~4.0.3 with the
 \pkg{fasano.franceschini.test}~1.0.0 package. \proglang{R} itself
-and all packages used are available from the Comprehensive
+and all package dependencies (\pkg{methods}~4.0.3; \pkg{parallel}~4.0.3) are available from the Comprehensive
 \proglang{R} Archive Network (CRAN) at
@@
 
-\RBnote{Any other package dependencies?}
+\RBnote{Any other package dependencies? - ENC: added dependenceis parallel and methods}
 

### 2021-04-26 journalMS/article.R

@@ -1,17 +1,29 @@
-## Package and options
+#' ## Figure Reproducibility Script
+#' ### Fasano--Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in R
+#' ### Authors: Elan Ness-Cohn and Rosemary Braun
+
+
+#+ setup, include=FALSE
+library(knitr)
+options(knitr.duplicate.label = "allow")
+opts_chunk$set(fig.width = 14, fig.height = 7)
+
+#' ## Package and options
 library(devtools)
 load_all()
 library("fasano.franceschini.test")
 library("parallel")
 library("ggplot2")
-library(viridis)
-library(mvtnorm)
-library(gridExtra)
-library(ggpubr)
+library("viridis")
+library("mvtnorm")
+library("gridExtra")
+library("ggpubr")
 
-#-------------------------------------#
-#    fasano.franceschini.test Usage
-#-------------------------------------#
+# set to T to plot figures to PDF Output
+pdfOutPuts <- F
+
+#' ## fasano.franceschini.test Usage
+
 # set seed for reproducible example
 set.seed(123)
 
@@ -27,9 +39,9 @@ sample2Data <- data.frame(
 
 fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data)
 
-#-------------------------------------#
-# fasano.franceschini.test Bootstrap
-#-------------------------------------#
+
+#' ## fasano.franceschini.test Bootstrap
+
 # set seed for reproducible example
 set.seed(123)
 
@@ -43,13 +55,16 @@ sample2Data <- data.frame(
   y = rnorm(n = 100, mean = 0, sd = 1)
 )
 
-fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data, nBootstrap = 1000, cores = 6)
+fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data, nBootstrap = 1000, cores = 1)
 
-#-------------------------------------#
-#  Visualization - 2-D KS test
-#-------------------------------------#
-png(file = paste0("~/Desktop/fftestOutput.png"),units = "in",res = 300,width = 14,height = 7)
[…]

### 2021-04-26 R/fasano_franceschini_test.R
updated package for dependenies

@@ -4,8 +4,8 @@
 #'
 #' Code adapted from Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688
 #'
-#' @param S1 a [n by 2] `data.frame` of x and y coordinates of sample 1
-#' @param S2 a [n by 2] `data.frame` of x and y coordinates of sample 2
+#' @param S1 a `[n by 2]` `data.frame` of x and y coordinates of sample 1
+#' @param S2 a `[n by 2]` `data.frame` of x and y coordinates of sample 2
 #' @param nBootstrap a `numeric` defining the number of bootstrapped samples to be generated for computing the empirical p-value (note this procedure is slow and computationally expensive on the order of nBootStrap*O(n^2). Default is set to 0.
 #' If nBootstrap is 0, the Fasano Franceschini distributional approximation is used for defining the p-value. See Fasano and Franceschini test (1987) for details.
 #' @param cores a `numeric` defining the number of cores to use for processing
@@ -42,9 +42,25 @@
 #'
 #' fasano.franceschini.test(sample1Data,sample2Data)
 #'
+#' @import methods
 #' @export
 
 fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
+
+  #validate inputs
+  if(!is.data.frame(S1) | !is.data.frame(S2)){
+    stop("S1 and S2 must be a `data.frame`")
+  }
+  if(ncol(S1) != 2 | ncol(S2) != 2){
+    stop("S1 and S2 must be a `data.frame` of dim nrows x 2 cols, more than 2 cols detected")
+  }
+  if(nBootstrap < 0){
+    stop("nBootstrap must be a positive value")
+  }
+
+  #cores can't be more than on computer
+
+
   start <- Sys.time()
   # determine number of samples in each data set
   n1 <- dim(S1)[1]
@@ -59,7 +75,7 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
   # average KS stat
   D <- (d1 + d2) / 2
 
-  # if bootstramp is enabled, compute the bootstrapped null
+  # if bootstrap is enabled, compute the bootstrapped null d
   if (nBootstrap > 0) {
     x_marg <- c(S1[, 1], S2[, 1])
[…]

### 2021-04-26 R/interal_functions_FasFran.R

@@ -58,73 +58,6 @@ quadCount <- function(x_origin, y_origin, x, y) {
 
   return(c(freqQuad1, freqQuad2, freqQuad3, freqQuad4))
 }
-# quadCount2 <- function(x_origin, y_origin, x, y) {
-#   n <- length(x)
-#   nq1 <- 0
-#   nq2 <- 0
-#   nq3 <- 0
-#   nq4 <- 0
-#   org <- 0
-#
-#   # think about less than or equal to?
-#   # i.e. ties
-#   # for (k in 1:n) {
-#   #     if (y[k] >= y_origin) {
-#   #       if (x[k] >= x_origin) {
-#   #         nq1 <- nq1 + 1
-#   #       } else {
-#   #         nq2 <- nq2 + 1
-#   #       }
-#   #     } else {
-#   #       if (x[k] >= x_origin) {
-#   #         nq4 <- nq4 + 1
-#   #       } else {
-#   #         nq3 <- nq3 + 1
-#   #       }
-#   #     }
-#   # }
-#   for (k in 1:n) {
-#     if (y[k] > y_origin) {
-#       if (x[k] > x_origin) {
-#         nq1 <- nq1 + 1
-#       } else if (x[k] < x_origin) {
-#         nq2 <- nq2 + 1
-#       } else {
-#         nq1 <- nq1 + 0.5
-#         nq2 <- nq2 + 0.5
-#       }
-#     } else if (y[k] < y_origin) {
-#       if (x[k] > x_origin) {
-#         nq4 <- nq4 + 1
-#       } else if (x[k] < x_origin) {
-#         nq3 <- nq3 + 1
-#       } else {
-#         nq3 <- nq3 + 0.5
-#         nq4 <- nq4 + 0.5
-#       }
-#     } else {
-#       if (y[k] == y_origin) {
-#         if (x[k] > x_origin) {
-#           nq1 <- nq1 + 0.5
-#           nq4 <- nq4 + 0.5
-#         } else if (x[k] < x_origin) {
-#           nq2 <- nq2 + 0.5
-#           nq3 <- nq3 + 0.5
-#         } else {
-#           org <- org + 1
-#         }
-#       }
-#     }
-#   }
-#
-#   n <- n - org
-#   freqQuad1 <- nq1 / n
-#   freqQuad2 <- nq2 / n
-#   freqQuad3 <- nq3 / n
-#   freqQuad4 <- nq4 / n
[…]

### 2021-06-08 journalMS/article.tex
RB edits finally

@@
 \Abstract{
-\RBnote{Are citations allowed in the abstract (per journal style)? - Yes, citation are found in the abstract both in the style guide template and also from checking published articles }
-The versatility of the univariate Kolmogorov-Smirnov (KS) test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines.
-Nonetheless, a generalization of the KS test as proposed by Kolmogorov and Smirnov does not naturally extend when working with data in more than one dimension.
-Here, we present the \pkg{fasano.franceschini.test} package, an  \proglang{R} implementation of the 2-D KS two--sample test as defined by Fasano and Franceschini~\citep{Fasano1987}.
-The \pkg{fasano.franceschini.test} package provides three improvements over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): (i) the Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$; (ii) the package implements a procedure for handling ties in the data; and (iii) the package implements a parallelized bootstrapping procedure for improved significance testing.
-Ultimately, the \pkg{fasano.franceschini.test} package presents a robust statistical test for analysing random samples defined in 2-dimensions.
-\RBnote{Summarize what the paper will cover. - ENC: Rewrote abstract}
+\RBnote{Perhaps briefly state what the KS test is.}
+The versatility of the univariate Kolmogorov-Smirnov (KS) test has
+made it a cornerstone of statistical analysis and is commonly used
+across the scientific disciplines. However, the test proposed by
+Kolmogorov and Smirnov does not naturally extend to multidimensional
+distributions.  Here, we present the \pkg{fasano.franceschini.test}
+package, an  \proglang{R} implementation of the 2-D KS two--sample
+test as defined by Fasano and Franceschini~\citep{Fasano1987}. The
+\pkg{fasano.franceschini.test} package provides three improvements
+over the current 2-D KS test on the Comprehensive \proglang{R} Archive
+Network (CRAN): (i) the Fasano and Franceschini test has been shown to
+run in $O(n^2)$ versus the Peacock implementation which runs in
+$O(n^3)$; (ii) the package implements a procedure for handling ties in
+the data; and (iii) the package implements a parallelized
+bootstrapping procedure for improved significance testing. Ultimately,
+the \pkg{fasano.franceschini.test} package presents a robust
+statistical test for analysing random samples defined in 2-dimensions.
 }
@@
 % C <- Implementation
+The Kolmogorov--Smirnov (KS) is a non--parametric, univariate
+statistical test designed to assess whether a set of data is
+consistent with a given probability distribution (or, in the two-sample 
+case,  whether the two 
+samples come from the same underlying distribution). First derived by
+Kolmogorov and Smirnov in a series of
+papers~\citep{Kolmogorov1933,Kolmogorov1933a,Smirnov1936,Smirnov1937,Smirnov1939,Smirnov1944,Smirnov1948},
+the one-sample KS test defines the distribution of the quantity
+$D_{KS}$, the maximal absolute difference between the empirical
+cumulative distribution function (CDF) of a set of values and a
+reference probability distribution. Kolmogorov and Smirnov's key
+insight was proving the distribution of $D_{KS}$ was independent of
+the CDFs being tested. Thus, the test can effectively be used to
+compare any univariate empirical data distribution to any continuous
+univariate reference distribution. The two-sample KS test could further
+be used to compare any two univariate empirical data distributions
+against each other to determine if they are drawn from the same
+underlying univariate distribution.
 
-\RBnote{Please use nonbreaking spaces (tilde rather than space) before all cite commands - ENC: tilde has be added before each cite, removing all spaces}
+The nonparametric versatility of the univariate KS test has made it a cornerstone of
+statistical analysis and is commonly used across the scientific
+disciplines~\citep{Atasoy2017,Chiang2018,Hahne2018,Hargreaves2020,Wong2020,Kaczanowska2021}.
+However, the KS test as proposed by Kolmogorov
+and Smirnov does not naturally extend to distributions in more
+than one dimension. Fortunately, a solution to the dimensionality
+issue was articulated by Peacock~\citep{Peacock1983} and later
+extended by Fasano and Franceschini~\citep{Fasano1987}.
 
-The Kolmogorov--Smirnov (KS) is a non--parametric, univariate statistical test designed to assess whether a set of data is consistent with a given probability distribution.
-First derived by Kolmogorov and Smirnov in a series of papers~\citep{Kolmogorov1933,Kolmogorov1933a,Smirnov1936,Smirnov1937,Smirnov1939,Smirnov1944,Smirnov1948}, the one-sample KS test defines the distribution of the quantity $D_{KS}$, the maximal absolute difference between the empirical cumulative distribution function (CDF) of a set of values and a reference probability distribution.
-Kolmogorov and Smirnov's key insight was proving the distribution of $D_{KS}$ was independent of the CDFs being tested.
-Thus, the test can effectively be used to compare any univariate empirical data distribution to any continuous univariate reference distribution.
-A two-sample KS test could further be used to compare any two univariate empirical data distributions against each other to determine if they are drawn from the same underlying univariate distribution.
+Currently, only the Peacock implementation of the 2-D two-sample KS
+test is available in \proglang{R}~\citep{R} with the
+\pkg{Peacock.test} package via the \fct{peacock2} function, but this
+has been shown to be markedly slower than the Fasano and Franceschini
+algorithm~\citep{Lopes2007}. A \proglang{C} implementation of the
+Fasano--Franceschini test is available in~\cite{numericalRecipes};
+however, arguments have been made to the validity of the
+implementation of the test not being
+distribution-free~\citep{Babu2006}. Furthermore, the \proglang{C}
+implementation statistical testing is based on a fit to Monte Carlo
+Simulation and is only valid for significance levels $\alpha \lessapprox 0.20$, 
+which may present potential issues in some analysis.
 
-The versatility of the univariate KS test has made it a cornerstone of statistical analysis and is commonly used across the scientific disciplines~\citep{Atasoy2017,Chiang2018,Hahne2018,Hargreaves2020,Wong2020,Kaczanowska2021} \RBnote{ENC: Cited a broad range of fields (neuropharmacology, climate change, covid-19, sustainable nutrition, robotic prostetics, cancer therapeutics)}.
-Nonetheless, a generalization of the KS test as proposed by Kolmogorov and Smirnov does not naturally extend when working with data in more than one dimension.
-Fortunately, a solution to the dimensionality issue was articulated by Peacock~\citep{Peacock1983} and later extended by Fasano and Franceschini~\citep{Fasano1987}.
-
-Currently, only the Peacock implementation of the 2-D two-sample KS test is available in \proglang{R}~\citep{R} with the \pkg{Peacock.test} package via the \fct{peacock2} function, but this has been shown to be markedly slower than the Fasano and Franceschini algorithm~\citep{Lopes2007}.
-A \proglang{C} implementation of the Fasano--Franceschini test is available in~\citep{numericalRecipes}; however, arguments have been made to the validity of the implementation of the test not being distribution-free~\citep{Babu2006}.
-Furthermore, the \proglang{C} implementation statistical testing is based on a fit to Monte Carlo Simulation and is only valid for significance levels $\alpha \lessapprox 0.20$, which may present potential issues in some analysis.
-
-Here we present the \pkg{fasano.franceschini.test} package as an \proglang{R} implementation of the 2-D two-sample KS test as defined by Fasano and Franceschini~\citep{Fasano1987} The \pkg{fasano.franceschini.test} package provides two improvements over the current 2-D KS test on the Comprehensive \proglang{R} Archive Network (CRAN): (i) the Fasano and Franceschini test has been shown to run in $O(n^2)$ versus the Peacock implementation which runs in $O(n^3)$; and (ii) the package implements a bootstrapping procedure for improved significance testing and mitigates the concerns over the distribution-free nature of the test brought up by~\citep{Babu2006}.
-
-% \begin{leftbar}
-% The introduction is in principle ``as usual''. However, it should usually embed
-% both the implemented \emph{methods} and the \emph{software} into the respective
-% relevant literature. For the latter both competing and complementary software
-% should be discussed (within the same software environment and beyond), bringing
-% out relative (dis)advantages. All software mentioned should be properly
-% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
-% \textsc{Bib}{\TeX}.)
-%
-% For writing about software JSS requires authors to use the markup
-% \verb|\proglang{}| (programming languages and large programmable systems),
-% \verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
-% arguments, etc.). If there is such markup in (sub)section titles (as above), a
-% plain text version has to be provided in the {\LaTeX} command as well. Below we
-% also illustrate how abbrevations should be introduced and citation commands can
-% be employed. See the {\LaTeX} code for more details.
-% \end{leftbar}
-%
-% Modeling count variables is a common task in economics and the social sciences.
-% The classical Poisson regression model for count data is often of limited use in
-% these disciplines because empirical count data sets typically exhibit
-% overdispersion and/or an excess number of zeros. The former issue can be
-% addressed by extending  the plain Poisson regression model in various
-% directions: e.g., using sandwich covariances or estimating an additional
-% dispersion parameter (in a so-called quasi-Poisson model). Another more formal
-% way is to use a negative binomial (NB) regression. All of these models belong to
-% the family of generalized linear models (GLMs). However, although these models
-% typically can capture overdispersion rather well, they are in many applications
-% not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
-% increased interest in zero-augmented models that address this issue by a second
-% model component capturing zero counts. An overview of count data models in
-% econometrics, including  hurdle and zero-inflated models, is provided in
-% \cite{Cameron+Trivedi:2013}.
-%
-% In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
-% \fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
-% \citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
-% diagnostics and inference. The manuscript that this document is based on
-% \citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
-% count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
-% package \citep{Jackman:2015}. Of course, much more software could be discussed
-% here, including (but not limited to) generalized additive models for count data
-% as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006},
-% \pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.
-
-%% -- Manuscript ---------------------------------------------------------------
-
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
+Here we present the \pkg{fasano.franceschini.test} package as an
+\proglang{R} implementation of the 2-D two-sample KS test described
+by Fasano and Franceschini~\citep{Fasano1987}. The
+\pkg{fasano.franceschini.test} package provides two improvements over
+the current 2-D KS test on the Comprehensive \proglang{R} Archive
+Network (CRAN): (i) the Fasano and Franceschini test has been shown to
+run in $O(n^2)$ versus the Peacock implementation which runs in
+$O(n^3)$; and (ii) the package implements a bootstrapping procedure
+for improved significance testing and mitigates the limitations of
+the test brought noted by~\citep{Babu2006}.
 
@@
 
-\RBnote{This can be shortened substantially.  I have done so, but please look closely.  In general, avoid writing textbook material in papers.}
-
-The Kolmogorov--Smirnov (KS) test is a non--parametric method for determining whether a sample is consistent with a given probability distribution.
+The Kolmogorov--Smirnov (KS) test is a non--parametric method for
+determining whether a sample is consistent with a given probability
+distribution~\citep{Stephens1992a}.
 %While a comprehensive discussion pertaining the intricacies of the 1-D KS test can be found elsewhere~\citep{Stephens1992a}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
@@
 %
-In one dimension,
-the Kolmogorov-Smirnov statistic ($D_{KS}$) is the defined by the maximum absolute difference between the cumulative density functions of the data and model (one--sample), or between the two data sets (two--sample),
-as illustrated in (\textbf{Figure~\ref{fig:kstest1D}}).
+In one dimension, the Kolmogorov-Smirnov statistic ($D_{KS}$) is the
+defined by the maximum absolute difference between the cumulative
+density functions of the data and model (one--sample), or between the
+two data sets (two--sample), as illustrated in
+\textbf{Figure~\ref{fig:kstest1D}}.
 %.
@@
 
-\begin{figure}[t!]
+\begin{figure}[hbt]
 \centering
 \includegraphics{pdfvsCDF}
-\caption{\label{fig:kstest1D} \textbf{LEFT:} Probability density function (PDF) of two normal distribution: orange sample1 $\mathcal{N}(\mu = 0,\,\sigma^{2} = 1)$; blue sample2 $\mathcal{N}(\mu = 5,\,\sigma^{2} = 1)$.
-\textbf{RIGHT:} Empirical cumulative density functions (eCDF) of the two PDFs in the left panel.
-Black dotted line represents the maximal absolute difference between the eCDFs ($D_{KS}$).
+\caption{\label{fig:kstest1D} \textbf{LEFT:} Probability density function (PDF) of two normal distributions: orange sample~1, $\mathcal{N}(\mu = 0,\,\sigma^{2} = 1)$; blue sample~2, $\mathcal{N}(\mu = 5,\,\sigma^{2} = 1)$.
+\textbf{RIGHT:} Cumulative density functions (CDF) of the two PDFs; the black dotted line represents the maximal absolute difference between the CDFs ($D_{KS}$).
 }
@@
 %\textbf{For the 1-Sample case}, the $p$-value probability of a given D is defined by
-In the one-sample case, the $p$~value is given by
+In the one-sample case with a sample of size $n$, the $p$~value is given by
 \begin{equation} \label{eq:2}
@@
 %
-%where $n_1$ and $n_2$ is the number of data points in the first and second samples respectively. A $p$-value $< 0.05$, indicates a significant difference between the samples.
-\RBnote{Note usage of left( and right) [alternatively, or amsmath big( ] to make bigger parens}
+where $n_1$ and $n_2$ are the number of observations in the first and second samples respectively. 
+%A $p$-value $< 0.05$, indicates a significant difference between the samples.
 
-\subsection{Higher Dimensional KS Test}
-
+\subsection{Higher dimensional variations: Peacock Test (1983) and Fasano--Franceschini Test (1987)}
 %Similar to the 1-D KS test, the obvious solution to extend the theory into higher dimensional would be to define the maximal cumulative difference between the two 2-D distributions.
@@
 In N dimensional space there are $2^{N}-1$ independent ways of defining the cumulative distribution function~\citep{Peacock1983}.
-Extending the above to two or higher dimension is complicated by the fact that CDFs are not well-defined in more than one dimension.
-In 2-D, there are 4 ways (3 independent ways) of defining the cumulative distribution, since the direction in which we order the $x$ and $y$ points are arbitrary (\textbf{Figure \ref{fig:kstest2Dissue}}); more generally, in $P$-dimensional space there are $2^{P}-1$ independent ways of defining the cumulative distribution function~\citep{Peacock1983}.
+Extending the above to two or higher dimension is complicated by the
+fact that CDFs are not well-defined in more than one dimension.  In
+2-D, there are 4 ways (3 independent ways) of defining the cumulative
+distribution, since the direction in which we order the $x$ and $y$
+points are arbitrary (\textbf{Figure \ref{fig:kstest2Dissue}}); more
+generally, in $k$-dimensional space there are $2^{k}-1$ independent
+ways of defining the cumulative distribution
+function~\citep{Peacock1983}.
 
@@
 \caption{\label{fig:kstest2Dissue} Four ways (3 independent) of defining the cumulative distribution for a given point in 2-D.
+Here, the orange point $(X,Y)$ is chosen as the origin; the density of observations may be integrated as
+$\mathbb{P}(x < X, y > Y)$ (A); 
+$\mathbb{P}(x < X \cup y < Y)$ (B); 
+$\mathbb{P}(x < X, y < Y)$ (C); 
+$\mathbb{P}(x > X, y > Y)$ (D). 
 %$P(x < X, y < Y) + P(x < X, y > Y) + P(x > X, y < Y) = 1 - P(x > X, y > Y)$
@@
 
-\subsection{Peacock Test (1983) and Fasano--Franceschini Test (1987)}
-Peacock solved the higher dimensionality issue by defining the 2-D test statistic as the largest difference between the empirical and theoretical cumulative distributions, after taking all possible ordering combinations into account.
-Peacock's test thus computes the total integrated probability -- i.e.\,fraction of data -- in each of the four quadrants around all possible points in the data.
-For example, for a given $n$ points in a two-dimensional space, the cumulative distribution functions is calculated in the $4n^2$ quadrants of the plane defined by all pairs ($X_i$ ,$Y_j$), $X_i$ and $Y_j$ being coordinates of any pairs of points in the given sample.
-By ranging over all possible pairs of data points and quadrants, the 2-dimensional $D$ statistic is defined by the maximal difference of the integrated probabilities between samples.
+\cite{Peacock1983} solved the higher dimensionality issue by defining
+the 2-D test statistic as the largest difference between the empirical
+and theoretical cumulative distributions, after taking all possible
+ordering combinations into account. Peacock's test thus computes the
+total integrated probability---i.e.\,fraction of data---tin each of
+the four quadrants around all possible points in the data. For
+example, for a given $n$ points in a two-dimensional space, the
+cumulative distribution functions is calculated in the $4n^2$
+quadrants of the plane defined by all pairs ($X_i$ ,$Y_j$), $X_i$ and
+$Y_j$ being coordinates of any pairs of points in the given sample. By
+ranging over all possible pairs of data points and quadrants, the
+2-dimensional $D$ statistic is defined by the maximal difference of
+the integrated probabilities between samples.
 
-The variation defined by Fasano and Franceschini was to only consider quadrants centered in each point of the given samples to compute the cumulative distribution functions.
-Rather than looking over all $X_i, Y_j: i,j={1,n}$, Fasano and Franceschini only used the detected points $X_i, Y_i : i={1,n}$.
-Thus for any given $n$ points in a two-dimensional space, those $n$ points define $4n$ -- rather than $4n^2$ -- quadrants.
-
-The 2-D two sample Fasano and Franceschini test is illustrated in the following example to compare the 2-D distributions of orange versus blue points (\textbf{Figure \ref{fig:kstest2D}}).
+The variation defined by \cite{Fasano1987}  was to only consider
+quadrants centered in each point of the given samples to compute the
+cumulative distribution functions.  Rather than looking over all 
+$X_i, Y_j: i,j={1,n}$, Fasano and Franceschini only used the detected points
+$X_i, Y_i : i={1,n}$. 
+\RBnote{I don't follow the preceding line? (Notation?)}
+Thus for any given $n$ points in a two-dimensional space, those $n$ points define $4n$ (rather than
+$4n^2$) quadrants.  The proceedure is illustrated in \textbf{Figure~\ref{fig:kstest2D}}.
+The algorithm loops through each point in one sample in turn to define
+the origin of 4 quadrants (grey dotted lines in
+\textbf{Figure~\ref{fig:kstest2D}}). The fraction of points in each
+sample is computed in each quadrant, and the quadrant with the maximal
+difference is designated with the current maximum for the specified
+origin. By iterating over all data points and quadrants, the test
+statistic $D_{FF,1}$ is defined by the maximal difference of the
+integrated probabilities between samples in any quadrant for any
+origin from the first sample.  In \textbf{Figure~\ref{fig:kstest2D}},
+using the orange point as the origin, the maximal difference is
+$D_{FF,1} = 0.52$.
 
@@
 
-The algorithm loops through each point in the orange sample in turn to define the origin of 4 quadrants (grey dotted line).
-The fraction of orange and blue points in each quadrant is computed (corners of plot) and the quadrant with the maximal difference is designated with the current maximum for the specified origin.
-By iterating over all pairs of data points and quadrants, the $D_{FF,1}$ is defined by the maximal difference of the integrated probabilities between samples.
-In this case using the orange point as the origin, the maximal difference is $D_{FF,1} = 0.52$.
-\RBnote{How are tied values counted? Describe that here.  Also discuss the treatment of the origin. - ENC: Discussion Added below}
+%\RBnote{Since $D_{FF}$ from one sample isn't the final $D_{FF}$, you should note it appropriately (eg, $D_{FF,1}$ and $D_{FF,2}$). - ENC: Text has been updated accordingly}
+This process is repeated using the points from \textit{other} sample
+as the origins to compute the maximal $D_{FF,2}$ with origins from the
+second sample. $D_{FF,1}$ and $D_{FF,2}$  are then averaged to compute
+the overall $D_{FF}$ for hypothesis testing. \RBnote{clarify: is
+$D_{FF}=(D_{FF,1}+D_{FF,2})/2$, or is it weighted by the $n$?}
 
-\RBnote{Since $D_{FF}$ from one sample isn't the final $D_{FF}$, you should note it appropriately (eg, $D_{FF,1}$ and $D_{FF,2}$). - ENC: Text has been updated accordingly}
-This process is repeated using the blue points as the origins to compute the maximal $D_{FF,2}$.
-Finally, both maximal $D_{FF,1}$ and $D_{FF,2}$ statistics using the orange and blue points as origins are averaged to compute the overall $D_{FF}$ for hypothesis testing.
-
-Note that in real data, it is possible for some points to be tied with the X and/or Y coordinates of the origin -- creating an ambuguity when computing the fraction of points in each quadrant.
-Since the Fransono Francheschini test attempts to define the maximal difference of the integrated probabilities, a natural solution would be to treat a point that is tied with the current X and/or Y coordinates of the origin as equally likely to have been drawn from any of the tied quadrants.
-In doing so, the integrated probability of the given point is exactly evenly distributed across the tied quadrants.
-For example, any data point sharing the same X or Y coordianted as the current origin is evenly distributed across the tied quadrants, with each of the two quadrants recieveing half a count.
-Any data point sharing the same X and Y coordiantes as the current origin  -- the origin included -- is evenly distributed across all quadrants, with all four quadrants recieveing a quarter count.
+It may be that some points are tied with the $X$ and/or $Y$
+coordinates of the origin, creating an ambiguity when computing the
+fraction of points in each quadrant. Since the 
+test attempts to define the maximal difference of the cumulative
+probabilities, a natural solution would be to treat a point that is
+tied with the current $X$ and/or $Y$ coordinates of the origin as equally
+likely to have been drawn from any of the tied quadrants. 
+Hence, any data point
+sharing the same $X$ or $Y$ coordinate as the origin is evenly
+distributed across the tied quadrants, with each of the two quadrants
+recieving half a count. Any data point sharing the both the same $X$ and $Y$
+coordinates as the current origin (including the origin itself) is evenly
+distributed across all quadrants, with all four quadrants recieveing a
+quarter count.
 
@@
 
-Using Monte Carlo simulation, Fasano and Franceschini created a look-up table of critical values of $D_{FF}$ as a function of $D_{FF}$, the sample size $N$, and the coefficient of correlation $r$.
-Teukolsky and Press later defined an approximate fit to this data in equation \ref{eq:4} for quick computation~\cite{numericalRecipes}, as follows.
-
-For the 1-Sample case,
+Using Monte Carlo simulation, Fasano and Franceschini created a
+look-up table of critical values of $D_{FF}$ as a function of
+$D_{FF}$, the sample size, and the coefficient of correlation $r$.
+\cite{numericalRecipes} later defined an approximate fit to the lookup table
+as follows. 
+%
+For the a single sample of size $n$,
 \begin{equation} \label{eq:4}
-\mathbb{P}(D > observed) = \Phi \left( \frac{D\sqrt{N}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{N})} \right) \, .
+\mathbb{P}(d_{FF} > D_{FF}) = \Phi \left( \frac{D_{FF}\sqrt{n}}{1+\sqrt{1-r^2}(0.25-0.75/\sqrt{n})} \right) \, .
 \end{equation}
-The 2-Sample case uses the same formula as above, but with the slight variation where
+\RBnote{Is $\Phi$ the same function defined in Eq~\ref{eq:1}?}
+The two sample case uses the same formula as above, but with the slight variation where
 \begin{equation} \label{eq:5}
-N = \frac{n_1n_2}{n_1+n_2}\, .
+n = \frac{n_1n_2}{n_1+n_2}\, .
 \end{equation}
-In both cases, $r$ is defined by
+In both cases, $r$ is defined in the usual way as
 \begin{equation} \label{eq:6}
-r = \frac{\sum_{i}^{}(x_i-\bar{x})(y_i-\bar{y})}{\sqrt{\sum_{i}^{}(x_i-\bar{x})^2}\sqrt{\sum_{i}^{}(y_i-\bar{y})^2}}\, .
+r = \frac{\sum_{i}(X_i-\bar{X})(Y_i-\bar{Y})}{\sqrt{\sum_{i}(X_i-\bar{X})^2}\sqrt{\sum_{i}(Y_i-\bar{Y})^2}}\, .
 \end{equation}
-
-\RBnote{In general, treat the equations as part of the sentences.  This means punctuating them -- see the examples above.}
-%use \verb|\,.| if the sentence ends with the equation, \verb|\,,| or \verb|\,;| if it's just the end of a clause.
-The \verb|\,| puts a little space before the punctuation to make it more readable; if you need more space, use \verb|\;| or \verb|$\quad$|.}
-
-% The basic Poisson regression model for count data is a special case of the GLM
-% framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
-% response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
-% $y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
-% $\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
-% log link and a linear predictor
-% %
-% \begin{equation} \label{eq:mean}
-% \log(\mu_i) \quad = \quad x_i^\top \beta,
-% \end{equation}
-% %
-% where the regression coefficients $\beta$ are estimated by maximum likelihood
-% (ML) using the iterative weighted least squares (IWLS) algorithm.
-%
-% \begin{leftbar}
-% Note that around the \verb|{equation}| above there should be no spaces (avoided
-% in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
-% not a new paragraph started.
-% \end{leftbar}
-%
-% \proglang{R} provides a very flexible implementation of the general GLM
-% framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
-% \pkg{stats} package. Its most important arguments are
-% \begin{Code}
-% glm(formula, data, subset, na.action, weights, offset,
-%   family = gaussian, start = NULL, control = glm.control(...),
-%   model = TRUE, y = TRUE, x = FALSE, ...)
-% \end{Code}
-% where \code{formula} plus \code{data} is the now standard way of specifying
-% regression relationships in \proglang{R}/\proglang{S} introduced in
-% \cite{Chambers+Hastie:1992}. The remaining arguments in the first line
-% (\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
-% standard  for setting up formula-based regression models in
-% \proglang{R}/\proglang{S}. The arguments in the second line control aspects
-% specific to GLMs while the arguments in the last line specify which components
-% are returned in the fitted model object (of class \class{glm} which inherits
-% from \class{lm}). For further arguments to \fct{glm} (including alternative
-% specifications of starting values) see \code{?glm}. For estimating a Poisson
-% model \code{family = poisson} has to be specified.
-%
-% \begin{leftbar}
-% As the synopsis above is a code listing that is not meant to be executed,
-% one can use either the dedicated \verb|{Code}| environment or a simple
-% \verb|{verbatim}| environment for this. Again, spaces before and after should be
-% avoided.
-%
-% Finally, there might be a reference to a \verb|{table}| such as
-% Table~\ref{tab:overview}. Usually, these are placed at the top of the page
-% (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
-% column headers and captions in sentence style, and if possible avoiding vertical
-% lines.
-% \end{leftbar}
-%
-% \begin{table}[t!]
-% \centering
-% \begin{tabular}{lllp{7.4cm}}
-% \hline
-% Type           & Distribution & Method   & Description \\ \hline
-% GLM            & Poisson      & ML       & Poisson regression: classical GLM,
-%                                            estimated by maximum likelihood (ML) \\
-%                &              & Quasi    & ``Quasi-Poisson regression'':
-%                                            same mean function, estimated by
-%                                            quasi-ML (QML) or equivalently
-%                                            generalized estimating equations (GEE),
-%                                            inference adjustment via estimated
-%                                            dispersion parameter \\
-%                &              & Adjusted & ``Adjusted Poisson regression'':
-%                                            same mean function, estimated by
-%                                            QML/GEE, inference adjustment via
-%                                            sandwich covariances\\
-%                & NB           & ML       & NB regression: extended GLM,
-%                                            estimated by ML including additional
-%                                            shape parameter \\ \hline
-% Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
-%                                            hurdle Poisson \\
-%                & NB           & ML       & Zero-inflated NB (ZINB),
-%                                            hurdle NB \\ \hline
-% \end{tabular}
-% \caption{\label{tab:overview} Overview of various count regression models. The
-% table is usually placed at the top of the page (\texttt{[t!]}), centered
-% (\texttt{centering}), has a caption below the table, column headers and captions
-% are in sentence style, and if possible vertical lines should be avoided.}
-% \end{table}
-
-%% -- Illustrations ------------------------------------------------------------
-
-%% - Virtually all JSS manuscripts list source code along with the generated
-%%   output. The style files provide dedicated environments for this.
-%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%%   or knitr using the render_sweave() hook - are used (without the need to
-%%   load Sweave.sty).
-%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-%% - The code input should use "the usual" command prompt in the respective
-%%   software system.
-%% - For R code, the prompt "R> " should be used with "+  " as the
-%%   continuation prompt.
-%% - Comments within the code chunks should be avoided - these should be made
-%%   within the regular LaTeX text.
 
@@
 In their paper, Fasano and Franceschini use Monte Carlo simulation to
-approximate the distribution of $D$ as a function of the sample
-size $N$ and the coefficient of correlation $r$. Unlike the 1-D KS
+approximate the distribution of $D_{FF}$ as a function of the sample
+size $n$ and the coefficient of correlation $r$. Notably, unlike the 1-D KS
 test, the distribution of
-$D_{FF}$ is not independent of the shape of the 2-D distribution of the
-underlying data. Fasano
-and Franceschini reasoned that the correlation between
-the two-dimensional distributions governs the resultant
-distribution of $D_{FF}$. For example, if the data sets are perfectly
+$D_{FF}$ is \textit{not} completely independent of the shape of the 2-D distribution of the
+underlying data, but depends on the correlations between the variables.
+In the case where the variables $X$ and $Y$  are perfectly
 correlated ($r = 1$), the 2-D distribution lies along a single line and thus
-the 1-D KS test could be used. At the other extreme, if the data sets are perfectly
-uncorrelated ($r = 0$), the 2-D distribution is independent in the $x$
-and $y$ directions and thus one could use a 1-D KS test on the marginal
+the 1-D KS test could be used; at the other extreme were $X$ and $Y$  are perfectly
+uncorrelated ($r = 0$), the 2-D distribution is independent in the $X$
+and $Y$ directions and one could apply the 1-D KS test on the marginal
 distributions on the 2-D plane.
 %
-Results from Monte Carlo simulation support these assumptions, showing
+Results from Monte Carlo simulation support these expectaions, showing
 that the distribution of $D$ is nearly identical for varying
@@
 The approximation
-by Teukolsky and Press (Eq~\ref{eq:4}--\ref{eq:5})
+by  \cite{numericalRecipes} (Eq~\ref{eq:4}--\ref{eq:5})
 can be used to test the significance levels for the 2-D K-S test using
@@
 
-As a caveat, to the fit produced by Teukolsky and Press are accurate when $N \gtrsim 20$, and the $p$-value is less than (more significant than) $\sim 0.20$.
-For $p > 0.20$, the distributional approximation may not be accurate.
-While this inaccuracy still allows a simple rejection decision to be made, it is sometimes useful to accurately quantify $p$ (such as if one was to do a cross-ctudy concordance analysis comparing $p$~values between studies as in~\citep{Ness-Cohn2020}).
+It has been noted that the approximation from  \citep{numericalRecipes} is only accurate when $n \gtrsim 20$ and the $p$-value is less than (more significant than) $\sim 0.2$ \citep{Babu2006}.
+While this inaccuracy still allows a simple rejection decision to be made at any $\alpha\leq0.2$, it is sometimes useful to quantify large $p$ more exactly (such as if one was to do a cross-study concordance analysis comparing $p$~values between studies as in~\cite{Ness-Cohn2020}), and to apply it to smaller datasets.
 %
-\RBnote{I'm not sure what this next sentence is getting at; elaborate or cut. - ENC: removed the next line}
+%\RBnote{I'm not sure what this next sentence is getting at; elaborate or cut. - ENC: removed the next line}
 %Moreover, additional arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
-
+%
 To address these limitations, one can bootstrap the significance levels for the particular multidimensional statistic directly from the particular data set under study.
@@
 
-The \pkg{fasano.franceschini.test} R package implements a parallelized bootstrapping procedure.
-The marginal distribution from 2-dimensional data set is resampled with replacement to generate randomized 2-dimensional data sets \code{nBootStrap} times.
-The frequency count by quadrant is performed for each bootstrapped resampling as described above to compute the $D_{FF}$.
-The observed $D_{FF}$ is then compared to the distribution of bootstrapped $D_{FF}$ to compute a $p$~value.
-The bootstrapped version of the Fasano--Franceschini test can be implemented as follows (see \fct{fasano.franceschini.test} for further source code details and implementation).
+The \pkg{fasano.franceschini.test} R package implements a parallelized
+bootstrapping procedure. The marginal distribution from 2-dimensional
+data set is resampled with replacement to generate randomized
+2-dimensional data sets \code{nBootStrap} times. The frequency count
+by quadrant is performed for each bootstrapped resampling as described
+above to compute the $D_{FF}$. The observed $D_{FF}$ is then compared
+to the distribution of bootstrapped $D_{FF}$ to compute a $p$~value.
+The bootstrapped version of the Fasano--Franceschini test can be
+implemented as follows (see \fct{fasano.franceschini.test} for further
+source code details and implementation).
 
@@
 \end{CodeChunk}
-
-Note, bootstrapping may significantly increase run time.
 To improve run time, one may adjust the \code{cores} parameter; see the R \pkg{parallel} package and the \fct{mclapply} function for further details.
 
-\subsection{Computational Efficency}
-The main distinction between the Peacock and Fasano--Franceschini test comes down to compute time. For a benchmarking of the algorithms see~\citep{Lopes2007}.  Briefly,
+\subsection{Computational efficency}
+\RBnote{This is weird without any data.  Can you benchmark using your package without bootstrap; your package with bootstrap and varying cores; and Peacock.test package to illustrate scaling?}
+The main distinction between the Peacock and Fasano--Franceschini tests is in computational efficiency, as
+illustrated by the benchmarking studies in \cite{Lopes2007}.  Briefly,
 %\begin{itemize}
@@
 (See the \pkg{Peacock.test} package for \proglang{R} implementation of the Peacock test.)
-
-% For a simple illustration of basic Poisson and NB count regression the
-% \code{quine} data from the \pkg{MASS} package is used. This provides the number
-% of \code{Days} that children were absent from school in Australia in a
-% particular year, along with several covariates that can be employed as regressors.
-% The data can be loaded by
-% %
-% \begin{CodeChunk}
-% \begin{CodeInput}
-% R> data("quine", package = "MASS")
-% \end{CodeInput}
-% \end{CodeChunk}
-% %
-% and a basic frequency distribution of the response variable is displayed in
-% Figure~\ref{fig:quine}.
-%
-% \begin{leftbar}
-% For code input and output, the style files provide dedicated environments.
-% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-% hook. Please make sure that all code is properly spaced, e.g., using
-% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-% use ``the usual'' command prompt in the respective software system. For
-% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-% the continuation prompt. Generally, comments within the code chunks should be
-% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-% before and after code input/output should be avoided (see above).
-% \end{leftbar}
-%
-% \begin{figure}[t!]
-% \centering
-% \includegraphics{article-visualization}
-% \caption{\label{fig:quine} Frequency distribution for number of days absent
-% from school.}
-% \end{figure}
-%
-% As a first model for the \code{quine} data, we fit the basic Poisson regression
-% model. (Note that JSS prefers when the second line of code is indented by two
-% spaces.)
-% %
-% \begin{CodeChunk}
-% \begin{CodeInput}
-% R> m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-% +    family = poisson)
-% \end{CodeInput}
-% \end{CodeChunk}
-% %
-% To account for potential overdispersion we also consider a negative binomial
-% GLM.
-% %
-% \begin{CodeChunk}
-% \begin{CodeInput}
-% R> library("MASS")
-% R> m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
-% \end{CodeInput}
-% \end{CodeChunk}
-% %
-% In a comparison with the BIC the latter model is clearly preferred.
-% %
-% \begin{CodeChunk}
-% \begin{CodeInput}
-% R> BIC(m_pois, m_nbin)
-% \end{CodeInput}
-% \begin{CodeOutput}
-%        df      BIC
-% m_pois 18 2046.851
-% m_nbin 19 1157.235
-% \end{CodeOutput}
-% \end{CodeChunk}
-% %
-% Hence, the full summary of that model is shown below.
-% %
-% \begin{CodeChunk}
-% \begin{CodeInput}
-% R> summary(m_nbin)
-% \end{CodeInput}
-% \begin{CodeOutput}
-% Call:
-% glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-%     init.theta = 1.60364105, link = log)
-%
-% Deviance Residuals:
-%     Min       1Q   Median       3Q      Max
-% -3.0857  -0.8306  -0.2620   0.4282   2.0898
-%
-% Coefficients: (1 not defined because of singularities)
-%             Estimate Std. Error z value Pr(>|z|)
-% (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
-% EthN        -0.24591    0.39135  -0.628  0.52977
-% SexM        -0.77181    0.38021  -2.030  0.04236 *
-% AgeF1       -0.02546    0.41615  -0.061  0.95121
-% AgeF2       -0.54884    0.54393  -1.009  0.31296
-% AgeF3       -0.25735    0.40558  -0.635  0.52574
-% LrnSL        0.38919    0.48421   0.804  0.42153
-% EthN:SexM    0.36240    0.29430   1.231  0.21818
-% EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876
-% EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 **
-% EthN:AgeF3   0.04721    0.44883   0.105  0.91622
-% EthN:LrnSL   0.06847    0.34040   0.201  0.84059
-% SexM:AgeF1   0.02257    0.47360   0.048  0.96198
-% SexM:AgeF2   1.55330    0.51325   3.026  0.00247 **
-% SexM:AgeF3   1.25227    0.45539   2.750  0.00596 **
-% SexM:LrnSL   0.07187    0.40805   0.176  0.86019
-% AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870
-% AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363
-% AgeF3:LrnSL       NA         NA      NA       NA
-% ---
-% Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
-%
-% (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)
-%
-%     Null deviance: 235.23  on 145  degrees of freedom
-% Residual deviance: 167.53  on 128  degrees of freedom
-% AIC: 1100.5
-%
-% Number of Fisher Scoring iterations: 1
-%
-%
-%               Theta:  1.604
-%           Std. Err.:  0.214
-%
-%  2 x log-likelihood:  -1062.546
-% \end{CodeOutput}
-% \end{CodeChunk}
 
@@
   (iii) parallelized bootstrapping for improved statistical testing.
-\RBnote{CRAN? - ENC: Ready to Submit to CRAN}
   See the complete package documentation and source code via the Comprehensive
@@
 
-\RBnote{Any other package dependencies? - ENC: added dependenceis parallel and methods}
+%\RBnote{Any other package dependencies? - ENC: added dependenceis parallel and methods}
 
@@
 
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-% \begin{appendix}
-%
-% \section{More technical details} \label{app:technical}
-%
-% \begin{leftbar}
-% Appendices can be included after the bibliography (with a page break). Each
-% section within the appendix should have a proper section title (rather than
-% just \emph{Appendix}).
-%
-% For more technical style details, please check out JSS's style FAQ at
-% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-% which includes the following topics:
-% \begin{itemize}
-%   \item Title vs.\ sentence case.
-%   \item Graphics formatting.
-%   \item Naming conventions.
-%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
-%   \item Trouble shooting.
-%   \item Many other potentially helpful details\dots
-% \end{itemize}
-% \end{leftbar}
-%
-%
-% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-%
-% \begin{leftbar}
-% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-% formats of author-year citations and allow to include additional details (e.g.,
-% pages, chapters, \dots) in brackets. In case you are not familiar with these
-% commands see the JSS style FAQ for details.
-%
-% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-% when acquiring the entries automatically from mixed online sources. However,
-% it is important that informations are complete and presented in a consistent
-% style to avoid confusions. JSS requires the following format.
-% \begin{itemize}
-%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-%     be used in the references.
-%   \item Titles should be in title case.
-%   \item Journal titles should not be abbreviated and in title case.
-%   \item DOIs should be included where available.
-%   \item Software should be properly cited as well. For \proglang{R} packages
-%     \code{citation("pkgname")} typically provides a good starting point.
-% \end{itemize}
-% \end{leftbar}
-%
-% \end{appendix}
-
-%% -----------------------------------------------------------------------------
-
 \end{document}

### 2021-06-08 journalMS/article.tex
updated minor edits

@@
 \Abstract{
-\RBnote{Perhaps briefly state what the KS test is.}
-The versatility of the univariate Kolmogorov-Smirnov (KS) test has
+The univariate Kolmogorov-Smirnov (KS) test is a non--parametric
+statistical test designed to assess whether a set of data is
+consistent with a given probability distribution (or, in the two-sample
+case,  whether the two
+samples come from the same underlying distribution).
+The versatility of the KS test has
 made it a cornerstone of statistical analysis and is commonly used
@@
 cumulative distribution functions.  Rather than looking over all
-$X_i, Y_j: i,j={1,n}$, Fasano and Franceschini only used the detected points
-$X_i, Y_i : i={1,n}$. 
-\RBnote{I don't follow the preceding line? (Notation?)}
+$X_i, Y_j \,;\; i,j \in [1,n]$, Fasano and Franceschini only used the detected points
+$X_i, Y_i \,;\; i \in [1,n]$.
 Thus for any given $n$ points in a two-dimensional space, those $n$ points define $4n$ (rather than
@@
 second sample. $D_{FF,1}$ and $D_{FF,2}$  are then averaged to compute
-the overall $D_{FF}$ for hypothesis testing. \RBnote{clarify: is
-$D_{FF}=(D_{FF,1}+D_{FF,2})/2$, or is it weighted by the $n$?}
+the overall $D_{FF}$ for hypothesis testing -- e.g.
+$D_{FF}=(D_{FF,1}+D_{FF,2})/2$.
 
@@
 \end{equation}
-\RBnote{Is $\Phi$ the same function defined in Eq~\ref{eq:1}?}
+where $\Phi$ is the same function defined in Eq~\ref{eq:1}.
 The two sample case uses the same formula as above, but with the slight variation where

### 2021-06-10 journalMS/article.tex
add FF test benchmarking

@@
 \subsection{Computational efficency}
-\RBnote{This is weird without any data.  Can you benchmark using your package without bootstrap; your package with bootstrap and varying cores; and Peacock.test package to illustrate scaling?}
+
+\begin{figure}[hbt]
+\centering
+\includegraphics{benchmark}
+\caption{\label{fig:bmark} \textbf{A:} Benchmarking of the Fasano--Franceschini versus the Peacock test under four distinct numbers of points N (e.g.
+10, 100, 1000, 5000).
+Points represent the average of 10 replicate runs.
+\textbf{B:} Benchmarking of Fasano--Franceschini test bootstrapping procedure broken up by the number of points N (e.g.
+10, 100, 1000, 5000).
+Within each panel, lines are colored by the number of bootstraps run (e.g.
+0, 10, 100, 1000).
+Points represent the average of 10 replicate runs.
+}
+\end{figure}
+
 The main distinction between the Peacock and Fasano--Franceschini tests is in computational efficiency, as
 illustrated by the benchmarking studies in \cite{Lopes2007}. Briefly,
-%\begin{itemize}
-%\item The Fasano--Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale.
-%\item The Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
-%\end{itemize}
-the Fasano--Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale, while the Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
+the Fasano--Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale, while the Peacock test is more computationally expensive, and runs in $O(n^3)$ time (\textbf{Figure~\ref{fig:bmark}A}).
+While the implentation of the bootstrapping procedure increases runtime in comparison to the approximate fit from  \citep{numericalRecipes}; parallelization of the Fasano--Franceschini test shows $\sim 4X$ reduction in run time from 1 to 8 cores across bootstraps and number of points.
+\textbf{Figure~\ref{fig:bmark}B}.
 %

### 2021-06-10 journalMS/article.R

@@ -17,10 +17,12 @@ library("ggplot2")
 library("viridis")
 library("mvtnorm")
 library("gridExtra")
+library("ggrepel")
 library("ggpubr")
+library("dplyr")
 
 # set to T to plot figures to PDF Output
-pdfOutPuts <- F
+pdfOutPuts <- T
 
 #' ## fasano.franceschini.test Usage
 
@@ -267,6 +269,76 @@ p4 <- ggplot(data, aes(x = x, y = y)) +
 pTotal <- ggarrange(p3, p2, p1, p4,
   labels = c("A", "B", "C", "D"),
   ncol = 2, nrow = 2
+)
+pTotal
+if (pdfOutPuts) {
+  dev.off()
+}
+
+#' ## Visualization - Benchmarking
+
+if (pdfOutPuts) {
+  png(file = paste0("journalMS/benchmark.png"), units = "in", res = 300, width = 12, height = 6)
+}
+#load files
+benchmarkingFiles <- list.files(path = "journalMS/benchmarking", recursive = TRUE,
+                                pattern = "\\.txt$",
+                                full.names = TRUE)
+
+benchmarkingDatalist <- lapply(benchmarkingFiles, function(x) {read.delim(x)})
+benchmarkingData <- do.call("rbind", benchmarkingDatalist)
+
+theme_update(
+  plot.title = element_text(hjust = 0.5, face = "bold"),
+  axis.title = element_text(face = "bold"), axis.text = element_text(face = "bold")
+)
+
+p5 <- benchmarkingData %>%
+    filter(cores == 1) %>%
+    filter(boots == 0) %>%
+    mutate(label = if_else(N == 5000, as.character(method), NA_character_)) %>%
+  ggplot(aes(x = N, y = time, col = method)) +
+  geom_line(linetype = "dashed") +
+  geom_point() +
+  scale_color_manual(labels = c("Fasano Franceschini", "Peacock"), values=c(S1col, S2col)) +
+  xlab("Number of Points (N)") +
+  ylab("Time (s)") +
+  guides(col=guide_legend(title="Method")) +
+  theme(legend.position = "bottom",
+        panel.background = element_blank(),
+        panel.grid.major.x = element_blank() ,
+        panel.grid.major.y = element_line( size=0.1, color="black" )) +
+  ggtitle("Fasano Franceschini versus Peacock Test Benchmark")
+
+
+nNames <- c(
+  `10` = "N = 10",
[…]

### 2021-06-10 journalMS/benchmarking.R

@@ -1,0 +1,69 @@
+# benchmarking Scripts
+# get arguments from command line
+args <- commandArgs(trailingOnly = TRUE)
+args <- as.numeric(args)
+print(args)
+
+# install.packages("Peacock.test")
+# devtools::install_github("eddelbuettel/rbenchmark")
+library(Peacock.test)
+library(devtools)
+load_all("./")
+library(fasano.franceschini.test)
+library(rbenchmark)
+
+sample1 <- data.frame(
+  x =  rnorm(args[1] / 2, mean = 0, sd = 1),
+  y =  rnorm(args[1] / 2, mean = 0, sd = 1)
+)
+
+sample2 <- data.frame(
+  x =  rnorm(args[1] / 2, mean = 0, sd = 1),
+  y =  rnorm(args[1] / 2, mean = 0, sd = 1)
+)
+
+if (args[2] == 0 & args[3] == 1) {
+  bench <- benchmark(
+    replications = 10,
+    "peacock" = {
+      peacock2(x = sample1, y = sample2)
+    },
+    "fasano" = {
+      fasano.franceschini.test(S1 = sample1, S2 = sample2, nBootstrap = args[2], cores = args[3])
+    }
+  )
+
+  out <- bench$elapsed
+
+  results <- data.frame(
+    N = args[1],
+    boots = args[2],
+    cores = args[3],
+    method = c("fasano", "peacock"),
+    time = out
+  )
+
+} else {
+
+  bench <- benchmark(
+    replications = 10,
+    "fasano" = {
+      fasano.franceschini.test(S1 = sample1, S2 = sample2, nBootstrap = args[2], cores = args[3])
+    }
+  )
+
+  out <- bench$elapsed
+
+  results <- data.frame(
+    N = args[1],
+    boots = args[2],
+    cores = args[3],
+    method = c("fasano"),
+    time = out
+  )
+}
+
+write.table(results,
+  file = paste0("~/Desktop/fasano.franceschini.test/journalMS/benchmarking/result_", args[1], "_", args[2], "_", args[3], ".txt"), sep = "\t",
+  row.names = FALSE, col.names = TRUE
+)

### 2021-06-10 journalMS/run_benchmark_local.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+while IFS=$'\t' read N boots cores
+do
+	Rscript benchmarking.R $N $boots $cores
+	# print out the job id for reference later
+	echo "JobID = ${JOB} for parameters $fileName submitted on `date`"
+done < parameters_benchmarking.txt
+exit
+
+# make this file executable and then run from the command line
+# chmod u+x run_limoRhyde_local.sh
+# ./run_limoRhyde_local.sh

### 2021-06-11 journalMS/article.R
added avg benchmark time for 10 runs

@@ -22,7 +22,7 @@ library("ggpubr")
 library("dplyr")
 
 # set to T to plot figures to PDF Output
-pdfOutPuts <- T
+pdfOutPuts <- F
 
 #' ## fasano.franceschini.test Usage
 
@@ -288,16 +288,23 @@ benchmarkingFiles <- list.files(path = "journalMS/benchmarking", recursive = TRU
 benchmarkingDatalist <- lapply(benchmarkingFiles, function(x) {read.delim(x)})
 benchmarkingData <- do.call("rbind", benchmarkingDatalist)
 
+#Compute avg time. Current time is for 10 replicates
+#see benchmarking.R
+benchmarkingData <- benchmarkingData %>%
+  mutate(avg_time = time/10)
+
 theme_update(
   plot.title = element_text(hjust = 0.5, face = "bold"),
   axis.title = element_text(face = "bold"), axis.text = element_text(face = "bold")
 )
 
+
+
 p5 <- benchmarkingData %>%
     filter(cores == 1) %>%
     filter(boots == 0) %>%
     mutate(label = if_else(N == 5000, as.character(method), NA_character_)) %>%
-  ggplot(aes(x = N, y = time, col = method)) +
+  ggplot(aes(x = N, y = avg_time, col = method)) +
   geom_line(linetype = "dashed") +
   geom_point() +
   scale_color_manual(labels = c("Fasano Franceschini", "Peacock"), values=c(S1col, S2col)) +
@@ -320,7 +327,7 @@ nNames <- c(
 
 p6 <- benchmarkingData %>%
   filter(method != "peacock") %>%
-  ggplot(aes(x = cores, y = time, group = boots, col = as.character(boots))) +
+  ggplot(aes(x = cores, y = avg_time, group = boots, col = as.character(boots))) +
   geom_line(linetype = "dashed") +
   geom_point() +
   scale_color_manual(values= c("#fef0d9","#fdcc8a","#fc8d59","#d7301f")) +
@@ -328,7 +335,7 @@ p6 <- benchmarkingData %>%
   xlab("Number of Cores") +
   ylab("Time (s)") +
   guides(col=guide_legend(title="Bootstraps")) +
-  scale_y_continuous(trans='log10') +
+  scale_y_continuous(trans='log10', labels = scales::comma) +
   theme(legend.position = "bottom",
         panel.background = element_blank(),
         panel.grid.major.x = element_blank(),

### 2021-06-17 journalMS/article.tex
RB edits, penultimate

--- 2021-06-08 journalMS/article.tex
@@
 \newcommand{\RBnote}[1]{\textcolor{red}{#1}}
+
+\hyphenation{Fa-sa-no=Fran-ces-chi-ni}
 
@@
 
-% R -> Peacock test (computationally inificent)
-% Python <- Implementation
-% C <- Implementation
 The Kolmogorov--Smirnov (KS) is a non--parametric, univariate
@@
 implementation of the test not being
-distribution-free~\citep{Babu2006}. Furthermore, the \proglang{C}
-implementation statistical testing is based on a fit to Monte Carlo
-Simulation and is only valid for significance levels $\alpha \lessapprox 0.20$,
-which may present potential issues in some analysis.
+distribution-free~\citep{Babu2006}. Furthermore, in the \proglang{C}
+implementation, statistical testing is based on a fit to Monte Carlo
+simulation that is only valid for significance levels $\alpha \lessapprox 0.20$.
 
@@
 \pkg{fasano.franceschini.test} package provides two improvements over
-the current 2-D KS test on the Comprehensive \proglang{R} Archive
+the current 2-D KS test available on the Comprehensive \proglang{R} Archive
 Network (CRAN): (i) the Fasano and Franceschini test has been shown to
@@
 for improved significance testing and mitigates the limitations of
-the test brought noted by~\citep{Babu2006}.
+the test brought noted by~\citet{Babu2006}.
 
@@
 distribution~\citep{Stephens1992a}.
-%While a comprehensive discussion pertaining the intricacies of the 1-D KS test can be found elsewhere~\citep{Stephens1992a}, we will present a brief overview as motivation for the basis and intuition of the 2-D KS test.
-%
-%As described previously, the 1-D KS test can be divided into two variants of the same statistical procedure:
-%
-%\begin{leftbar}
-%\textbf{The One-Sample Test}
-%\begin{itemize}
-%\item Compares the data to analytical distribution.
-%\end{itemize}
-%\textbf{The Two-Sample Test}
-%\begin{itemize}
-%\item Compares two data samples to determine if they are drawn from the same underlying probability distribution.
-%\end{itemize}
-%\end{leftbar}
-%
-%In both cases, the KS test works as follows:
 %
@@
 \textbf{Figure~\ref{fig:kstest1D}}.
-%.
-In the following example, the orange and blue curve in the left panel represents the underlying distribution (pdf) for sample1 and sample2, respectively (\textbf{Figure \ref{fig:kstest1D}}).
-The eCDF is then defined by data drawn from these underlying distribution, and the D--stat is the computed as the largest between difference between the eCDF (black dotted line) in the right panel.
 
@@
 \includegraphics{pdfvsCDF}
-\caption{\label{fig:kstest1D} \textbf{LEFT:} Probability density function (PDF) of two normal distributions: orange sample~1, $\mathcal{N}(\mu = 0,\,\sigma^{2} = 1)$; blue sample~2, $\mathcal{N}(\mu = 5,\,\sigma^{2} = 1)$.
-\textbf{RIGHT:} Cumulative density functions (CDF) of the two PDFs; the black dotted line represents the maximal absolute difference between the CDFs ($D_{KS}$).
-}
+\caption{\label{fig:kstest1D} \textbf{LEFT:} Probability density
+function (PDF) of two normal distributions: orange sample~1,
+$\mathcal{N}(\mu = 0,\,\sigma^{2} = 1)$; blue sample~2,
+$\mathcal{N}(\mu = 5,\,\sigma^{2} = 1)$. \textbf{RIGHT:} Cumulative
+density functions (CDF) of the two PDFs; the black dotted line
+represents the maximal absolute difference between the CDFs
+($D_{KS}$).  }
 \end{figure}
 
-%Kolmogorov and Smirnov work mathematically defined the distribution for $D_{KS}$ to be distribution free, and further derived a table of critical values for hypothesis testing.
-Kolmogorov and Smirnov work was later refined by Kendall and Stuart ~\citep{Kendall1946} to define the integrated probability distribution for large values of n  ($n \geq 80$), having the asymptotic expression:
-
-In the large--sample limit ($n \geq 80$), it can be shown~\citep{Kendall1946} that $D_{KS}$ converges in distribution to
+In the large--sample limit ($n \geq 80$), it can be
+shown~\citep{Kendall1946} that $D_{KS}$ converges in distribution to
 \begin{equation} \label{eq:1}
-\Phi(\lambda) = 2 \sum_{k=1}^{\infty} -1^{k-1}e^{-2k^2\lambda^2} \,.
+D_{KS} \overset{d}{\rightarrow} \Phi(\lambda) = 2 \sum_{k=1}^{\infty} -1^{k-1}e^{-2k^2\lambda^2} \,.
 \end{equation}
-%Given the maximum absolute difference ($D_{KS}$) and \textbf{Equation~\ref{eq:1}}, one can  compute a $p$-value for  hypothesis testing to determine if there is a significance difference between the cumulative density functions of the data and model (one-sample), or between the two data sets (two-sample).
-%
-%\textbf{For the 1-Sample case}, the $p$-value probability of a given D is defined by
 In the one-sample case with a sample of size $n$, the $p$~value is given by
@@
 in the two-sample case, the $p$~value is given by
-%where $N$ is the number of data points sampled. A $p$-value $< 0.05$, indicates a significant difference between the samples.
-%
-%\textbf{In the 2-Sample case}, the $p$-value probability of a given D is defined by
 \begin{equation} \label{eq:3}
@@
 \end{equation}
-%
 where $n_1$ and $n_2$ are the number of observations in the first and second samples respectively.
-%A $p$-value $< 0.05$, indicates a significant difference between the samples.
 
 \subsection{Higher dimensional variations: Peacock Test (1983) and Fasano--Franceschini Test (1987)}
-%Similar to the 1-D KS test, the obvious solution to extend the theory into higher dimensional would be to define the maximal cumulative difference between the two 2-D distributions.
-This however, presents an issue as the cumulative distribution function is not well defined in more than one dimension.
-In 2-D, there are 4 ways (3 independent ways) of defining the cumulative distribution, since the direction in which we define choose to order the points in the x and y are arbitrary (\textbf{Figure \ref{fig:kstest2Dissue}}).
-In N dimensional space there are $2^{N}-1$ independent ways of defining the cumulative distribution function~\citep{Peacock1983}.
 Extending the above to two or higher dimension is complicated by the
 fact that CDFs are not well-defined in more than one dimension.  In
-2-D, there are 4 ways (3 independent ways) of defining the cumulative
+2-D, there are 4 ways (3 independent) of defining the cumulative
 distribution, since the direction in which we order the $x$ and $y$
-points are arbitrary (\textbf{Figure \ref{fig:kstest2Dissue}}); more
+points is arbitrary (\textbf{Figure \ref{fig:kstest2Dissue}}); more
 generally, in $k$-dimensional space there are $2^{k}-1$ independent
@@
 $\mathbb{P}(x > X, y > Y)$ (D).
-%$P(x < X, y < Y) + P(x < X, y > Y) + P(x > X, y < Y) = 1 - P(x > X, y > Y)$
 }
@@
 ordering combinations into account. Peacock's test thus computes the
-total integrated probability---i.e.\,fraction of data---tin each of
-the four quadrants around all possible points in the data. For
-example, for a given $n$ points in a two-dimensional space, the
+total probability---i.e.\,fraction of data---in each of
+the four quadrants around all possible tuples in the data. For
+example, for $n$ points in a two-dimensional space, the empirical
 cumulative distribution functions is calculated in the $4n^2$
-quadrants of the plane defined by all pairs ($X_i$ ,$Y_j$), $X_i$ and
-$Y_j$ being coordinates of any pairs of points in the given sample. By
-ranging over all possible pairs of data points and quadrants, the
+quadrants of the plane defined by all pairs $(X_i, Y_j): i,j\in[1,n]$, where $X_i$ and
+$Y_j$ are any observed value of $x$ and $y$ (whether or not they are observed as a pair).
+There are $n^2$ such pairs, each of which can define four quadrants in the 2-D plane;
+by ranging over all possible pairs of data points and quadrants, the
 2-dimensional $D$ statistic is defined by the maximal difference of
@@
 The variation defined by \cite{Fasano1987}  was to only consider
-quadrants centered in each point of the given samples to compute the
-cumulative distribution functions.  Rather than looking over all
-$X_i, Y_j \,;\; i,j \in [1,n]$, Fasano and Franceschini only used the detected points
-$X_i, Y_i \,;\; i \in [1,n]$.
+quadrants centered on each observed $(x, y)$ pair to compute the
+cumulative distribution functions. That is, rather than looking over all $n^2$ points
+${(X_i, Y_j): i,j \in [1,n]}$, Fasano and Franceschini only use the observed $n$ points
+${(X_i, Y_i): i \in [1,n]}$.
 Thus for any given $n$ points in a two-dimensional space, those $n$ points define $4n$ (rather than
@@
 second sample. $D_{FF,1}$ and $D_{FF,2}$  are then averaged to compute
-the overall $D_{FF}$ for hypothesis testing -- e.g.
+the overall $D_{FF}$ for hypothesis testing,
 $D_{FF}=(D_{FF,1}+D_{FF,2})/2$.
@@
 distributed across the tied quadrants, with each of the two quadrants
-recieving half a count. Any data point sharing the both the same $X$ and $Y$
+receiving half a count. Any data point sharing the both the same $X$ and $Y$
 coordinates as the current origin (including the origin itself) is evenly
-distributed across all quadrants, with all four quadrants recieveing a
+distributed across all quadrants, with all four quadrants receiving a
 quarter count.
 
-\subsection{Defining the Fasano and Franceschini null distribution}
+\subsection{Null distribution of $D_{FF}$}
 
@@
 %
-For the a single sample of size $n$,
+For a single sample of size $n$,
 \begin{equation} \label{eq:4}
@@
 \end{equation}
-where $\Phi$ is the same function defined in Eq~\ref{eq:1}.
+where $\Phi(\cdot)$ is as defined in Eq~\ref{eq:1}.
 The two sample case uses the same formula as above, but with the slight variation where
@@
 
-\subsection{Fasano--Franceschini Test Usage}
+\subsection{Fasano--Franceschini test usage}
 
@@
 and $Y$ directions and one could apply the 1-D KS test on the marginal
-distributions on the 2-D plane.
+distributions.
 %
-Results from Monte Carlo simulation support these expectaions, showing
+Results from Monte Carlo simulation support these expectations, showing
 that the distribution of $D$ is nearly identical for varying
@@
 
-\subsection{Bootstrap Version of the Fasano--Franceschini Test}
+\subsection{Bootstrap version of the Fasano--Franceschini test}
 
-It has been noted that the approximation from  \citep{numericalRecipes} is only accurate when $n \gtrsim 20$ and the $p$-value is less than (more significant than) $\sim 0.2$ \citep{Babu2006}.
-While this inaccuracy still allows a simple rejection decision to be made at any $\alpha\leq0.2$, it is sometimes useful to quantify large $p$ more exactly (such as if one was to do a cross-study concordance analysis comparing $p$~values between studies as in~\cite{Ness-Cohn2020}), and to apply it to smaller datasets.
+It has been noted that the approximation from
+\citet{numericalRecipes} is only accurate when $n \gtrsim 20$ and the
+$p$-value is less than (more significant than) $\sim 0.2$
+\citep{Babu2006}. While this inaccuracy still allows a simple
+rejection decision to be made at any $\alpha\leq0.2$, it is sometimes
+useful to quantify large $p$ more exactly (such as if one was to do a
+cross-study concordance analysis comparing $p$~values between studies
+as in~\cite{Ness-Cohn2020}), and to apply it to smaller datasets.
 %
-%\RBnote{I'm not sure what this next sentence is getting at; elaborate or cut. - ENC: removed the next line}
-%Moreover, additional arguments have been made to the validity of the $>1$-D KS tests as not being distribution-free \citep{Babu2006}.
-%
-To address these limitations, one can bootstrap the significance levels for the particular multidimensional statistic directly from the particular data set under study.
-As Fasano and Franceschini's paper was originally released in 1987, this approach was unfeasible at scale.
-Today, modern computers can rapidly compute a bootstrapped null distribution of $D_{FF}$ from the data to test significance.
+To address these limitations, one can bootstrap the significance
+levels for the particular multidimensional statistic directly from the
+particular data set under study. As Fasano and Franceschini's paper
+was originally released in 1987, this approach was unfeasible at
+scale.  Today, modern computers can rapidly compute a bootstrapped
+null distribution of $D_{FF}$ from the data to test significance.
 
@@
 The bootstrapped version of the Fasano--Franceschini test can be
-implemented as follows (see \fct{fasano.franceschini.test} for further
+run as follows (see \fct{fasano.franceschini.test} for further
 source code details and implementation).
@@
 \subsection{Computational efficency}
-\RBnote{This is weird without any data.  Can you benchmark using your package without bootstrap; your package with bootstrap and varying cores; and Peacock.test package to illustrate scaling?}
-The main distinction between the Peacock and Fasano--Franceschini tests is in computational efficiency, as
-illustrated by the benchmarking studies in \cite{Lopes2007}.  Briefly,
-%\begin{itemize}
-%\item The Fasano--Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale.
-%\item The Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
-%\end{itemize}
-the Fasano--Franceschini test runs in $O(n^2)$ time, and is advantageous to the Peacock test as data set sizes scale, while the Peacock test is more computationally expensive, and runs in $O(n^3)$ time.
-%
-(See the \pkg{Peacock.test} package for \proglang{R} implementation of the Peacock test.)
+
+\begin{figure}[hbt]
+\centering
+\includegraphics{benchmark}
+\caption{\label{fig:bmark} Computational efficiency benchmarks.
+\textbf{A:} Runtime of the 
+Fasano--Franceschini test relative to the Peacock test at four different sample sizes
+($n=10, 100, 1000, 5000$).  Points represent the
+average of 10 benchmark runs. \textbf{B:} Runtime of the
+Fasano--Franceschini bootstrapping procedure for various sample sizes
+($n= 10, 100, 1000, 5000$) as a function of the number of cores used. Within each panel,
+lines are colored by the number of bootstrap iterations (no bootstrap, 10, 100,
+1000). Points represent the average of 10 benchmark runs. Note the logarithmic $y$-axis
+in (B).}
+\end{figure}
+
+To assess the computational efficiency, we benchmarked the package as follows.
+\RBnote{Please explain the benchmark here; make note also of the specs of the machine that was used,
+inc # cores, processor speed, memory, and OS.}
+The main distinction between the Peacock and Fasano--Franceschini
+tests is in computational efficiency, with  Fasano--Franceschini
+scaling as $O(n^2)$ relative to Peacock's complexity of  $O(n^3)$
+\citep{Lopes2007}.  Our benchmarks also show this advantage, as shown
+in \textbf{Figure~\ref{fig:bmark}A}.  While the implentation of the
+bootstrapping procedure increases runtime in comparison to the
+approximate fit from  \citet{numericalRecipes}, parallelization of the
+Fasano--Franceschini test shows a four-fold reduction in run time when
+parallelized across 8 cores. \textbf{Figure~\ref{fig:bmark}B}.
+% (See the \pkg{Peacock.test} package for \proglang{R} implementation of the Peacock test.)
 
@@
 It improves upon existing packages by implementing
-  (i) a fast algorithm;
-  (ii) a extension to the Fasano and Franceschini test for dealing with ties; and
-  (iii) parallelized bootstrapping for improved statistical testing.
-  See the complete package documentation and source code via the Comprehensive
+  a fast algorithm and 
+  a parallelized bootstrapping procedure for improved statistical testing.
+  Complete package documentation and source code is available via the Comprehensive
 \proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/} or the package website at \url{https://nesscoder.github.io/fasano.franceschini.test/}.
+\url{https://CRAN.R-project.org/} and the package website at \url{https://nesscoder.github.io/fasano.franceschini.test/}.
 

### 2021-06-17 journalMS/article.tex
added benchmarking method info

@@
 \usepackage{amssymb}
+\usepackage{amsmath}
 
@@
 
-\hyphenation{Fa-sa-no=Fran-ces-chi-ni}
+%\hyphenation{Fa-sa-no=Fran-ces-chi-ni}
 
@@
 To assess the computational efficiency, we benchmarked the package as follows.
-\RBnote{Please explain the benchmark here; make note also of the specs of the machine that was used,
-inc # cores, processor speed, memory, and OS.}
+Using the \pkg{rbenchmark} package to evaluate runtime, the Fasano--Franceschini
+test and Peacock test were run under four different samples sizes
+($n=10, 100, 1000, 5000$), with 10 replicates for each run. The
+Fasano--Franceschini test bootstrap precedure was further evaluated
+under four different bootstrap iterations (no bootstrap, 10, 100, 1000),
+again using 10 replicates for each run. Reported results represent the
+average run time of the 10 replicate benchmarks. All benchmark tests were run on a 2018 macBook
+Pro Mac (macOS Catalina) with a 2.7-GHz Quad-Core Intel Core i7 processor and 16 GB of 2133 MHz LPDDR3 memory.
+
 The main distinction between the Peacock and Fasano--Franceschini
@@
 Fasano--Franceschini test shows a four-fold reduction in run time when
-parallelized across 8 cores. \textbf{Figure~\ref{fig:bmark}B}.
+parallelized across 8 cores (\textbf{Figure~\ref{fig:bmark}B}).
 % (See the \pkg{Peacock.test} package for \proglang{R} implementation of the Peacock test.)

### 2021-06-17 journalMS/article.tex
RB minor edits, good to go

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Fasano--Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in \proglang{R}}
-\Plaintitle{Fasano--Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in R}
-\Shorttitle{2-Dimensional Kolmogorov--Smirnov in \proglang{R}}
+\title{Fasano--Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov test in \proglang{R}}
+\Plaintitle{Fasano--Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov test in R}
+\Shorttitle{2-dimensional KS test in \proglang{R}}
 

### 2021-06-17 journalMS/article.tex
fixed spelling and grammar mistakes

@@
 the \pkg{fasano.franceschini.test} package presents a robust
-statistical test for analysing random samples defined in 2-dimensions.
+statistical test for analyzing random samples defined in 2-dimensions.
 }
@@
 Thus for any given $n$ points in a two-dimensional space, those $n$ points define $4n$ (rather than
-$4n^2$) quadrants.  The proceedure is illustrated in \textbf{Figure~\ref{fig:kstest2D}}.
+$4n^2$) quadrants.  The procedure is illustrated in \textbf{Figure~\ref{fig:kstest2D}}.
 The algorithm loops through each point in one sample in turn to define
@@
 ($n=10, 100, 1000, 5000$), with 10 replicates for each run. The
-Fasano--Franceschini test bootstrap precedure was further evaluated
+Fasano--Franceschini test bootstrap procedure was further evaluated
 under four different bootstrap iterations (no bootstrap, 10, 100, 1000),
@@
 \citep{Lopes2007}.  Our benchmarks also show this advantage, as shown
-in \textbf{Figure~\ref{fig:bmark}A}.  While the implentation of the
+in \textbf{Figure~\ref{fig:bmark}A}.  While the implementation of the
 bootstrapping procedure increases runtime in comparison to the

### 2021-06-17 R/fasano_franceschini_test.R
updated doi links

@@ -14,8 +14,8 @@
 #'
 #' @references{
 #' \itemize{
-#' \item{Fasano, G., Franceschini, A. (1987)<DOI:10.1093/mnras/225.1.155>. A multidimensional version of the Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society 225:155-170.}
-#' \item{Peacock J.A. (1983) <DOI:10.1093/mnras/202.3.615>. Two-dimensional goodness-of-fit testing in astronomy. Monthly Notices of the Royal Astronomical Society 202:615-627.}
+#' \item{Fasano, G., Franceschini, A. (1987) \doi{10.1093/mnras/225.1.155}. A multidimensional version of the Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society 225:155-170.}
+#' \item{Peacock J.A. (1983) \doi{10.1093/mnras/202.3.615}. Two-dimensional goodness-of-fit testing in astronomy. Monthly Notices of the Royal Astronomical Society 202:615-627.}
 #' \item{Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688}
 #' }
 #' }

### 2021-06-18 journalMS/article.R
removed journal MS from github page

@@ -1,356 +1,0 @@
-#' ## Figure Reproducibility Script
-#' ### Fasano--Franceschini Test: an Implementation of a 2-Dimensional Kolmogorov--Smirnov in R
-#' ### Authors: Elan Ness-Cohn and Rosemary Braun
-
-
-#+ setup, include=FALSE
-library(knitr)
-options(knitr.duplicate.label = "allow")
-opts_chunk$set(fig.width = 14, fig.height = 7)
-
-#' ## Package and options
-library(devtools)
-load_all()
-library("fasano.franceschini.test")
-library("parallel")
-library("ggplot2")
-library("viridis")
-library("mvtnorm")
-library("gridExtra")
-library("ggrepel")
-library("ggpubr")
-library("dplyr")
-
-# set to T to plot figures to PDF Output
-pdfOutPuts <- F
-
-#' ## fasano.franceschini.test Usage
-
-# set seed for reproducible example
-set.seed(123)
-
-# create 2-D samples with the same underlying distributions
-sample1Data <- data.frame(
-  x = rnorm(n = 100, mean = 0, sd = 1),
-  y = rnorm(n = 100, mean = 0, sd = 1)
-)
-sample2Data <- data.frame(
-  x = rnorm(n = 100, mean = 0, sd = 1),
-  y = rnorm(n = 100, mean = 0, sd = 1)
-)
-
-fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data)
-
-
-#' ## fasano.franceschini.test Bootstrap
-
-# set seed for reproducible example
-set.seed(123)
-
-# create 2-D samples with the same underlying distributions
-sample1Data <- data.frame(
-  x = rnorm(n = 100, mean = 0, sd = 1),
-  y = rnorm(n = 100, mean = 0, sd = 1)
-)
-sample2Data <- data.frame(
-  x = rnorm(n = 100, mean = 0, sd = 1),
-  y = rnorm(n = 100, mean = 0, sd = 1)
-)
-
-fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data, nBootstrap = 1000, cores = 1)
-
-
-#' ## Visualization - 2-D KS test
-
-if (pdfOutPuts) {
[…]

### 2021-06-18 journalMS/benchmarking.R

@@ -1,69 +1,0 @@
-# benchmarking Scripts
-# get arguments from command line
-args <- commandArgs(trailingOnly = TRUE)
-args <- as.numeric(args)
-print(args)
-
-# install.packages("Peacock.test")
-# devtools::install_github("eddelbuettel/rbenchmark")
-library(Peacock.test)
-library(devtools)
-load_all("./")
-library(fasano.franceschini.test)
-library(rbenchmark)
-
-sample1 <- data.frame(
-  x =  rnorm(args[1] / 2, mean = 0, sd = 1),
-  y =  rnorm(args[1] / 2, mean = 0, sd = 1)
-)
-
-sample2 <- data.frame(
-  x =  rnorm(args[1] / 2, mean = 0, sd = 1),
-  y =  rnorm(args[1] / 2, mean = 0, sd = 1)
-)
-
-if (args[2] == 0 & args[3] == 1) {
-  bench <- benchmark(
-    replications = 10,
-    "peacock" = {
-      peacock2(x = sample1, y = sample2)
-    },
-    "fasano" = {
-      fasano.franceschini.test(S1 = sample1, S2 = sample2, nBootstrap = args[2], cores = args[3])
-    }
-  )
-
-  out <- bench$elapsed
-
-  results <- data.frame(
-    N = args[1],
-    boots = args[2],
-    cores = args[3],
-    method = c("fasano", "peacock"),
-    time = out
-  )
-
-} else {
-
-  bench <- benchmark(
-    replications = 10,
-    "fasano" = {
-      fasano.franceschini.test(S1 = sample1, S2 = sample2, nBootstrap = args[2], cores = args[3])
-    }
-  )
-
-  out <- bench$elapsed
-
-  results <- data.frame(
-    N = args[1],
-    boots = args[2],
-    cores = args[3],
-    method = c("fasano"),
-    time = out
-  )
-}
-
-write.table(results,
-  file = paste0("~/Desktop/fasano.franceschini.test/journalMS/benchmarking/result_", args[1], "_", args[2], "_", args[3], ".txt"), sep = "\t",
-  row.names = FALSE, col.names = TRUE
-)

### 2021-06-18 journalMS/run_benchmark_local.sh

@@ -1,12 +1,0 @@
-#!/bin/bash
-while IFS=$'\t' read N boots cores
-do
-	Rscript benchmarking.R $N $boots $cores
-	# print out the job id for reference later
-	echo "JobID = ${JOB} for parameters $fileName submitted on `date`"
-done < parameters_benchmarking.txt
-exit
-
-# make this file executable and then run from the command line
-# chmod u+x run_limoRhyde_local.sh
-# ./run_limoRhyde_local.sh

### 2021-06-18 README.Rmd
fixed spelling in readme

@@
 
-# Fasano Franceschini Test
+# Fasano-Franceschini Test
 
@@
 
-The fasano.franceschini.test package is an R implementation of the 2-D Kolmogorov-Smirnov (**KS**) two-sample test as defined by Fasano and Franceschini (1987).
+The `fasano.franceschini.test` package is an R implementation of the 2-D Kolmogorov-Smirnov (**KS**) two-sample test as defined by Fasano and Franceschini (1987).
 This is a variant of the 2-D two-sample KS test as originally defined by Peacock (1983).
@@
 
-You can install the released version of fasano.franceschini.test from [CRAN](https://CRAN.R-project.org) with:
+<!-- You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://CRAN.R-project.org) with:
 
@@
 
-And the development version from [GitHub](https://github.com/) with:
+And the --> 
+You can install the development version of the `fasano.franceschini.test` package from [GitHub](https://github.com/) with:
 

### 2021-06-18 R/fasano_franceschini_test.R
set default parallel cores of ff test from 2 to 1

@@ -26,10 +26,10 @@
 #' set.seed(123)
 #'
 #' #create 2-D samples with different underlying distributions
-#' sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 3), y = rnorm(n = 100,mean = 0, sd = 1))
-#' sample2Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1), y = rnorm(n = 100,mean = 0, sd = 3))
+#' sample1Data <- data.frame(x = rnorm(n = 10,mean = 0, sd = 3), y = rnorm(n = 10,mean = 0, sd = 1))
+#' sample2Data <- data.frame(x = rnorm(n = 10,mean = 0, sd = 1), y = rnorm(n = 10,mean = 0, sd = 3))
 #'
-#' fasano.franceschini.test(sample1Data,sample2Data)
+#' fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data)
 #'
 #'
 #' #Underlying distributions are the same
@@ -37,15 +37,15 @@
 #' set.seed(123)
 #'
 #' #create 2-D samples with the same underlying distributions
-#' sample1Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1), y = rnorm(n = 100,mean = 0, sd = 1))
-#' sample2Data <- data.frame(x = rnorm(n = 100,mean = 0, sd = 1), y = rnorm(n = 100,mean = 0, sd = 1))
+#' sample1Data <- data.frame(x = rnorm(n = 10,mean = 0, sd = 1), y = rnorm(n = 10,mean = 0, sd = 1))
+#' sample2Data <- data.frame(x = rnorm(n = 10,mean = 0, sd = 1), y = rnorm(n = 10,mean = 0, sd = 1))
 #'
-#' fasano.franceschini.test(sample1Data,sample2Data)
+#' fasano.franceschini.test(S1 = sample1Data, S2 = sample2Data)
 #'
 #' @import methods
 #' @export
 
-fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
+fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 1) {
 
   #validate inputs
   if(!is.data.frame(S1) | !is.data.frame(S2)){
@@ -58,15 +58,12 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, cores = 2) {
     stop("nBootstrap must be a positive value")
[…]

### 2021-06-18 README.Rmd
updated readme

@@
 sample1Data <- data.frame(
-  x = rnorm(n = 100, mean = 0, sd = 3),
-  y = rnorm(n = 100, mean = 0, sd = 1)
+  x = rnorm(n = 50, mean = 0, sd = 3),
+  y = rnorm(n = 50, mean = 0, sd = 1)
 )
 sample2Data <- data.frame(
-  x = rnorm(n = 100, mean = 0, sd = 1),
-  y = rnorm(n = 100, mean = 0, sd = 3)
+  x = rnorm(n = 50, mean = 0, sd = 1),
+  y = rnorm(n = 50, mean = 0, sd = 3)
 )
@@
 sample1Data <- data.frame(
-  x = rnorm(n = 100, mean = 0, sd = 1),
-  y = rnorm(n = 100, mean = 0, sd = 1)
+  x = rnorm(n = 50, mean = 0, sd = 1),
+  y = rnorm(n = 50, mean = 0, sd = 1)
 )
 sample2Data <- data.frame(
-  x = rnorm(n = 100, mean = 0, sd = 1),
-  y = rnorm(n = 100, mean = 0, sd = 1)
+  x = rnorm(n = 50, mean = 0, sd = 1),
+  y = rnorm(n = 50, mean = 0, sd = 1)
 )

### 2021-06-19 README.Rmd
added ff test logo to website

@@
   comment = "#>",
-  fig.path = "man/figures/README-",
+  fig.path = "man/figures/",
   out.width = "100%"
 )
+library(badger)
 ```
+
+<img src="man/figures/logo.png" width="200" align="right"/>
 
@@
 <!-- badges: start -->
+
+<!-- `r badge_doi("10.1101/2020.11.19.389981", "yellow")` -->
+[![R build status](https://github.com/nesscoder/fasano.franceschini.test/workflows/R-CMD-check/badge.svg)](https://github.com/nesscoder/fasano.franceschini.test/actions)
+[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
+`r badge_devel("nesscoder/fasano.franceschini.test", "blue")`
+`r badge_code_size("nesscoder/fasano.franceschini.test")`
+
 <!-- badges: end -->

### 2021-06-22 README.Rmd
updated package for CRAN installation

--- 2021-06-18 README.Rmd
@@
 <!-- badges: start -->
+
+`badge_custom(x = "arXiv", y = "abs/2106.10539", color = "yellow", url = "https://arxiv.org/abs/2106.10539")` 
+[![R build status](https://github.com/nesscoder/fasano.franceschini.test/workflows/R-CMD-check/badge.svg)](https://github.com/nesscoder/fasano.franceschini.test/actions) 
+[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
+`r badge_devel("nesscoder/fasano.franceschini.test", "blue")` 
+`r badge_code_size("nesscoder/fasano.franceschini.test")`
+`r badge_cran_download("fasano.franceschini.test", "grand-total", "blue")`
+
 <!-- badges: end -->
@@
 
-<!-- You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://CRAN.R-project.org) with:
+You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://CRAN.R-project.org) with:
 
@@
 
-And the --> 
-You can install the development version of the `fasano.franceschini.test` package from [GitHub](https://github.com/) with:
+And the development version of the `fasano.franceschini.test` package from [GitHub](https://github.com/) with:
 
@@
 ```
+
 ## Example

### 2021-09-03 R/fasano_franceschini_test.R
dev version with defunc of nBootstrap for nPermute

@@ -6,7 +6,6 @@
 #'
 #' @param S1 a `[n by 2]` `data.frame` of x and y coordinates of sample 1
 #' @param S2 a `[n by 2]` `data.frame` of x and y coordinates of sample 2
-#' @param nBootstrap a depreciated `numeric` argument defining the number of bootstrapped samples to be generated for computing the empirical p-value. nBootstrap is set to be replaced by nPermute in the next released version of the package.
 #' @param nPermute a `numeric` defining the number of permuted samples to be generated for computing the empirical p-value (note this procedure is slow and computationally expensive on the order of nPermute*O(n^2). Default is set to 0.
 #' If nPermute is 0, the Fasano Franceschini distributional approximation is used for defining the p-value. See Fasano and Franceschini test (1987) for details.
 #' @param cores a `numeric` defining the number of cores to use for processing
@@ -46,7 +45,7 @@
 #' @import methods
 #' @export
 
-fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, nPermute = 0, cores = 1) {
+fasano.franceschini.test <- function(S1, S2, nPermute = 0, cores = 1) {
 
   #validate inputs
   if(!is.data.frame(S1) | !is.data.frame(S2)){
@@ -54,13 +53,6 @@ fasano.franceschini.test <- function(S1, S2, nBootstrap = 0, nPermute = 0, cores
   }
   if(ncol(S1) != 2 | ncol(S2) != 2){
     stop("S1 and S2 must be a `data.frame` of dim nrows x 2 cols, more than 2 cols detected")
-  }
-
-  if (!missing(nBootstrap)) {
-    warning("The 'nBootstrap' argument is deprecated, and will be removed in the next version of this package.
-             Use 'nPermute' instead of 'nBootstrap' to compute the empirical null distribution for your data.
-             The 'nPermute' argument has been set equal to 'nBootstrap'.")
-    nPermute <- nBootstrap
   }
 
   if(nPermute < 0){

### 2021-09-03 tests/testthat/test-fasano_franceschini_test.R

@@ -34,12 +34,11 @@ test_that("FF test computes with bootstrap, bootstrap is depreciated", {
 
   set.seed(123)
 
-  output <- fasano.franceschini.test(S1 = S1, S2 = S2, nBootstrap = 1000, cores = 1)
+  output <- fasano.franceschini.test(S1 = S1, S2 = S2, nPermute = 1000, cores = 1)
   expect_equal(unname(output$p.value), 0.001998002)
   expect_named(output$p.value, "p-value")
   expect_equal(unname(output$statistic), 0.525)
   expect_named(output$statistic, "D-stat")
-  expect_warning(fasano.franceschini.test(S1 = S1, S2 = S2, nBootstrap = 1000, cores = 1),"The 'nPermute' argument has been set equal to 'nBootstrap'.")
 })
 
 test_that("FF test computes with Permute", {

### 2022-05-17 README.Rmd
Version 2.0.0 initial commit

@@
 
-`badge_custom(x = "arXiv", y = "abs/2106.10539", color = "yellow", url = "https://arxiv.org/abs/2106.10539")` 
-[![R build status](https://github.com/nesscoder/fasano.franceschini.test/workflows/R-CMD-check/badge.svg)](https://github.com/nesscoder/fasano.franceschini.test/actions) 
-[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
+`r badge_custom(x = "arXiv", y = "abs/2106.10539", color = "yellow", url = "https://arxiv.org/abs/2106.10539")` 
+`r badge_cran_release("fasano.franceschini.test", "orange")`
 `r badge_devel("nesscoder/fasano.franceschini.test", "blue")` 
-`r badge_code_size("nesscoder/fasano.franceschini.test")`
-`r badge_cran_download("fasano.franceschini.test", "grand-total", "blue")`
-
+`r badge_cran_download("fasano.franceschini.test", "grand-total", "green")`
+[![R-CMD-check](https://github.com/cpuritz/fftest/workflows/R-CMD-check/badge.svg)](https://github.com/cpuritz/fftest/actions)
 <!-- badges: end -->
 
-The `fasano.franceschini.test` package is an R implementation of the 2-D Kolmogorov-Smirnov (**KS**) two-sample test as defined by Fasano and Franceschini (1987).
-This is a variant of the 2-D two-sample KS test as originally defined by Peacock (1983).
+The `fasano.franceschini.test` package is an R implementation of the multidimensional Kolmogorov-Smirnov two-sample test as defined by Fasano and Franceschini (1987).
+
+```
+Fasano, G. & Franceschini, A. (1987). A multidimensional version of the
+Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society,
+225:155-170. doi: 10.1093/mnras/225.1.155.
+```
 
@@
 
-## Example
+## Examples
 
-#### Underlying Distributions Are Different
+#### Underlying distributions are the same
 
@@
 
-#set seed for reproducible example
-set.seed(123)
+# set seed for reproducibility
+set.seed(0)
 
-#create 2-D samples with the different underlying distributions
-sample1Data <- data.frame(
-  x = rnorm(n = 50, mean = 0, sd = 3),
-  y = rnorm(n = 50, mean = 0, sd = 1)
-)
-sample2Data <- data.frame(
-  x = rnorm(n = 50, mean = 0, sd = 1),
-  y = rnorm(n = 50, mean = 0, sd = 3)
-)
+# create 2D samples from the same underlying distribution
+S1 <- data.frame(x = rnorm(n = 50, mean = 0, sd = 1),
+                 y = rnorm(n = 50, mean = 0, sd = 3))
+S2 <- data.frame(x = rnorm(n = 100, mean = 0, sd = 1),
+                 y = rnorm(n = 100, mean = 0, sd = 3))
 
-fasano.franceschini.test(sample1Data,sample2Data)
+fasano.franceschini.test(S1, S2, seed = 0)
 ```
 
-#### Underlying Distributions Are The Same
+#### Underlying distributions are different
 
 ```{r}
-#set seed for reproducible example
-set.seed(123)
+# set seed for reproducibility
+set.seed(1)
 
-#create 2-D samples with the same underlying distributions
-sample1Data <- data.frame(
-  x = rnorm(n = 50, mean = 0, sd = 1),
-  y = rnorm(n = 50, mean = 0, sd = 1)
-)
-sample2Data <- data.frame(
-  x = rnorm(n = 50, mean = 0, sd = 1),
-  y = rnorm(n = 50, mean = 0, sd = 1)
-)
+# create 3D samples from different underlying distributions
+S1 <- cbind(rgamma(n = 43, shape = 2),
+            rpois(n = 43, lambda = 5),
+            rpois(n = 43, lambda = 3.5))
+S2 <- cbind(rgamma(n = 72, shape = 2),
+            rpois(n = 72, lambda = 5),
+            rpois(n = 72, lambda = 5))
 
-fasano.franceschini.test(sample1Data,sample2Data)
+fasano.franceschini.test(S1, S2, seed = 1)
 ```

### 2022-05-17 v1/README.Rmd
Version 2.0.0 initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
library(badger)
```

<img src="man/figures/logo.png" width="200" align="right"/>

# Fasano-Franceschini Test

<!-- badges: start -->

`r badge_custom(x = "arXiv", y = "abs/2106.10539", color = "yellow", url = "https://arxiv.org/abs/2106.10539")` 
[![R build status](https://github.com/nesscoder/fasano.franceschini.test/workflows/R-CMD-check/badge.svg)](https://github.com/nesscoder/fasano.franceschini.test/actions) 
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
`r badge_devel("nesscoder/fasano.franceschini.test", "blue")` 
`r badge_code_size("nesscoder/fasano.franceschini.test")`
`r badge_cran_download("fasano.franceschini.test", "grand-total", "blue")`

<!-- badges: end -->

The `fasano.franceschini.test` package is an R implementation of the 2-D Kolmogorov-Smirnov (**KS**) two-sample test as defined by Fasano and Franceschini (1987). This is a variant of the 2-D two-sample KS test as originally defined by Peacock (1983).

## Installation

You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("fasano.franceschini.test")
```

And the development version of the `fasano.franceschini.test` package from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nesscoder/fasano.franceschini.test")
```

## Example

#### Underlying Distributions Are Different

```{r}
library(fasano.franceschini.test)

#set seed for reproducible example
set.seed(123)

#create 2-D samples with the different underlying distributions
sample1Data <- data.frame(
  x = rnorm(n = 50, mean = 0, sd = 3),
  y = rnorm(n = 50, mean = 0, sd = 1)
)
sample2Data <- data.frame(
  x = rnorm(n = 50, mean = 0, sd = 1),
  y = rnorm(n = 50, mean = 0, sd = 3)
)

fasano.franceschini.test(sample1Data,sample2Data)
```

#### Underlying Distributions Are The Same

```{r}
#set seed for reproducible example
set.seed(123)

#create 2-D samples with the same underlying distributions
sample1Data <- data.frame(
  x = rnorm(n = 50, mean = 0, sd = 1),
  y = rnorm(n = 50, mean = 0, sd = 1)
)
sample2Data <- data.frame(
  x = rnorm(n = 50, mean = 0, sd = 1),
  y = rnorm(n = 50, mean = 0, sd = 1)
)

fasano.franceschini.test(sample1Data,sample2Data)
```

### 2022-05-17 R/RcppExports.R

@@ -1,0 +1,19 @@
+# Generated by using Rcpp::compileAttributes() -> do not edit by hand
+# Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+ffTestStatistic <- function(S1, S2) {
+    .Call(`_fasano_franceschini_test_ffTestStatistic`, S1, S2)
+}
+
+permutationTestSeeded <- function(S1, S2, nPermute, verbose, seed) {
+    .Call(`_fasano_franceschini_test_permutationTestSeeded`, S1, S2, nPermute, verbose, seed)
+}
+
+permutationTest <- function(S1, S2, nPermute, verbose) {
+    .Call(`_fasano_franceschini_test_permutationTest`, S1, S2, nPermute, verbose)
+}
+
+permutationTestParallel <- function(S1, S2, nPermute) {
+    .Call(`_fasano_franceschini_test_permutationTestParallel`, S1, S2, nPermute)
+}
+

### 2022-05-17 R/ff.test-package.R

@@ -1,0 +1,7 @@
+## usethis namespace: start
+#' @importFrom Rcpp evalCpp
+#' @importFrom RcppParallel RcppParallelLibs
+#' @importFrom stats binom.test
+#' @useDynLib fasano.franceschini.test, .registration = TRUE
+## usethis namespace: end
+NULL

### 2022-05-17 R/ff_function.R

@@ -1,0 +1,159 @@
+#' Fasano-Franceschini Test
+#'
+#' Performs a multidimensional two-sample Kolmogorov-Smirnov test as described by Fasano and Franceschini (1987).
+#' This test evaluates the null hypothesis that two i.i.d. random samples were drawn from the same underlying
+#' probability distribution. The data can be of any dimension, and can be of any type (continuous, discrete, or mixed).
+#'
+#' @param S1 \code{matrix} or \code{data.frame}.
+#' @param S2 \code{matrix} or \code{data.frame}.
+#' @param nPermute a nonnegative \code{integer} setting the number of permuted samples to generate
+#' when estimating the permutation test p-value. Default is 100. If set to 0, no p-value is estimated.
+#' @param threads a positive \code{integer} or \code{"auto"} setting the number of threads used for performing
+#' the permutation test. If set to \code{"auto"}, the number of threads is determined by
+#' \code{RcppParallel::defaultNumThreads()}. Default is 1.
+#' @param verbose a \code{boolean} indicating whether to display a progress bar. Default is \code{FALSE}. Only available for serial
+#' version (\code{threads} = 1).
+#' @param seed optional integer to seed the PRNG used for the permutation test. Default is \code{NULL}. Only available for serial
+#' version (\code{threads} = 1).
+#' @param p.conf.level confidence level for the confidence interval of the permutation test p-value.
+#' @return A list with class \code{htest} containing the following components:
+#'   \item{statistic}{the value of the test statistic Z.}
+#'   \item{estimate}{the value of the sample test statistics D1 and D2.}
+#'   \item{p.value}{the permutation test p-value.}
+#'   \item{conf.int}{a binomial confidence interval for the p-value.}
+#'   \item{method}{a character string indicating what type of test was performed.}
+#'   \item{data.name}{a character string giving the names of the data.}
+#' @references{
+#' \itemize{
+#'   \item{Fasano, G. & Franceschini, A. (1987). A multidimensional version of the Kolmogorov-Smirnov test. \emph{Monthly Notices of the Royal Astronomical Society}, 225:155-170. \doi{10.1093/mnras/225.1.155}.}
[…]

### 2022-05-17 R/test_functions.R

@@ -1,0 +1,177 @@
+# Naive 2D count
+#
+# @param origin numeric vector
+# @param x numeric vector
+# @param y numeric vector
+# @return numeric vector
+count2D <- function(origin, x, y) {
+    xLess <- x < origin[[1]]
+    yLess <- y < origin[[2]]
+    xGreater <- x > origin[[1]]
+    yGreater <- y > origin[[2]]
+
+    nq1 <- sum(yGreater & xGreater)
+    nq2 <- sum(yGreater & xLess)
+    nq4 <- sum(yLess & xGreater)
+    nq3 <- sum(yLess & xLess)
+
+    return(c(nq1, nq2, nq3, nq4)/length(x))
+}
+
+# Naive 3D count
+#
+# @param origin numeric vector
+# @param x numeric vector
+# @param y numeric vector
+# @param z numeric vector
+# @return numeric vector
+count3D <- function(origin, x, y, z) {
+    xGreater <- x > origin[[1]]
+    yGreater <- y > origin[[2]]
+    zGreater <- z > origin[[3]]
+    xLess <- x < origin[[1]]
+    yLess <- y < origin[[2]]
+    zLess <- z < origin[[3]]
+
+    nq1 <- sum(xLess & yLess & zLess)
+    nq2 <- sum(xGreater & yLess & zLess)
+    nq3 <- sum(xLess & yGreater & zLess)
+    nq4 <- sum(xLess & yLess & zGreater)
+    nq5 <- sum(xGreater & yGreater & zLess)
+    nq6 <- sum(xGreater & yLess & zGreater)
+    nq7 <- sum(xLess & yGreater & zGreater)
+    nq8 <- sum(xGreater & yGreater & zGreater)
+
+    return(c(nq1, nq2, nq3, nq4, nq5, nq6, nq7, nq8)/length(x))
+}
+
+# Naive 4D count
+#
+# @param origin numeric vector
+# @param x numeric vector
+# @param y numeric vector
+# @param z numeric vector
+# @param t numeric vector
+# @return numeric vector
+count4D <- function(origin, x, y, z, t) {
+    xLess <- x < origin[[1]]
+    yLess <- y < origin[[2]]
+    zLess <- z < origin[[3]]
+    tLess <- t < origin[[4]]
+    xGreater <- x > origin[[1]]
[…]

### 2022-05-17 src/ProgressBar.h

@@ -1,0 +1,56 @@
+// Create a simple Rcpp progress bar
+
+#ifndef PROGRESS_H
+#define PROGRESS_H
+
+#include <Rcpp.h>
+
+class ProgressBar {
+private:
+    // Number times the progress bar will be incremented
+    const int nsteps;
+    // When printing, how many ticks should be printed
+    const int ticksPerStep;
+    // Whether to display output
+    const bool verbose;
+    // How many ticks have been printed so far
+    int ticks;
+    // Percentage of task completed so far
+    double progress;
+    // Number of ticks to print
+    static const int MAX_TICKS = 51;
+public:
+    ProgressBar(int nsteps, bool verbose) :
+        nsteps(nsteps), ticksPerStep(MAX_TICKS/nsteps+1), verbose(verbose), ticks(0), progress(0) {
+        if (verbose) {
+            Rcpp::Rcerr << "0%   10   20   30   40   50   60   70   80   90   100%\n" <<
+                           "|----|----|----|----|----|----|----|----|----|----|\n";
+        }
+    }
+
+    void step() {
+        if (verbose) {
+            progress += 1/static_cast<double>(nsteps);
+            if (progress > ticks/static_cast<double>(MAX_TICKS) && ticks < MAX_TICKS) {
+                for (int i = 0; i < ticksPerStep; ++i) {
+                    Rcpp::Rcerr << "*";
+                    ++ticks;
+                    if (ticks == MAX_TICKS) {
+                        break;
+                    }
+                }
+            }
+        }
+    }
+
+    // Double check all ticks have been printed at the end
+    void finalize() {
+        if (verbose) {
+            for (int i = ticks; i < MAX_TICKS; ++i) {
+                Rcpp::Rcerr << "*";
+            }
+        }
+    }
+};
+
+#endif //PROGRESS_H

### 2022-05-17 src/RangeTree.h

@@ -1,0 +1,1291 @@
+/*
+ * MIT License
+ *
+ * Copyright (c) 2016 Luca Weihs
+ *
+ * Permission is hereby granted, free of charge, to any person obtaining a copy
+ * of this software and associated documentation files (the "Software"), to deal
+ * in the Software without restriction, including without limitation the rights
+ * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+ * copies of the Software, and to permit persons to whom the Software is
+ * furnished to do so, subject to the following conditions:
+ *
+ * The above copyright notice and this permission notice shall be included in all
+ * copies or substantial portions of the Software.
+ *
+ * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+ * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+ * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+ * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+ * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+ * OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+ * SOFTWARE.
+ */
+
+/**
+* Implements the RangeTree datastructure.
+*
+* See the documentation of the class RangeTree for more details regarding the purpose
+* of RangeTree's.
+*
+* This file contains two main abstractions meant for use:
+* 1. The RangeTree class.
+* 2. A Point class which captures the idea of a d-dimensional euclidean point.
+*/
+
+#ifndef RANGETREE_H
+#define RANGETREE_H
+
+#include <vector>
+#include <iostream>
+#include <sstream>
+#include <numeric>
+#include <type_traits>
+#include <deque>
+#include <cmath>
+#include <algorithm>
+
+namespace RangeTree {
+
+    /**
+    * A point in euclidean space.
+    *
+    * A class that represents a multi-dimensional euclidean point
+    * with some associated value. We allow for each point to have an
+    * associated value so that some more information can be stored with
+    * each point. Points can also have a multiplicity/count, this corresponds
+    * to having several duplicates of the same point.
+    */
[…]

### 2022-05-17 src/RcppExports.cpp

@@ -1,0 +1,79 @@
+// Generated by using Rcpp::compileAttributes() -> do not edit by hand
+// Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+#include <Rcpp.h>
+
+using namespace Rcpp;
+
+#ifdef RCPP_USE_GLOBAL_ROSTREAM
+Rcpp::Rostream<true>&  Rcpp::Rcout = Rcpp::Rcpp_cout_get();
+Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
+#endif
+
+// ffTestStatistic
+NumericVector ffTestStatistic(const NumericMatrix S1, const NumericMatrix S2);
+RcppExport SEXP _fasano_franceschini_test_ffTestStatistic(SEXP S1SEXP, SEXP S2SEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< const NumericMatrix >::type S1(S1SEXP);
+    Rcpp::traits::input_parameter< const NumericMatrix >::type S2(S2SEXP);
+    rcpp_result_gen = Rcpp::wrap(ffTestStatistic(S1, S2));
+    return rcpp_result_gen;
+END_RCPP
+}
+// permutationTestSeeded
+unsigned int permutationTestSeeded(const NumericMatrix S1, const NumericMatrix S2, int nPermute, bool verbose, int seed);
+RcppExport SEXP _fasano_franceschini_test_permutationTestSeeded(SEXP S1SEXP, SEXP S2SEXP, SEXP nPermuteSEXP, SEXP verboseSEXP, SEXP seedSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< const NumericMatrix >::type S1(S1SEXP);
+    Rcpp::traits::input_parameter< const NumericMatrix >::type S2(S2SEXP);
+    Rcpp::traits::input_parameter< int >::type nPermute(nPermuteSEXP);
+    Rcpp::traits::input_parameter< bool >::type verbose(verboseSEXP);
+    Rcpp::traits::input_parameter< int >::type seed(seedSEXP);
+    rcpp_result_gen = Rcpp::wrap(permutationTestSeeded(S1, S2, nPermute, verbose, seed));
[…]

### 2022-05-17 src/ff_internal_functions.cpp

@@ -1,0 +1,204 @@
+#include <Rcpp.h>
+#include <RcppParallel.h>
+#include <vector>
+#include <random>
+#include <cmath>
+#include "RangeTree.h"
+#include "matrix_util.h"
+#include "permute.h"
+#include "range_count.h"
+#include "ProgressBar.h"
+
+using namespace Rcpp;
+
+// Compute FF test statistics
+//
+// @param M a matrix of type NumericMatrix or RMatrix<double>
+// @param r1 number of rows in first sample
+// @param r2 number of rows in second sample
+// @param shuffle whether to shuffle the samples
+// @param prng a pseudorandom number generator (unused if shuffle == false)
+// @return FF test statistics
+template<typename MatrixT>
+std::vector<double> testStatistics(const MatrixT M,
+                                   std::size_t r1,
+                                   std::size_t r2,
+                                   bool shuffle,
+                                   std::mt19937& prng) {
+    double n1 = static_cast<double>(r1);
+    double n2 = static_cast<double>(r2);
+
+    // Build range trees
+    std::vector<std::size_t> s;
+    if (shuffle) {
+        s = permutation(r1 + r2, prng);
+    } else {
+        s = seq(r1 + r2);
+    }
+    std::vector<RTree> trees = buildRangeTrees<MatrixT>(M, r1, r2, s);
+
+    // Compute test statistic using first sample as origins
+    double d1 = -1;
+    for (std::size_t i = 0; i < r1; ++i) {
+        std::vector<double> org = getRow<MatrixT>(M, s[i]);
+        std::vector<double> c1 = rangeCount(trees[0], org);
+        std::vector<double> c2 = rangeCount(trees[1], org);
+        for (std::size_t j = 0; j < c1.size(); ++j) {
+            d1 = std::max(d1, abs(c1[j]/n1 - c2[j]/n2));
+        }
+    }
+
+    // Compute test statistic using second sample as origins
+    double d2 = -1;
+    for (std::size_t i = 0; i < r2; ++i) {
+        std::vector<double> org = getRow<MatrixT>(M, s[i+r1]);
+        std::vector<double> c1 = rangeCount(trees[0], org);
+        std::vector<double> c2 = rangeCount(trees[1], org);
+        for (std::size_t j = 0; j < c1.size(); ++j) {
+            d2 = std::max(d2, abs(c1[j]/n1 - c2[j]/n2));
+        }
+    }
+
[…]

### 2022-05-17 src/matrix_util.h

@@ -1,0 +1,46 @@
+// Utility functions for dealing with Rcpp matrices
+
+#ifndef MATRIXUTIL_H
+#define MATRIXUTIL_H
+
+#include <Rcpp.h>
+#include <vector>
+
+using namespace Rcpp;
+
+// Row bind two matrices
+//
+// @param M1 first matrix with dimensions r1 x c
+// @param M2 second matrix with dimensions r2 x c
+// @return the input matrices row-binded
+template<typename MatrixT>
+MatrixT rbind(const MatrixT M1, const MatrixT M2) {
+    std::size_t r1 = M1.nrow();
+    std::size_t r2 = M2.nrow();
+
+    MatrixT S(r1 + r2, M1.ncol());
+    for (std::size_t i = 0; i < r1; ++i) {
+        S(i,_) = M1(i,_);
+    }
+    for (std::size_t i = 0; i < r2; ++i) {
+        S(i+r1,_) = M2(i,_);
+    }
+    return S;
+}
+
+// Extract a row of a numeric matrix as a std::vector
+//
+// @param M input matrix
+// @param row index
+// @return the specified row represented as a std::vector
+template<typename MatrixT>
+std::vector<double> getRow(MatrixT M, std::size_t row) {
+    std::size_t ncol = M.ncol();
+    std::vector<double> v(ncol);
+    for (std::size_t i = 0; i < ncol; ++i) {
+        v[i] = M(row, i);
+    }
+    return v;
+}
+
+#endif //MATRIXUTIL_H

### 2022-05-17 src/permute.h

@@ -1,0 +1,31 @@
+// Utility functions for performing a permutation test
+
+#ifndef PERMUTE_H
+#define PERMUTE_H
+
+#include <vector>
+#include <random>
+#include <algorithm>
+
+// Return the sequence 0,1,...,n-1
+// @param n length of the sequence
+// @return the vector 0,1,...,n-1
+std::vector<std::size_t> seq(std::size_t n) {
+    std::vector<std::size_t> ix(n);
+    std::iota(ix.begin(), ix.end(), 0);
+    return ix;
+}
+
+// Create a permutation of the indices 0,1,...,n-1
+//
+// @param n length of permutation
+// @param prng a pseudorandom number generator
+// @return a permutation of {0,1,...,n-1}
+std::vector<std::size_t> permutation(std::size_t n,
+                                     std::mt19937& prng) {
+    std::vector<std::size_t> ix = seq(n);
+    std::shuffle(ix.begin(), ix.end(), prng);
+    return ix;
+}
+
+#endif //PERMUTE_H

### 2022-05-17 src/range_count.h

@@ -1,0 +1,78 @@
+// Functions to perform range counting
+
+#ifndef RANGECOUNT_H
+#define RANGECOUNT_H
+
+#include <Rcpp.h>
+#include <vector>
+#include <limits>
+#include "RangeTree.h"
+#include "matrix_util.h"
+
+using namespace Rcpp;
+
+typedef RangeTree::RangeTree<double,int> RTree;
+
+// Build range trees for two samples
+//
+// @param M matrix of two data samples row binded
+// @param r1 number of points in first sample
+// @param r2 number of points in second sample
+// @param s a permutation of {0,1,...,(r1+r2-1)} which indicates which points belong to each sample
+// @return range trees built on each sample
+template<typename MatrixT>
+std::vector<RTree> buildRangeTrees(const MatrixT M,
+                                   std::size_t r1,
+                                   std::size_t r2,
+                                   const std::vector<std::size_t>& s) {
+    std::vector<RangeTree::Point<double,int> > pts1(r1);
+    std::vector<RangeTree::Point<double,int> > pts2(r2);
+
+    // Extract points from first sample
+    for (std::size_t i = 0; i < r1; ++i) {
+        pts1[i] = RangeTree::Point<double,int>(getRow<MatrixT>(M, s[i]), 0);
+    }
+
+    // Extract points from second sample
+    for (std::size_t i = 0; i < r2; ++i) {
+        pts2[i] = RangeTree::Point<double,int>(getRow<MatrixT>(M, s[i+r1]), 0);
+    }
+
+    // Construct range trees
+    return {RTree(pts1), RTree(pts2)};
+}
+
+// Count the number of points in each axis-aligned orthant defined by the origin
+//
+// @param rtree range tree
+// @param origin vector
+// @return vector counting points in each orthant
+std::vector<double> rangeCount(const RTree& rtree,
+                               const std::vector<double>& origin) {
+    std::size_t ndim = origin.size();
+    std::size_t noct = 1<<ndim;
+    std::vector<double> res(noct);
+    std::vector<bool> strict(ndim, false);
+    double inf = std::numeric_limits<double>::infinity();
+
+    // Count the number of points in each axis-aligned orthant using specified origin
+    for (std::size_t i = 0; i < noct; ++i) {
+        // Determine the upper and lower bounds on each dimension for the given orthant.
+        // That is, the orthant can be represented as the Cartesian product of the intervals
[…]

### 2022-05-17 tests/testthat/test_2d.R

@@ -1,0 +1,33 @@
+test_that("compare with naive 2d implementation", {
+    set.seed(2)
+
+    ffcomp <- function(S1, S2) {
+        res <- fasano.franceschini.test(S1, S2, nPermute = 0)
+        d12 <- c(getDstat_naive(S1, S1, S2), getDstat_naive(S2, S1, S2))
+        names(d12) <- names(res$estimate)
+
+        n1 <- as.numeric(dim(S1)[1])
+        n2 <- as.numeric(dim(S2)[1])
+        z <- sqrt(n1*n2/(n1+n2))*((d12[1] + d12[2])/2.0)
+        names(z) <- names(res$statistic)
+
+        expect_equal(d12, res$estimate, tolerance = 1e-14)
+        expect_equal(z, res$statistic, tolerance = 1e-14)
+    }
+
+    S1 <- data.frame(rnorm(n=50, mean=0, sd=1), rnorm(n=50, mean=1, sd=1))
+    S2 <- data.frame(rnorm(n=83, mean=0.2, sd=1), rnorm(n=83, mean=0.1, sd=2))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rnorm(n=132, mean=10, sd=1), rnorm(n=132, mean=10, sd=1))
+    S2 <- data.frame(rnorm(n=47, mean=10, sd=1), rnorm(n=47, mean=10, sd=1))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n=100, shape1=1.3, shape2=2.73), rbeta(n=100, shape1=1.3, shape2=2.73))
+    S2 <- data.frame(rbeta(n=23, shape1=1.3, shape2=2.73), rbeta(n=23, shape1=1.3, shape2=2.73))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n=224, shape1=10.1, shape2=0.1), rbeta(n=224, shape1=10.1, shape2=0.1))
+    S2 <- data.frame(rbeta(n=1557, shape1=10.1, shape2=0.1), rbeta(n=1557, shape1=10.1, shape2=0.1))
+    ffcomp(S1, S2)
+})

### 2022-05-17 tests/testthat/test_3d.R

@@ -1,0 +1,49 @@
+test_that("compare with naive 3d implementation", {
+    set.seed(3)
+
+    ffcomp <- function(S1, S2) {
+        res <- fasano.franceschini.test(S1, S2, nPermute = 0)
+        d12 <- c(getDstat_naive(S1, S1, S2), getDstat_naive(S2, S1, S2))
+        names(d12) <- names(res$estimate)
+
+        n1 <- as.numeric(dim(S1)[1])
+        n2 <- as.numeric(dim(S2)[1])
+        z <- sqrt(n1*n2/(n1+n2))*((d12[1] + d12[2])/2.0)
+        names(z) <- names(res$statistic)
+
+        expect_equal(d12, res$estimate, tolerance = 1e-14)
+        expect_equal(z, res$statistic, tolerance = 1e-14)
+    }
+
+    S1 <- data.frame(rnorm(n=33, mean=1.4, sd=0.3),
+                     rnorm(n=33, mean=1.5, sd=1),
+                     rnorm(n=33, mean=1.6, sd=1.8))
+    S2 <- data.frame(rnorm(n=206, mean=1.4, sd=0.3),
+                     rnorm(n=206, mean=1.5, sd=1),
+                     rnorm(n=206, mean=1.6, sd=1.8))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n=100, shape1=1.3, shape2=2.73),
+                     rbeta(n=100, shape1=1.2, shape2=0.11),
+                     rbeta(n=100, shape1=2.1, shape2=1.39))
+    S2 <- data.frame(rbeta(n=65, shape1=1.3, shape2=2.73),
+                     rbeta(n=65, shape1=1.2, shape2=0.11),
+                     rbeta(n=65, shape1=2.1, shape2=1.39))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n=400, shape1=1.3, shape2=2.73),
+                     rbeta(n=400, shape1=1.2, shape2=0.11),
+                     rbeta(n=400, shape1=2.1, shape2=1.39))
+    S2 <- data.frame(rbeta(n=761, shape1=1.3, shape2=2.73),
+                     rbeta(n=761, shape1=1.2, shape2=0.11),
+                     rbeta(n=761, shape1=0.2, shape2=1.39))
+    ffcomp(S1, S2)
+
[…]

### 2022-05-17 tests/testthat/test_4d.R

@@ -1,0 +1,47 @@
+test_that("compare with naive 4d implementation", {
+    set.seed(4)
+
+    ffcomp <- function(S1, S2) {
+        res <- fasano.franceschini.test(S1, S2, nPermute = 0)
+        d12 <- c(getDstat_naive(S1, S1, S2), getDstat_naive(S2, S1, S2))
+        names(d12) <- names(res$estimate)
+
+        n1 <- as.numeric(dim(S1)[1])
+        n2 <- as.numeric(dim(S2)[1])
+        z <- sqrt(n1*n2/(n1+n2))*((d12[1] + d12[2])/2.0)
+        names(z) <- names(res$statistic)
+
+        expect_equal(d12, res$estimate, tolerance = 1e-14)
+        expect_equal(z, res$statistic, tolerance = 1e-14)
+    }
+
+    S1 <- data.frame(rnorm(n=70, mean=1.4, sd=0.3),
+                     rnorm(n=70, mean=1.5, sd=1),
+                     rnorm(n=70, mean=1.6, sd=1.8),
+                     rnorm(n=70, mean=1.6, sd=1.8))
+    S2 <- data.frame(rnorm(n=206, mean=1.4, sd=0.3),
+                     rnorm(n=206, mean=1.5, sd=1),
+                     rnorm(n=206, mean=1.6, sd=1.8),
+                     rnorm(n=206, mean=2.2, sd=0.01))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n=102, shape1=1.3, shape2=2.73),
+                     rgamma(n=102, shape=4.4, rate=0.4),
+                     rbeta(n=102, shape1=1.2, shape2=0.11),
+                     rbeta(n=102, shape1=2.1, shape2=1.39))
+    S2 <- data.frame(rbeta(n=65, shape1=1.3, shape2=2.73),
+                     rbeta(n=65, shape1=1.2, shape2=0.11),
+                     rbeta(n=65, shape1=2.1, shape2=1.39),
+                     rexp(n=65, rate=33.3))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n=400, shape1=1.3, shape2=2.73),
+                     rbeta(n=400, shape1=1.2, shape2=0.11),
+                     rbeta(n=400, shape1=2.1, shape2=1.39),
+                     rgamma(n=400, shape=0.9, rate=3.01))
[…]

### 2022-05-17 tests/testthat/test_5d.R

@@ -1,0 +1,53 @@
+test_that("compare with naive 5d implementation", {
+    set.seed(5)
+
+    ffcomp <- function(S1, S2) {
+        res <- fasano.franceschini.test(S1, S2, nPermute = 0)
+        d12 <- c(getDstat_naive(S1, S1, S2), getDstat_naive(S2, S1, S2))
+        names(d12) <- names(res$estimate)
+
+        n1 <- as.numeric(dim(S1)[1])
+        n2 <- as.numeric(dim(S2)[1])
+        z <- sqrt(n1*n2/(n1+n2))*((d12[1] + d12[2])/2.0)
+        names(z) <- names(res$statistic)
+
+        expect_equal(d12, res$estimate, tolerance = 1e-14)
+        expect_equal(z, res$statistic, tolerance = 1e-14)
+    }
+
+    S1 <- data.frame(rnorm(n=70, mean=1.4, sd=0.3),
+                     rnorm(n=70, mean=1.5, sd=1),
+                     rnorm(n=70, mean=1.6, sd=1.8),
+                     rnorm(n=70, mean=1.6, sd=1.8),
+                     rnorm(n=70, mean=1.6, sd=1.8))
+    S2 <- data.frame(rnorm(n=206, mean=1.4, sd=0.3),
+                     rnorm(n=206, mean=1.5, sd=1),
+                     rnorm(n=206, mean=1.6, sd=1.8),
+                     rnorm(n=206, mean=2.2, sd=0.01),
+                     rnorm(n=206, mean=2.2, sd=0.01))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n=102, shape1=1.3, shape2=2.73),
+                     rgamma(n=102, shape=4.4, rate=0.4),
+                     rbeta(n=102, shape1=1.2, shape2=0.11),
+                     rbeta(n=102, shape1=2.1, shape2=1.39),
+                     rbeta(n=102, shape1=2.1, shape2=1.39))
+    S2 <- data.frame(rbeta(n=65, shape1=1.3, shape2=2.73),
+                     rbeta(n=65, shape1=1.2, shape2=0.11),
+                     rbeta(n=65, shape1=2.1, shape2=1.39),
+                     rexp(n=65, rate=33.3),
+                     rexp(n=65, rate=33.3))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n=400, shape1=1.3, shape2=2.73),
[…]

### 2022-05-17 tests/testthat/test_errors.R

@@ -1,0 +1,34 @@
+test_that("test for bad inputs", {
+    set.seed(1)
+
+    S1 <- "a"
+    S2 <- data.frame(rnorm(n=50))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0))
+    expect_error(fasano.franceschini.test(S2, S1, nPermute = 0))
+
+    S1 <- data.frame(rnorm(n=61))
+    S2 <- data.frame(rnorm(n=72), rnorm(n=72))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0))
+    expect_error(fasano.franceschini.test(S2, S1, nPermute = 0))
+
+    S1 <- data.frame(rnorm(n=61), rnorm(n=61))
+    S2 <- data.frame(rnorm(n=72), rnorm(n=72))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = -1))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0.1))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 1.2))
+
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, cores = -1))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, cores = 0))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, cores = 0.1))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, cores = 1.1))
+
+    S1 <- rnorm(n=50)
+    S2 <- data.frame(rnorm(n=50))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0))
+    expect_error(fasano.franceschini.test(S2, S1, nPermute = 0))
+
+    S1 <- rnorm(n=60)
+    S2 <- as.matrix(rnorm(60))
+    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0))
+    expect_error(fasano.franceschini.test(S2, S1, nPermute = 0))
+})

### 2022-05-17 tests/testthat/test_seed.R

@@ -1,0 +1,30 @@
+test_that("test that seed works for serial version", {
+    set.seed(0)
+
+    S1 <- data.frame(rnorm(n=50, mean=0.2, sd=2), rnorm(n=50, mean=1.3, sd=1.1))
+    S2 <- data.frame(rnorm(n=83, mean=0.2, sd=2), rnorm(n=83, mean=1.3, sd=1.1))
+    test1 <- fasano.franceschini.test(S1, S2, nPermute = 100, seed = 0)
+    test2 <- fasano.franceschini.test(S1, S2, nPermute = 100, seed = 0)
+    expect_equal(test1$statistic, test2$statistic, tolerance = 1e-14)
+    expect_equal(test1$estimate, test2$estimate, tolerance = 1e-14)
+    expect_equal(test1$p.value, test2$p.value, tolerance = 1e-14)
+
+    S1 <- data.frame(rnorm(n=33, mean=1.4, sd=0.3),
+                     rnorm(n=33, mean=1.5, sd=1),
+                     rnorm(n=33, mean=1.6, sd=1.8))
+    S2 <- data.frame(rnorm(n=206, mean=1.4, sd=0.3),
+                     rnorm(n=206, mean=1.5, sd=1),
+                     rnorm(n=206, mean=1.6, sd=1.8))
+
+    test1 <- fasano.franceschini.test(S1, S2, nPermute = 20, seed = 1)
+    test2 <- fasano.franceschini.test(S1, S2, nPermute = 20, seed = 1)
+    expect_equal(test1$statistic, test2$statistic, tolerance = 1e-14)
+    expect_equal(test1$estimate, test2$estimate, tolerance = 1e-14)
+    expect_equal(test1$p.value, test2$p.value, tolerance = 1e-14)
+
+    test1 <- fasano.franceschini.test(S1, S2, nPermute = 31, seed = 2)
+    test2 <- fasano.franceschini.test(S1, S2, nPermute = 31, seed = 2)
+    expect_equal(test1$statistic, test2$statistic, tolerance = 1e-14)
+    expect_equal(test1$estimate, test2$estimate, tolerance = 1e-14)
+    expect_equal(test1$p.value, test2$p.value, tolerance = 1e-14)
+})

### 2022-05-17 v1/R/fasano_franceschini_test.R

@@ -1,0 +1,123 @@
+#' Fasano Franceschini Test
+#'
+#' Computes the 2-D Kolmogorov-Smirnov two-sample test as described by Fasano and Franceschini (1987).
+#'
+#' Code adapted from Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688
+#'
+#' @param S1 a `[n by 2]` `data.frame` of x and y coordinates of sample 1
+#' @param S2 a `[n by 2]` `data.frame` of x and y coordinates of sample 2
+#' @param nPermute a `numeric` defining the number of permuted samples to be generated for computing the empirical p-value (note this procedure is slow and computationally expensive on the order of nPermute*O(n^2). Default is set to 0.
+#' If nPermute is 0, the Fasano Franceschini distributional approximation is used for defining the p-value. See Fasano and Franceschini test (1987) for details.
+#' @param cores a `numeric` defining the number of cores to use for processing
+#'
+#' @return the 2-D ks statistic and p-value
+#'
+#' @references{
+#' \itemize{
+#' \item{Fasano, G., Franceschini, A. (1987) \doi{10.1093/mnras/225.1.155}. A multidimensional version of the Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society 225:155-170.}
+#' \item{Peacock J.A. (1983) \doi{10.1093/mnras/202.3.615}. Two-dimensional goodness-of-fit testing in astronomy. Monthly Notices of the Royal Astronomical Society 202:615-627.}
+#' \item{Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688}
+#' }
+#' }
+#'
+#' @examples
+#' #Underlying distributions are different
+#' #set seed for reproducible example
+#' set.seed(123)
+#'
+#' #create 2-D samples with different underlying distributions
+#' sample1Data <- data.frame(x = rnorm(n = 50, mean = 0, sd = 3), y = rnorm(n = 50,mean = 0, sd = 1))
[…]

### 2022-05-17 v1/R/interal_functions_FasFran.R

@@ -1,0 +1,130 @@
+#' Quad Count
+#'
+#' Counts the frequency of points in the four quadrants - starting from the upper right going counter clockwise. Quadrants defined by the origin points x and y.
+#'
+#' Code adapted from Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688
+#'
+#'
+#' @param x_origin a `numeric` defining the x coordinate of the origin defining the 4 quadrants
+#' @param y_origin a `numeric` defining the y coordinate of the origin defining the 4 quadrants
+#' @param x a `vector` of `numeric` x coordinates
+#' @param y a `vector` of `numeric` y coordinates
+#'
+#' @return a `vector` of frequencies of the number of points in each of the four quadrants defined by the origin point
+#'
+#'
+#'
+quadCount <- function(x_origin, y_origin, x, y) {
+
+  #number of points being checked
+  n <- length(x)
+
+  # sum number of points in each quadrant (counter clockwise from 1 to 4)
+  # sign(x - x_origin)
+  # sign(y - y_origin)
+  yGreater <- y > y_origin
+  xGreater <- x > x_origin
+  yLess <- y < y_origin
+  xLess <- x < x_origin
+  xequals <- x == x_origin
+  yequals <- y == y_origin
+
+  nq1 <- sum(yGreater & xGreater)
+  nq2 <- sum(yGreater & xLess)
+  nq4 <- sum(yLess & xGreater)
+  nq3 <- sum(yLess & xLess)
+
+  #origin point is divided equally across the four quadrants
+  #as well as any other point they may be tied with the origin
+  origin <- 0.25*sum(xequals & yequals)
+  nq1 <- nq1 + origin
+  nq2 <- nq2 + origin
+  nq3 <- nq3 + origin
+  nq4 <- nq4 + origin
+
+  #if there are any ties that are NOT the origin
+  #i.e. a point is exactly on the line between two quadrants
+  #split the point count equally between the quadrants
+  if(any(xor(xequals, yequals))){
+    #add warning message
+    yties_12 <- sum(yGreater & xequals)
[…]

### 2022-05-17 v1/tests/testthat/test-fasano_franceschini_test.R

@@ -1,0 +1,90 @@
+test_that("FF test uses validated inputs", {
+  S1 <- data.frame(
+    x = c(1, 1),
+    y = c(1, -1)
+  )
+  S2 <- data.frame(
+    x = c(0, 1),
+    y = c(0, -1)
+  )
+
+  expect_error(fasano.franceschini.test(S1 = NULL, S2 = S2, cores = 1), "S1 and S2 must be a `data.frame`")
+  expect_error(fasano.franceschini.test(S1 = NULL, S2 = NULL, cores = 1), "S1 and S2 must be a `data.frame`")
+  expect_error(fasano.franceschini.test(S1 = S1, S2 = NULL, cores = 1), "S1 and S2 must be a `data.frame`")
+  expect_error(fasano.franceschini.test(S1 = as.matrix(S1), S2 = S2, cores = 1), "S1 and S2 must be a `data.frame`")
+  expect_error(fasano.franceschini.test(S1 = S1, S2 = as.matrix(S2), cores = 1), "S1 and S2 must be a `data.frame`")
+  expect_error(fasano.franceschini.test(S1 = as.matrix(S1), S2 = as.matrix(S2), cores = 1), "S1 and S2 must be a `data.frame`")
+  expect_error(fasano.franceschini.test(S1 = S1, S2 = S2, nPermute = -1, cores = 1), "nPermute must be a positive value")
+})
+
+
+
+test_that("FF test computes with bootstrap, bootstrap is depreciated", {
+
+  # sample 1 - 30 points in quad 1
+  S1 <- data.frame(
+    x = c(rep(1, 30), rep(-1, 4), rep(-1, 3), rep(1, 2), 0),
+    y = c(rep(1, 30), rep(1, 4), rep(-1, 3), rep(-1, 2), 0)
+  )
+  # sample 2 - 1 points in quad 1
+  S2 <- data.frame(
+    x = c(1, rep(-1, 4), rep(-1, 3), rep(1, 2)),
+    y = c(1, rep(1, 4), rep(-1, 3), rep(-1, 2))
+  )
+
+  set.seed(123)
+
+  output <- fasano.franceschini.test(S1 = S1, S2 = S2, nPermute = 1000, cores = 1)
+  expect_equal(unname(output$p.value), 0.001998002)
+  expect_named(output$p.value, "p-value")
+  expect_equal(unname(output$statistic), 0.525)
[…]

### 2022-05-17 v1/R/fasano_franceschini_test.R
Moved v1 to local storage

@@ -1,123 +1,0 @@
-#' Fasano Franceschini Test
-#'
-#' Computes the 2-D Kolmogorov-Smirnov two-sample test as described by Fasano and Franceschini (1987).
-#'
-#' Code adapted from Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688
-#'
-#' @param S1 a `[n by 2]` `data.frame` of x and y coordinates of sample 1
-#' @param S2 a `[n by 2]` `data.frame` of x and y coordinates of sample 2
-#' @param nPermute a `numeric` defining the number of permuted samples to be generated for computing the empirical p-value (note this procedure is slow and computationally expensive on the order of nPermute*O(n^2). Default is set to 0.
-#' If nPermute is 0, the Fasano Franceschini distributional approximation is used for defining the p-value. See Fasano and Franceschini test (1987) for details.
-#' @param cores a `numeric` defining the number of cores to use for processing
-#'
-#' @return the 2-D ks statistic and p-value
-#'
-#' @references{
-#' \itemize{
-#' \item{Fasano, G., Franceschini, A. (1987) \doi{10.1093/mnras/225.1.155}. A multidimensional version of the Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society 225:155-170.}
-#' \item{Peacock J.A. (1983) \doi{10.1093/mnras/202.3.615}. Two-dimensional goodness-of-fit testing in astronomy. Monthly Notices of the Royal Astronomical Society 202:615-627.}
-#' \item{Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688}
-#' }
-#' }
-#'
-#' @examples
-#' #Underlying distributions are different
-#' #set seed for reproducible example
-#' set.seed(123)
-#'
-#' #create 2-D samples with different underlying distributions
-#' sample1Data <- data.frame(x = rnorm(n = 50, mean = 0, sd = 3), y = rnorm(n = 50,mean = 0, sd = 1))
[…]

### 2022-05-17 v1/R/interal_functions_FasFran.R

@@ -1,130 +1,0 @@
-#' Quad Count
-#'
-#' Counts the frequency of points in the four quadrants - starting from the upper right going counter clockwise. Quadrants defined by the origin points x and y.
-#'
-#' Code adapted from Press, W. H., Teukolsky, S. A., Vetterling, W. T.,, Flannery, B. P. (2007). Numerical Recipes 3rd Edition: The Art of Scientific Computing. Cambridge University Press. ISBN: 0521880688
-#'
-#'
-#' @param x_origin a `numeric` defining the x coordinate of the origin defining the 4 quadrants
-#' @param y_origin a `numeric` defining the y coordinate of the origin defining the 4 quadrants
-#' @param x a `vector` of `numeric` x coordinates
-#' @param y a `vector` of `numeric` y coordinates
-#'
-#' @return a `vector` of frequencies of the number of points in each of the four quadrants defined by the origin point
-#'
-#'
-#'
-quadCount <- function(x_origin, y_origin, x, y) {
-
-  #number of points being checked
-  n <- length(x)
-
-  # sum number of points in each quadrant (counter clockwise from 1 to 4)
-  # sign(x - x_origin)
-  # sign(y - y_origin)
-  yGreater <- y > y_origin
-  xGreater <- x > x_origin
-  yLess <- y < y_origin
-  xLess <- x < x_origin
-  xequals <- x == x_origin
-  yequals <- y == y_origin
-
-  nq1 <- sum(yGreater & xGreater)
-  nq2 <- sum(yGreater & xLess)
-  nq4 <- sum(yLess & xGreater)
-  nq3 <- sum(yLess & xLess)
-
-  #origin point is divided equally across the four quadrants
-  #as well as any other point they may be tied with the origin
-  origin <- 0.25*sum(xequals & yequals)
-  nq1 <- nq1 + origin
-  nq2 <- nq2 + origin
-  nq3 <- nq3 + origin
-  nq4 <- nq4 + origin
-
-  #if there are any ties that are NOT the origin
-  #i.e. a point is exactly on the line between two quadrants
-  #split the point count equally between the quadrants
-  if(any(xor(xequals, yequals))){
-    #add warning message
-    yties_12 <- sum(yGreater & xequals)
[…]

### 2022-05-17 v1/tests/testthat.R

@@ -1,4 +1,0 @@
-library(testthat)
-library(fasano.franceschini.test)
-
-test_check("fasano.franceschini.test")

### 2022-05-17 v1/tests/testthat/test-fasano_franceschini_test.R

@@ -1,90 +1,0 @@
-test_that("FF test uses validated inputs", {
-  S1 <- data.frame(
-    x = c(1, 1),
-    y = c(1, -1)
-  )
-  S2 <- data.frame(
-    x = c(0, 1),
-    y = c(0, -1)
-  )
-
-  expect_error(fasano.franceschini.test(S1 = NULL, S2 = S2, cores = 1), "S1 and S2 must be a `data.frame`")
-  expect_error(fasano.franceschini.test(S1 = NULL, S2 = NULL, cores = 1), "S1 and S2 must be a `data.frame`")
-  expect_error(fasano.franceschini.test(S1 = S1, S2 = NULL, cores = 1), "S1 and S2 must be a `data.frame`")
-  expect_error(fasano.franceschini.test(S1 = as.matrix(S1), S2 = S2, cores = 1), "S1 and S2 must be a `data.frame`")
-  expect_error(fasano.franceschini.test(S1 = S1, S2 = as.matrix(S2), cores = 1), "S1 and S2 must be a `data.frame`")
-  expect_error(fasano.franceschini.test(S1 = as.matrix(S1), S2 = as.matrix(S2), cores = 1), "S1 and S2 must be a `data.frame`")
-  expect_error(fasano.franceschini.test(S1 = S1, S2 = S2, nPermute = -1, cores = 1), "nPermute must be a positive value")
-})
-
-
-
-test_that("FF test computes with bootstrap, bootstrap is depreciated", {
-
-  # sample 1 - 30 points in quad 1
-  S1 <- data.frame(
-    x = c(rep(1, 30), rep(-1, 4), rep(-1, 3), rep(1, 2), 0),
-    y = c(rep(1, 30), rep(1, 4), rep(-1, 3), rep(-1, 2), 0)
-  )
-  # sample 2 - 1 points in quad 1
-  S2 <- data.frame(
-    x = c(1, rep(-1, 4), rep(-1, 3), rep(1, 2)),
-    y = c(1, rep(1, 4), rep(-1, 3), rep(-1, 2))
-  )
-
-  set.seed(123)
-
-  output <- fasano.franceschini.test(S1 = S1, S2 = S2, nPermute = 1000, cores = 1)
-  expect_equal(unname(output$p.value), 0.001998002)
-  expect_named(output$p.value, "p-value")
-  expect_equal(unname(output$statistic), 0.525)
[…]

### 2022-05-17 v1/tests/testthat/test-interal_functions_FasFran.R

@@ -1,104 +1,0 @@
-test_that("ksCDF output has limits", {
-  # input between 0 and Inf
-  # output between 0 and 1
-  expect_equal(ksCDF(Inf), 0)
-  expect_equal(ksCDF(0), 1)
-  expect_error(ksCDF(-0.1), "lambda must be between 0 and Inf, negative number supplied")
-})
-
-
-
-test_that("quadCount computes accurate normalized fractions", {
-  base <- quadCount(
-    x_origin = 0,
-    y_origin = 0,
-    x = c(rep(1, 4), rep(-1, 3), rep(-1, 2), 1),
-    y = c(rep(1, 4), rep(1, 3), rep(-1, 2), -1)
-  )
-
-  expect_length(base, 4L)
-  expect_equal(base, c(0.4, 0.3, 0.2, 0.1))
-})
-
-
-
-test_that("quadCount handles ties", {
-
-  # Tied with quad 1 and quad 2
-  expect_equal(
-    quadCount(
-      x_origin = 0,
-      y_origin = 0,
-      x = c(1, 0, -1, 1),
-      y = c(1, 1, -1, -1)
-    ),
-    c(0.375, 0.125, 0.25, 0.25)
-  )
-
-  # Tied with quad 3 and quad 4
-  expect_equal(
-    quadCount(
-      x_origin = 0,
-      y_origin = 0,
-      x = c(1, -1, 0, 1),
-      y = c(1, 1, -1, -1)
-    ),
-    c(0.25, 0.25, 0.125, 0.375)
-  )
-
-
-  # Tied with quad 1 and quad 4
-  expect_equal(
-    quadCount(
-      x_origin = 0,
-      y_origin = 0,
-      x = c(1, -1, -1, 1),
-      y = c(0, 1, -1, -1)
-    ),
-    c(0.125, 0.25, 0.25, 0.375)
-  )
-
-  # Tied with quad 2 and quad 3
-  expect_equal(
-    quadCount(
-      x_origin = 0,
-      y_origin = 0,
-      x = c(1, -1, -1, 1),
-      y = c(1, 0, -1, -1)
-    ),
-    c(0.25, 0.125, 0.375, 0.25)
-  )
-
-  # Tied with origin
-  expect_equal(
-    quadCount(
-      x_origin = 0,
-      y_origin = 0,
-      x = c(0, -1, -1, 1),
[…]

### 2022-05-18 R/ff_function.R
Minor documentation changes

@@ -18,7 +18,7 @@
 #' @param p.conf.level confidence level for the confidence interval of the permutation test p-value.
 #' @return A list with class \code{htest} containing the following components:
 #'   \item{statistic}{the value of the test statistic Z.}
-#'   \item{estimate}{the value of the sample test statistics D1 and D2.}
+#'   \item{estimate}{the value of the difference statistics D1 and D2.}
 #'   \item{p.value}{the permutation test p-value.}
 #'   \item{conf.int}{a binomial confidence interval for the p-value.}
 #'   \item{method}{a character string indicating what type of test was performed.}

### 2022-05-18 README.Rmd
Added R CMD CHECK badge

@@
 `r badge_cran_download("fasano.franceschini.test", "grand-total", "green")`
-[![R-CMD-check](https://github.com/cpuritz/fftest/workflows/R-CMD-check/badge.svg)](https://github.com/cpuritz/fftest/actions)
+`r badge_github_actions("nesscoder/fasano.franceschini.test")`
 <!-- badges: end -->

### 2022-05-18 R/ff_function.R
Changed argument order

@@ -11,11 +11,11 @@
 #' @param threads a positive \code{integer} or \code{"auto"} setting the number of threads used for performing
 #' the permutation test. If set to \code{"auto"}, the number of threads is determined by
 #' \code{RcppParallel::defaultNumThreads()}. Default is 1.
-#' @param verbose a \code{boolean} indicating whether to display a progress bar. Default is \code{FALSE}. Only available for serial
-#' version (\code{threads} = 1).
 #' @param seed optional integer to seed the PRNG used for the permutation test. Default is \code{NULL}. Only available for serial
 #' version (\code{threads} = 1).
 #' @param p.conf.level confidence level for the confidence interval of the permutation test p-value.
+#' @param verbose a \code{boolean} indicating whether to display a progress bar. Default is \code{TRUE}. Only available for serial
+#' version (\code{threads} = 1).
 #' @return A list with class \code{htest} containing the following components:
 #'   \item{statistic}{the value of the test statistic Z.}
 #'   \item{estimate}{the value of the difference statistics D1 and D2.}
@@ -43,8 +43,6 @@
 #' fasano.franceschini.test(S1, S2, nPermute = 200)
 #' # set seed for reproducible p-value
 #' fasano.franceschini.test(S1, S2, seed = 0)
-#' # display progress bar
-#' fasano.franceschini.test(S1, S2, verbose = TRUE)
 #' # change confidence level for p-value confidence interval
 #' fasano.franceschini.test(S1, S2, p.conf.level = 0.99)
 #'
@@ -75,7 +73,7 @@
 #' If there are ties in the data, points are counted with multiplicity.
 #'
 #' @export
-fasano.franceschini.test <- function(S1, S2, nPermute = 100, threads = 1, verbose = FALSE, seed = NULL, p.conf.level = 0.95) {
+fasano.franceschini.test <- function(S1, S2, nPermute = 100, threads = 1, seed = NULL, p.conf.level = 0.95, verbose = TRUE) {
     # Store names of samples for output
     dname <- paste(deparse(substitute(S1)), "and", deparse(substitute(S2)))
 

### 2022-05-18 README.Rmd
Updated readme

@@
   fig.path = "man/figures/",
-  out.width = "100%"
+  out.width = "100%",
+  warning = TRUE,
+  message = TRUE
 )

### 2022-05-18 R/ff_function.R
Updated docs

@@ -39,16 +39,20 @@
 #'
 #' # perform test (serial version)
 #' fasano.franceschini.test(S1, S2)
+#'
 #' # perform test with more permutations
 #' fasano.franceschini.test(S1, S2, nPermute = 200)
+#'
 #' # set seed for reproducible p-value
 #' fasano.franceschini.test(S1, S2, seed = 0)
+#'
 #' # change confidence level for p-value confidence interval
 #' fasano.franceschini.test(S1, S2, p.conf.level = 0.99)
 #'
 #' # perform test (parallel version, 2 threads)
 #' \dontrun{
-#' fasano.franceschini.test(S1, S2, threads = 2)}
+#' fasano.franceschini.test(S1, S2, threads = 2)
+#' }
 #'
 #'
 #' # create 3-D mixed samples using matrices

### 2022-05-20 R/ff_function.R
parallel test

@@ -50,6 +50,8 @@
 #' fasano.franceschini.test(S1, S2, p.conf.level = 0.99)
 #'
 #' # perform test (parallel version, 2 threads)
+#' fasano.franceschini.test(S1, S2, threads = 2)
+#' fasano.franceschini.test(S1, S2, threads = 4)
 #' \dontrun{
 #' fasano.franceschini.test(S1, S2, threads = 2)
 #' }

### 2022-05-20 R/ff_function.R
undo last change

@@ -50,8 +50,6 @@
 #' fasano.franceschini.test(S1, S2, p.conf.level = 0.99)
 #'
 #' # perform test (parallel version, 2 threads)
-#' fasano.franceschini.test(S1, S2, threads = 2)
-#' fasano.franceschini.test(S1, S2, threads = 4)
 #' \dontrun{
 #' fasano.franceschini.test(S1, S2, threads = 2)
 #' }

### 2022-05-27 R/RcppExports.R
Added in option to use brute force computation of test statistic

@@ -1,19 +1,19 @@
 # Generated by using Rcpp::compileAttributes() -> do not edit by hand
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
-ffTestStatistic <- function(S1, S2) {
-    .Call(`_fasano_franceschini_test_ffTestStatistic`, S1, S2)
+ffTestStatistic <- function(S1, S2, method) {
+    .Call(`_fasano_franceschini_test_ffTestStatistic`, S1, S2, method)
 }
 
-permutationTestSeeded <- function(S1, S2, nPermute, verbose, seed) {
-    .Call(`_fasano_franceschini_test_permutationTestSeeded`, S1, S2, nPermute, verbose, seed)
+permutationTestSeeded <- function(S1, S2, nPermute, verbose, seed, method) {
+    .Call(`_fasano_franceschini_test_permutationTestSeeded`, S1, S2, nPermute, verbose, seed, method)
 }
 
-permutationTest <- function(S1, S2, nPermute, verbose) {
-    .Call(`_fasano_franceschini_test_permutationTest`, S1, S2, nPermute, verbose)
+permutationTest <- function(S1, S2, nPermute, verbose, method) {
+    .Call(`_fasano_franceschini_test_permutationTest`, S1, S2, nPermute, verbose, method)
 }
 
-permutationTestParallel <- function(S1, S2, nPermute) {
-    .Call(`_fasano_franceschini_test_permutationTestParallel`, S1, S2, nPermute)
+permutationTestParallel <- function(S1, S2, nPermute, method) {
+    .Call(`_fasano_franceschini_test_permutationTestParallel`, S1, S2, nPermute, method)
 }
 

### 2022-05-27 R/ff_function.R

@@ -23,6 +23,10 @@
 #' permutation test p-value.
 #' @param verbose a \code{boolean} indicating whether to display a progress bar.
 #' Default is \code{TRUE}. Only available for serial version (\code{threads} = 1).
+#' @param method a \code{character} indicating which method to use to compute the
+#' test statistic. Must be either 'r' for the range-tree method (default), or 'b'
+#' for the brute force method. Both return the same results, but may vary in
+#' computation speed. See Details section for more information.
 #' @return A list with class \code{htest} containing the following components:
 #'   \item{statistic}{the value of the test statistic Z.}
 #'   \item{estimate}{the value of the difference statistics D1 and D2.}
@@ -74,8 +78,10 @@
 #'             rpois(n = 72, lambda = 5),
 #'             rpois(n = 72, lambda = 5))
 #'
-#' # perform test
-#' fasano.franceschini.test(S1, S2)
+#' # perform test using range tree method
+#' fasano.franceschini.test(S1, S2, seed = 0, method = 'r')
+#' # perform test using brute force method
+#' fasano.franceschini.test(S1, S2, seed = 0, method = 'b')
 #'
 #' @details The test statistic is computed using an efficient range tree
 #' implementation with a time complexity of \emph{O(n*log(n)^(d-1))}, where
@@ -101,7 +107,8 @@ fasano.franceschini.test <- function(S1,
                                      cores,
                                      seed = NULL,
                                      p.conf.level = 0.95,
-                                     verbose = TRUE) {
+                                     verbose = TRUE,
+                                     method = c('r','b')) {
     # Store names of samples for output
     dname <- paste(deparse(substitute(S1)), "and", deparse(substitute(S2)))
 
@@ -142,9 +149,11 @@ fasano.franceschini.test <- function(S1,
     if (!is.numeric(p.conf.level) || p.conf.level <= 0 || p.conf.level >= 1) {
         stop("'p.conf.level' must be a number between 0 and 1")
     }
+    # Validate method
+    method <- match.arg(method)
 
     # Perform FF test
-    ffStats <- ffTestStatistic(S1, S2)
+    ffStats <- ffTestStatistic(S1, S2, method)
     estimate <- c(ffStats[1], ffStats[2])
     names(estimate) <- c("D1", "D2")
     Z <- ffStats[3]
@@ -159,13 +168,13 @@ fasano.franceschini.test <- function(S1,
                 warning("Test cannot be seeded if using multiple threads.")
             }
[…]

### 2022-05-27 src/RcppExports.cpp

@@ -11,20 +11,21 @@ Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
 #endif
 
 // ffTestStatistic
-NumericVector ffTestStatistic(const NumericMatrix S1, const NumericMatrix S2);
-RcppExport SEXP _fasano_franceschini_test_ffTestStatistic(SEXP S1SEXP, SEXP S2SEXP) {
+NumericVector ffTestStatistic(const NumericMatrix S1, const NumericMatrix S2, char method);
+RcppExport SEXP _fasano_franceschini_test_ffTestStatistic(SEXP S1SEXP, SEXP S2SEXP, SEXP methodSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< const NumericMatrix >::type S1(S1SEXP);
     Rcpp::traits::input_parameter< const NumericMatrix >::type S2(S2SEXP);
-    rcpp_result_gen = Rcpp::wrap(ffTestStatistic(S1, S2));
+    Rcpp::traits::input_parameter< char >::type method(methodSEXP);
+    rcpp_result_gen = Rcpp::wrap(ffTestStatistic(S1, S2, method));
     return rcpp_result_gen;
 END_RCPP
 }
 // permutationTestSeeded
-unsigned int permutationTestSeeded(const NumericMatrix S1, const NumericMatrix S2, int nPermute, bool verbose, int seed);
-RcppExport SEXP _fasano_franceschini_test_permutationTestSeeded(SEXP S1SEXP, SEXP S2SEXP, SEXP nPermuteSEXP, SEXP verboseSEXP, SEXP seedSEXP) {
+unsigned int permutationTestSeeded(const NumericMatrix S1, const NumericMatrix S2, int nPermute, bool verbose, int seed, char method);
+RcppExport SEXP _fasano_franceschini_test_permutationTestSeeded(SEXP S1SEXP, SEXP S2SEXP, SEXP nPermuteSEXP, SEXP verboseSEXP, SEXP seedSEXP, SEXP methodSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
@@ -33,13 +34,14 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< int >::type nPermute(nPermuteSEXP);
[…]

### 2022-05-27 src/ff_internal_functions.cpp

@@ -7,6 +7,7 @@
 #include "matrix_util.h"
 #include "permute.h"
 #include "range_count.h"
+#include "range_count_brute.h"
 #include "ProgressBar.h"
 
 using namespace Rcpp;
@@ -18,13 +19,15 @@ using namespace Rcpp;
 // @param r2 number of rows in second sample
 // @param shuffle whether to shuffle the samples
 // @param prng a pseudorandom number generator (unused if shuffle == false)
+// @param method what method to use to compute the test statistic
 // @return FF test statistics
 template<typename MatrixT>
 std::vector<double> testStatistics(const MatrixT M,
                                    std::size_t r1,
                                    std::size_t r2,
                                    bool shuffle,
-                                   std::mt19937& prng) {
+                                   std::mt19937& prng,
+                                   char method) {
     double n1 = static_cast<double>(r1);
     double n2 = static_cast<double>(r2);
 
@@ -35,27 +38,51 @@ std::vector<double> testStatistics(const MatrixT M,
     } else {
         s = seq(r1 + r2);
     }
-    std::vector<RTree> trees = buildRangeTrees<MatrixT>(M, r1, r2, s);
 
-    // Compute test statistic using first sample as origins
     double d1 = -1;
-    for (std::size_t i = 0; i < r1; ++i) {
-        std::vector<double> org = getRow<MatrixT>(M, s[i]);
-        std::vector<double> c1 = rangeCount(trees[0], org);
-        std::vector<double> c2 = rangeCount(trees[1], org);
-        for (std::size_t j = 0; j < c1.size(); ++j) {
-            d1 = std::max(d1, abs(c1[j]/n1 - c2[j]/n2));
-        }
-    }
-
-    // Compute test statistic using second sample as origins
     double d2 = -1;
-    for (std::size_t i = 0; i < r2; ++i) {
-        std::vector<double> org = getRow<MatrixT>(M, s[i+r1]);
-        std::vector<double> c1 = rangeCount(trees[0], org);
-        std::vector<double> c2 = rangeCount(trees[1], org);
-        for (std::size_t j = 0; j < c1.size(); ++j) {
-            d2 = std::max(d2, abs(c1[j]/n1 - c2[j]/n2));
+    if (method == 'r') {
+        // Compute range trees
+        std::vector<RTree> trees = buildRangeTrees<MatrixT>(M, r1, r2, s);
+
+        // Compute test statistic using first sample as origins
[…]

### 2022-05-27 src/range_count.h

@@ -25,17 +25,19 @@ std::vector<RTree> buildRangeTrees(const MatrixT M,
                                    std::size_t r1,
                                    std::size_t r2,
                                    const std::vector<std::size_t>& s) {
-    std::vector<RangeTree::Point<double,int> > pts1(r1);
-    std::vector<RangeTree::Point<double,int> > pts2(r2);
+    std::vector<RangeTree::Point<double,int> > pts1;
+    pts1.reserve(r1);
+    std::vector<RangeTree::Point<double,int> > pts2;
+    pts2.reserve(r2);
 
     // Extract points from first sample
     for (std::size_t i = 0; i < r1; ++i) {
-        pts1[i] = RangeTree::Point<double,int>(getRow<MatrixT>(M, s[i]), 0);
+        pts1.push_back(RangeTree::Point<double,int>(getRow<MatrixT>(M, s[i]), 0));
     }
 
     // Extract points from second sample
     for (std::size_t i = 0; i < r2; ++i) {
-        pts2[i] = RangeTree::Point<double,int>(getRow<MatrixT>(M, s[i+r1]), 0);
+        pts2.push_back(RangeTree::Point<double,int>(getRow<MatrixT>(M, s[i+r1]), 0));
     }
 
     // Construct range trees

### 2022-05-27 src/range_count_brute.h

@@ -1,0 +1,27 @@
+template<typename MatrixT>
+std::vector<double> rangeCountBrute(const MatrixT M,
+                                    std::size_t npts,
+                                    std::size_t offset,
+                                    const std::vector<std::size_t>& s,
+                                    const std::vector<double>& origin) {
+    std::size_t ndim = origin.size();
+    std::size_t noct = 1<<ndim;
+    std::vector<double> counts(noct);
+
+    for (std::size_t i = 0; i < npts; ++i) {
+        std::vector<double> pt = getRow<MatrixT>(M, s[i + offset]);
+        for (std::size_t j = 0; j < noct; ++j) {
+            bool in_oct = true;
+            for (std::size_t k = 0; k < ndim; ++k) {
+                bool s = j & (1 << (ndim-1-k));
+                bool check = s ? (pt[k] < origin[k]) : (pt[k] > origin[k]);
+                in_oct = in_oct && check;
+            }
+            if (in_oct) {
+                ++counts[j];
+                break;
+            }
+        }
+    }
+    return counts;
+}

### 2022-05-27 tests/testthat/test_2d.R

@@ -1,4 +1,4 @@
-test_that("compare with naive 2d implementation", {
+test_that("compare with naive R 2d implementation", {
     set.seed(2)
 
     ffcomp <- function(S1, S2) {

### 2022-05-27 tests/testthat/test_2d_2.R

@@ -1,0 +1,34 @@
+test_that("compare with brute force 2d implementation", {
+    set.seed(0)
+
+    ffcomp <- function(S1, S2) {
+        res1 <- fasano.franceschini.test(S1, S2, nPermute = 0, method = 'r')
+        res2 <- fasano.franceschini.test(S1, S2, nPermute = 0, method = 'b')
+        expect_equal(res1$estimate, res2$estimate, tolerance = 1e-14)
+        expect_equal(res1$statistic, res2$statistic, tolerance = 1e-14)
+    }
+
+    S1 <- data.frame(rnorm(n = 50, mean = 0, sd = 1),
+                     rnorm(n = 50, mean = 1, sd = 1))
+    S2 <- data.frame(rnorm(n = 83, mean = 0.2, sd = 1),
+                     rnorm(n = 83, mean = 0.1, sd = 2))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rnorm(n = 134, mean = 11, sd = 2),
+                     rnorm(n = 134, mean = 11, sd = 2))
+    S2 <- data.frame(rnorm(n = 45, mean = 11, sd = 2),
+                     rnorm(n = 45, mean = 11, sd = 2))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 200, shape1 = 1.4, shape2 = 1.73),
+                     rbeta(n = 200, shape1 = 1.4, shape2 = 1.73))
+    S2 <- data.frame(rbeta(n = 23, shape1 = 1.4, shape2 = 1.73),
+                     rbeta(n = 23, shape1 = 1.4, shape2 = 1.73))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 324, shape1 = 10.2, shape2 = 0.3),
+                     rbeta(n = 324, shape1 = 10.2, shape2 = 0.3))
+    S2 <- data.frame(rbeta(n = 1559, shape1 = 10.2, shape2 = 0.3),
+                     rbeta(n = 1559, shape1 = 10.2, shape2 = 0.3))
+    ffcomp(S1, S2)
+})

### 2022-05-27 tests/testthat/test_3d.R

@@ -1,4 +1,4 @@
-test_that("compare with naive 3d implementation", {
+test_that("compare with naive R 3d implementation", {
     set.seed(3)
 
     ffcomp <- function(S1, S2) {

### 2022-05-27 tests/testthat/test_3d_2.R

@@ -1,0 +1,42 @@
+test_that("compare with brute force 3d implementation", {
+    set.seed(3)
+
+    ffcomp <- function(S1, S2) {
+        res1 <- fasano.franceschini.test(S1, S2, nPermute = 0, method = 'r')
+        res2 <- fasano.franceschini.test(S1, S2, nPermute = 0, method = 'b')
+        expect_equal(res1$estimate, res2$estimate, tolerance = 1e-14)
+        expect_equal(res1$statistic, res2$statistic, tolerance = 1e-14)
+    }
+
+    S1 <- data.frame(rnorm(n = 35, mean = 1.4, sd = 0.2),
+                     rnorm(n = 35, mean = 1.5, sd = 2),
+                     rnorm(n = 35, mean = 1.6, sd = 1.8))
+    S2 <- data.frame(rnorm(n = 207, mean = 1.4, sd = 0.2),
+                     rnorm(n = 207, mean = 1.5, sd = 2),
+                     rnorm(n = 207, mean = 1.6, sd = 1.8))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 101, shape1 = 1.2, shape2 = 2.73),
+                     rbeta(n = 101, shape1 = 1.2, shape2 = 0.11),
+                     rbeta(n = 101, shape1 = 2.1, shape2 = 1.39))
+    S2 <- data.frame(rbeta(n = 66, shape1 = 1.3, shape2 = 2.73),
+                     rbeta(n = 66, shape1 = 1.2, shape2 = 0.12),
+                     rbeta(n = 66, shape1 = 2.1, shape2 = 1.39))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 403, shape1 = 1.1, shape2 = 2.71),
+                     rbeta(n = 403, shape1 = 1.3, shape2 = 0.14),
+                     rbeta(n = 403, shape1 = 2.4, shape2 = 1.35))
+    S2 <- data.frame(rbeta(n = 780, shape1 = 1.5, shape2 = 2.76),
+                     rbeta(n = 780, shape1 = 1.7, shape2 = 0.18),
+                     rbeta(n = 780, shape1 = 0.6, shape2 = 1.38))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 402, shape1 = 1.3, shape2 = 2.43),
[…]

### 2022-05-27 tests/testthat/test_4d.R

@@ -1,4 +1,4 @@
-test_that("compare with naive 4d implementation", {
+test_that("compare with naive R 4d implementation", {
     set.seed(4)
 
     ffcomp <- function(S1, S2) {

### 2022-05-27 tests/testthat/test_4d_2.R

@@ -1,0 +1,40 @@
+test_that("compare with brute force 4d implementation", {
+    set.seed(4)
+
+    ffcomp <- function(S1, S2) {
+        res1 <- fasano.franceschini.test(S1, S2, nPermute = 0, method = 'r')
+        res2 <- fasano.franceschini.test(S1, S2, nPermute = 0, method = 'b')
+        expect_equal(res1$estimate, res2$estimate, tolerance = 1e-14)
+        expect_equal(res1$statistic, res2$statistic, tolerance = 1e-14)
+    }
+
+    S1 <- data.frame(rnorm(n = 70, mean = 1.4, sd = 0.3),
+                     rnorm(n = 70, mean = 1.5, sd = 1),
+                     rnorm(n = 70, mean = 1.6, sd = 1.8),
+                     rnorm(n = 70, mean = 1.6, sd = 1.8))
+    S2 <- data.frame(rnorm(n = 206, mean = 1.4, sd = 0.3),
+                     rnorm(n = 206, mean = 1.5, sd = 1),
+                     rnorm(n = 206, mean = 1.6, sd = 1.8),
+                     rnorm(n = 206, mean = 2.2, sd = 0.01))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 104, shape1 = 1.3, shape2 = 2.53),
+                     rgamma(n = 104, shape = 4.4, rate = 0.4),
+                     rbeta(n = 104, shape1 = 1.2, shape2 = 0.31),
+                     rbeta(n = 104, shape1 = 2.1, shape2 = 1.39))
+    S2 <- data.frame(rbeta(n = 67, shape1 = 1.3, shape2 = 2.63),
+                     rbeta(n = 67, shape1 = 1.2, shape2 = 0.12),
+                     rbeta(n = 67, shape1 = 2.1, shape2 = 1.39),
+                     rexp(n = 67, rate = 32.3))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 401, shape1 = 2.3, shape2 = 2.73),
+                     rbeta(n = 401, shape1 = 2.2, shape2 = 0.12),
+                     rbeta(n = 401, shape1 = 1.1, shape2 = 1.39),
+                     rgamma(n = 401, shape = 0.9, rate = 3.01))
+    S2 <- data.frame(rbeta(n = 736, shape1 = 2.3, shape2 = 2.23),
[…]

### 2022-05-27 tests/testthat/test_5d.R

@@ -1,4 +1,4 @@
-test_that("compare with naive 5d implementation", {
+test_that("compare with naive R 5d implementation", {
     set.seed(5)
 
     ffcomp <- function(S1, S2) {

### 2022-05-27 tests/testthat/test_5d_2.R

@@ -1,0 +1,46 @@
+test_that("compare with brute force 5d implementation", {
+    set.seed(5)
+
+    ffcomp <- function(S1, S2) {
+        res1 <- fasano.franceschini.test(S1, S2, nPermute = 0, method = 'r')
+        res2 <- fasano.franceschini.test(S1, S2, nPermute = 0, method = 'b')
+        expect_equal(res1$estimate, res2$estimate, tolerance = 1e-14)
+        expect_equal(res1$statistic, res2$statistic, tolerance = 1e-14)
+    }
+
+    S1 <- data.frame(rnorm(n = 70, mean = 1.4, sd = 0.3),
+                     rnorm(n = 70, mean = 1.5, sd = 1),
+                     rnorm(n = 70, mean = 2.2, sd = 1.2),
+                     rnorm(n = 70, mean = 2.2, sd = 1.8),
+                     rnorm(n = 70, mean = -1.1, sd = 1.8))
+    S2 <- data.frame(rnorm(n = 106, mean = 1.4, sd = 0.3),
+                     rnorm(n = 106, mean = 1.5, sd = 1),
+                     rnorm(n = 106, mean = 1.7, sd = 1.8),
+                     rnorm(n = 106, mean = 2.2, sd = 0.01),
+                     rnorm(n = 106, mean = 2.2, sd = 0.05))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 102, shape1 = 1.3, shape2 = 2.63),
+                     rgamma(n = 102, shape = 4.4, rate = 0.4),
+                     rbeta(n = 102, shape1 = 1.2, shape2 = 0.11),
+                     rbeta(n = 102, shape1 = 2.1, shape2 = 1.39),
+                     rbeta(n = 102, shape1 = 2.1, shape2 = 1.39))
+    S2 <- data.frame(rbeta(n = 65, shape1 = 1.3, shape2 = 2.73),
+                     rbeta(n = 65, shape1 = 1.2, shape2 = 0.11),
+                     rbeta(n = 65, shape1 = 2.1, shape2 = 1.39),
+                     rexp(n = 65, rate = 33.3),
+                     rexp(n = 65, rate = 33.3))
+    ffcomp(S1, S2)
+
+    S1 <- data.frame(rbeta(n = 300, shape1 = 1.3, shape2 = 2.73),
+                     rbeta(n = 300, shape1 = 1.2, shape2 = 0.11),
[…]

### 2022-05-27 tests/testthat/test_errors.R

@@ -22,11 +22,6 @@ test_that("test for bad inputs", {
     expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, threads = 0.1))
     expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, threads = 1.1))
 
-    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, cores = -1))
-    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, cores = 0))
-    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, cores = 0.1))
-    expect_error(fasano.franceschini.test(S1, S2, nPermute = 0, cores = 1.1))
-
     ## Check for deprecation warning, remove in future releases
     expect_warning(fasano.franceschini.test(S1, S2, nPermute = 0, cores = 1))
     ##
@@ -40,4 +35,7 @@ test_that("test for bad inputs", {
     S2 <- as.matrix(rnorm(n = 60))
     expect_error(fasano.franceschini.test(S1, S2, nPermute = 0))
     expect_error(fasano.franceschini.test(S2, S1, nPermute = 0))
+
+    expect_error(fasano.franceschini.test(S1, S2, method = "a"))
+    expect_error(fasano.franceschini.test(S1, S2, method = 1))
 })

### 2022-05-27 src/ProgressBar.cpp
Minor reorganization

@@ -1,0 +1,32 @@
+#include "ProgressBar.h"
+
+ProgressBar::ProgressBar(int nsteps, bool verbose) :
+    nsteps(nsteps), ticksPerStep(MAX_TICKS/nsteps+1), verbose(verbose), ticks(0), progress(0) {
+    if (verbose) {
+        Rcpp::Rcerr << "0%   10   20   30   40   50   60   70   80   90   100%\n" <<
+            "|----|----|----|----|----|----|----|----|----|----|\n";
+    }
+}
+
+void ProgressBar::step() {
+    if (verbose) {
+        progress += 1/static_cast<double>(nsteps);
+        if (progress > ticks/static_cast<double>(MAX_TICKS) && ticks < MAX_TICKS) {
+            for (int i = 0; i < ticksPerStep; ++i) {
+                Rcpp::Rcerr << "*";
+                ++ticks;
+                if (ticks == MAX_TICKS) {
+                    break;
+                }
+            }
+        }
+    }
+}
+
+void ProgressBar::finalize() {
+    if (verbose) {
+        for (int i = ticks; i < MAX_TICKS; ++i) {
+            Rcpp::Rcerr << "*";
+        }
+    }
+}

### 2022-05-27 src/ProgressBar.h

@@ -20,37 +20,11 @@ private:
     // Number of ticks to print
     static const int MAX_TICKS = 51;
 public:
-    ProgressBar(int nsteps, bool verbose) :
-        nsteps(nsteps), ticksPerStep(MAX_TICKS/nsteps+1), verbose(verbose), ticks(0), progress(0) {
-        if (verbose) {
-            Rcpp::Rcerr << "0%   10   20   30   40   50   60   70   80   90   100%\n" <<
-                           "|----|----|----|----|----|----|----|----|----|----|\n";
-        }
-    }
-
-    void step() {
-        if (verbose) {
-            progress += 1/static_cast<double>(nsteps);
-            if (progress > ticks/static_cast<double>(MAX_TICKS) && ticks < MAX_TICKS) {
-                for (int i = 0; i < ticksPerStep; ++i) {
-                    Rcpp::Rcerr << "*";
-                    ++ticks;
-                    if (ticks == MAX_TICKS) {
-                        break;
-                    }
-                }
-            }
-        }
-    }
-
+    ProgressBar(int nsteps, bool verbose);
+    // Increment the progress bar
+    void step();
     // Double check all ticks have been printed at the end
-    void finalize() {
-        if (verbose) {
-            for (int i = ticks; i < MAX_TICKS; ++i) {
-                Rcpp::Rcerr << "*";
-            }
-        }
-    }
+    void finalize();
 };
 
 #endif //PROGRESS_H

### 2022-05-27 src/ff_internal_functions.cpp

@@ -7,10 +7,21 @@
 #include "matrix_util.h"
 #include "permute.h"
 #include "range_count.h"
-#include "range_count_brute.h"
 #include "ProgressBar.h"
 
 using namespace Rcpp;
+
+// Helper function for maximizing difference statistics
+double propDiff(const std::vector<double>& c1,
+                const std::vector<double>& c2,
+                std::size_t n1,
+                std::size_t n2) {
+    double d = -1;
+    for (std::size_t i = 0; i < c1.size(); ++i) {
+        d = std::max(d, abs(c1[i]/n1 - c2[i]/n2));
+    }
+    return d;
+}
 
 // Compute FF test statistics
 //
@@ -42,50 +53,41 @@ std::vector<double> testStatistics(const MatrixT M,
     double d1 = -1;
     double d2 = -1;
     if (method == 'r') {
-        // Compute range trees
+        // Build range trees
         std::vector<RTree> trees = buildRangeTrees<MatrixT>(M, r1, r2, s);
 
         // Compute test statistic using first sample as origins
         for (std::size_t i = 0; i < r1; ++i) {
             std::vector<double> org = getRow<MatrixT>(M, s[i]);
-            std::vector<double> c1 = rangeCount(trees[0], org);
-            std::vector<double> c2 = rangeCount(trees[1], org);
-            for (std::size_t j = 0; j < c1.size(); ++j) {
-                d1 = std::max(d1, abs(c1[j]/n1 - c2[j]/n2));
-            }
+            d1 = std::max(d1, propDiff(rangeCountTree(trees[0], org),
+                                       rangeCountTree(trees[1], org),
+                                       n1, n2));
         }
 
         // Compute test statistic using second sample as origins
         for (std::size_t i = 0; i < r2; ++i) {
             std::vector<double> org = getRow<MatrixT>(M, s[i+r1]);
-            std::vector<double> c1 = rangeCount(trees[0], org);
-            std::vector<double> c2 = rangeCount(trees[1], org);
-            for (std::size_t j = 0; j < c1.size(); ++j) {
-                d2 = std::max(d2, abs(c1[j]/n1 - c2[j]/n2));
-            }
+            d2 = std::max(d2, propDiff(rangeCountTree(trees[0], org),
+                                       rangeCountTree(trees[1], org),
+                                       n1, n2));
         }
     } else {
[…]

### 2022-05-27 src/range_count.h

@@ -3,13 +3,10 @@
 #ifndef RANGECOUNT_H
 #define RANGECOUNT_H
 
-#include <Rcpp.h>
 #include <vector>
 #include <limits>
 #include "RangeTree.h"
 #include "matrix_util.h"
-
-using namespace Rcpp;
 
 typedef RangeTree::RangeTree<double,int> RTree;
 
@@ -44,13 +41,14 @@ std::vector<RTree> buildRangeTrees(const MatrixT M,
     return {RTree(pts1), RTree(pts2)};
 }
 
-// Count the number of points in each axis-aligned orthant defined by the origin
+// Count the number of points in each axis-aligned orthant defined by the origin.
+// Uses range tree method.
 //
 // @param rtree range tree
 // @param origin vector
 // @return vector counting points in each orthant
-std::vector<double> rangeCount(const RTree& rtree,
-                               const std::vector<double>& origin) {
+std::vector<double> rangeCountTree(const RTree& rtree,
+                                   const std::vector<double>& origin) {
     std::size_t ndim = origin.size();
     std::size_t noct = 1<<ndim;
     std::vector<double> res(noct);
@@ -75,6 +73,47 @@ std::vector<double> rangeCount(const RTree& rtree,
         res[i] = rtree.countInRange(lowerLims, upperLims, strict, strict);
     }
     return res;
+}
+
+// Count the number of points in each axis-aligned orthant defined by the origin.
+// Uses brute force method.
+//
+// @param M a matrix of type NumericMatrix or RMatrix<double>
+// @param npts how many points are in the sample
+// @param offset where to start counting the points
+// @param shuffle whether to shuffle the samples
+// @param s a permutation of {0,1,...,(r1+r2-1)} which indicates which points belong to each sample
+// @param origin vector
+// @return vector counting points in each orthant
+template<typename MatrixT>
+std::vector<double> rangeCountBrute(const MatrixT M,
+                                    std::size_t npts,
+                                    std::size_t offset,
+                                    const std::vector<std::size_t>& s,
+                                    const std::vector<double>& origin) {
+    std::size_t ndim = origin.size();
+    std::size_t noct = 1<<ndim;
+    std::vector<double> counts(noct);
+
+    for (std::size_t i = 0; i < npts; ++i) {
+        std::vector<double> pt = getRow<MatrixT>(M, s[i + offset]);
+        for (std::size_t j = 0; j < noct; ++j) {
+            bool in_oct = true;
[…]

### 2022-05-27 src/range_count_brute.h

@@ -1,27 +1,0 @@
-template<typename MatrixT>
-std::vector<double> rangeCountBrute(const MatrixT M,
-                                    std::size_t npts,
-                                    std::size_t offset,
-                                    const std::vector<std::size_t>& s,
-                                    const std::vector<double>& origin) {
-    std::size_t ndim = origin.size();
-    std::size_t noct = 1<<ndim;
-    std::vector<double> counts(noct);
-
-    for (std::size_t i = 0; i < npts; ++i) {
-        std::vector<double> pt = getRow<MatrixT>(M, s[i + offset]);
-        for (std::size_t j = 0; j < noct; ++j) {
-            bool in_oct = true;
-            for (std::size_t k = 0; k < ndim; ++k) {
-                bool s = j & (1 << (ndim-1-k));
-                bool check = s ? (pt[k] < origin[k]) : (pt[k] > origin[k]);
-                in_oct = in_oct && check;
-            }
-            if (in_oct) {
-                ++counts[j];
-                break;
-            }
-        }
-    }
-    return counts;
-}

### 2022-05-27 R/ff_function.R
Minor cleanups

@@ -15,8 +15,8 @@
 #' of threads used for performing the permutation test. If set to \code{"auto"},
 #' the number of threads is determined by \code{RcppParallel::defaultNumThreads()}.
 #' Default is 1.
-#' @param cores allowed for backwards compatibility. \code{threads} is now the preferred
-#' argument name.
+#' @param cores allowed for backwards compatibility. \code{threads} is now the
+#' preferred argument name.
 #' @param seed optional integer to seed the PRNG used for the permutation test.
 #' Default is \code{NULL}. Only available for serial version (\code{threads} = 1).
 #' @param p.conf.level confidence level for the confidence interval of the
@@ -24,9 +24,9 @@
 #' @param verbose a \code{boolean} indicating whether to display a progress bar.
 #' Default is \code{TRUE}. Only available for serial version (\code{threads} = 1).
 #' @param method a \code{character} indicating which method to use to compute the
-#' test statistic. Must be either 'r' for the range-tree method (default), or 'b'
-#' for the brute force method. Both return the same results, but may vary in
-#' computation speed. See Details section for more information.
+#' test statistic. Must be either \code{'r'} for the range-tree method (default),
+#' or \code{'b'} for the brute force method. Both return the same results, but
+#' may vary in computation speed. See the Details section for more information.
 #' @return A list with class \code{htest} containing the following components:
 #'   \item{statistic}{the value of the test statistic Z.}
 #'   \item{estimate}{the value of the difference statistics D1 and D2.}
@@ -80,13 +80,23 @@
 #'
 #' # perform test using range tree method
 #' fasano.franceschini.test(S1, S2, seed = 0, method = 'r')
+#'
 #' # perform test using brute force method
 #' fasano.franceschini.test(S1, S2, seed = 0, method = 'b')
 #'
-#' @details The test statistic is computed using an efficient range tree
-#' implementation with a time complexity of \emph{O(n*log(n)^(d-1))}, where
-#' \emph{n} is the size of the largest sample and \emph{d} is the dimension
-#' of the data.
+#' @details The test statistic can be computed using two different methods.
+#' Both methods return identical results, but vary in computation time.
+#' \itemize{
+#'   \item Range tree method (\code{method = 'r'}): This method has a time
+#'   complexity of \emph{O(n*log(n)^(d-1))}, where \emph{n} is the size of
+#'   the larger sample and \emph{d} is the dimension of the data.
[…]

### 2022-05-28 tests/testthat/test_2d.R
Added a few tests

@@ -38,4 +38,18 @@ test_that("compare with naive R 2d implementation", {
     S2 <- data.frame(rbeta(n = 1557, shape1 = 10.1, shape2 = 0.1),
                      rbeta(n = 1557, shape1 = 10.1, shape2 = 0.1))
     ffcomp(S1, S2)
+
+    S1 <- cbind(rpois(n = 100, lambda = 4),
+                rpois(n = 100, lambda = 3))
+    S2 <- cbind(rpois(n = 100, lambda = 4),
+                rpois(n = 100, lambda = 3))
+    ffcomp(S1, S2)
+
+    S1 <- cbind(rep(1, 100), rep(2, 100))
+    S2 <- cbind(rep(1, 100), rep(2, 100))
+    ffcomp(S1, S2)
+
+    S1 <- cbind(rep(1, 100), rep(2, 100))
+    S2 <- cbind(rep(2, 100), rep(1, 100))
+    ffcomp(S1, S2)
 })

### 2022-05-28 tests/testthat/test_2d_2.R

@@ -31,4 +31,18 @@ test_that("compare with brute force 2d implementation", {
     S2 <- data.frame(rbeta(n = 1559, shape1 = 10.2, shape2 = 0.3),
                      rbeta(n = 1559, shape1 = 10.2, shape2 = 0.3))
     ffcomp(S1, S2)
+
+    S1 <- cbind(rpois(n = 100, lambda = 4),
+                rpois(n = 100, lambda = 3))
+    S2 <- cbind(rpois(n = 100, lambda = 4),
+                rpois(n = 100, lambda = 3))
+    ffcomp(S1, S2)
+
+    S1 <- cbind(rep(1, 100), rep(2, 100))
+    S2 <- cbind(rep(1, 100), rep(2, 100))
+    ffcomp(S1, S2)
+
+    S1 <- cbind(rep(1, 100), rep(2, 100))
+    S2 <- cbind(rep(2, 100), rep(1, 100))
+    ffcomp(S1, S2)
 })

### 2022-05-28 R/ff_function.R
Minor documentation changes

@@ -93,10 +93,10 @@
 #'   \item Brute force method (\code{method = 'b'}): This method has a time
 #'   complexity of \emph{O(n^2)}.
 #' }
-#' When \emph{d = 2}, or \emph{d > 2} and \emph{n} is large, the range tree
-#' method tends to outperform the brute force method. When \emph{d > 2} and
-#' \emph{n} is small, the brute force method tends to outperform the range
-#' tree method.
+#' When \emph{d = 2} (regardless of \emph{n}), or \emph{d > 2} and \emph{n} is
+#' large, the range tree method tends to outperform the brute force method.
+#' When \emph{d > 2} and \emph{n} is small, the brute force method tends to
+#' outperform the range tree method.
 #'
 #' The p-value for the test is computed empirically using a permutation test. As
 #' it is almost always infeasible to compute the exact permutation test p-value,

### 2023-05-04 R/ff_function.R
Temp add of old pvalue

@@ -172,11 +172,14 @@ fasano.franceschini.test <- function(S1,
             pval <- permutationTestPvalueSeeded(counts[1], counts[2], nPermute, seed)
         }
         names(pval) <- "p-value"
+
+        pval2 <- (1 + counts[1] + counts[2]) / (1 + nPermute)
     }
 
     # Construct output
     result <- list(statistic = Dff,
                    p.value = pval,
+                   p.value.old = pval2,
                    method = "Fasano-Franceschini Test",
                    data.name = dname)
     class(result) <- "htest"

### 2023-05-04 R/ff_function.R
Probably final update

@@ -172,14 +172,11 @@ fasano.franceschini.test <- function(S1,
             pval <- permutationTestPvalueSeeded(counts[1], counts[2], nPermute, seed)
         }
         names(pval) <- "p-value"
-
-        pval2 <- (1 + counts[1] + counts[2]) / (1 + nPermute)
     }
 
     # Construct output
     result <- list(statistic = Dff,
                    p.value = pval,
-                   p.value.old = pval2,
                    method = "Fasano-Franceschini Test",
                    data.name = dname)
     class(result) <- "htest"

### 2023-05-04 README.Rmd
Minor documentation changes
Updated docs

@@
 
-You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://CRAN.R-project.org) with:
+You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://cran.r-project.org/web/packages/fasano.franceschini.test/index.html) with:
 
@@
 
-And the development version of the `fasano.franceschini.test` package from [GitHub](https://github.com/) with:
+The development version of the `fasano.franceschini.test` package can be installed from [GitHub](https://github.com/braunlab-nu/fasano.franceschini.test/) with:
 

### 2023-05-08 README.Rmd
Prepare for 2.2.0 release

--- 2022-05-18 README.Rmd
@@
 `r badge_cran_release("fasano.franceschini.test", "orange")`
-`r badge_devel("nesscoder/fasano.franceschini.test", "blue")` 
+`r badge_devel("braunlab-nu/fasano.franceschini.test", "blue")` 
 `r badge_cran_download("fasano.franceschini.test", "grand-total", "green")`
-`r badge_github_actions("nesscoder/fasano.franceschini.test")`
+`r badge_github_actions("braunlab-nu/fasano.franceschini.test")`
 <!-- badges: end -->
@@
 
-You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://CRAN.R-project.org) with:
+You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://cran.r-project.org/) with:
 
@@
 
-And the development version of the `fasano.franceschini.test` package from [GitHub](https://github.com/) with:
+The development version of the `fasano.franceschini.test` package can be installed from Github with:
 
@@
 # install.packages("devtools")
-devtools::install_github("nesscoder/fasano.franceschini.test")
+devtools::install_github("braunlab-nu/fasano.franceschini.test")
 ```

### 2023-05-08 src/ff_internal_functions.cpp
Fixed issue with p-value computation

@@ -149,7 +149,7 @@ long ffTestStatistic(const NumericMatrix& S1,
 
 /*******************************************************************************/
 
-/***************** This function is for computing the p-value. *****************/
+/************** These functions are for computing the p-value. *****************/
 
 // [[Rcpp::export(permutationTestPvalue)]]
 double permutationTestPvalue(unsigned int zLess,
@@ -157,7 +157,7 @@ double permutationTestPvalue(unsigned int zLess,
                              unsigned int nPermute) {
     std::mt19937 prng(std::random_device{}());
     std::uniform_real_distribution<double> dis(0.0, 1.0);
-    return (zLess + (1 + zEqual) * dis(prng)) / static_cast<double>(nPermute);
+    return (zLess + (1 + zEqual) * dis(prng)) / static_cast<double>(1 + nPermute);
 }
 
 // [[Rcpp::export(permutationTestPvalueSeeded)]]
@@ -167,7 +167,7 @@ double permutationTestPvalueSeeded(unsigned int zLess,
                                    int seed) {
     std::mt19937 prng(seed);
     std::uniform_real_distribution<double> dis(0.0, 1.0);
-    return (zLess + (1 + zEqual) * dis(prng)) / static_cast<double>(nPermute);
+    return (zLess + (1 + zEqual) * dis(prng)) / static_cast<double>(1 + nPermute);
 }
 
 /*******************************************************************************/

### 2023-05-08 R/ff_function.R
Temp commit for pval comparison

@@ -172,11 +172,13 @@ fasano.franceschini.test <- function(S1,
             pval <- permutationTestPvalueSeeded(counts[1], counts[2], nPermute, seed)
         }
         names(pval) <- "p-value"
+        pval_old <- permutationTestPvalueOld(counts[1], counts[2], nPermute)
     }
 
     # Construct output
     result <- list(statistic = Dff,
                    p.value = pval,
+                   p.value.old = pval_old,
                    method = "Fasano-Franceschini Test",
                    data.name = dname)
     class(result) <- "htest"

### 2023-05-08 src/ff_internal_functions.cpp

@@ -160,6 +160,13 @@ double permutationTestPvalue(unsigned int zLess,
     return (zLess + (1 + zEqual) * dis(prng)) / static_cast<double>(1 + nPermute);
 }
 
+// [[Rcpp::export(permutationTestPvalueOld)]]
+double permutationTestPvalue(unsigned int zLess,
+                             unsigned int zEqual,
+                             unsigned int nPermute) {
+    return (1 + zLess + zEqual) / static_cast<double>(1 + nPermute);
+}
+
 // [[Rcpp::export(permutationTestPvalueSeeded)]]
 double permutationTestPvalueSeeded(unsigned int zLess,
                                    unsigned int zEqual,

### 2023-05-08 src/ff_internal_functions.cpp
Temp commit for pval comparison

@@ -161,7 +161,7 @@ double permutationTestPvalue(unsigned int zLess,
 }
 
 // [[Rcpp::export(permutationTestPvalueOld)]]
-double permutationTestPvalue(unsigned int zLess,
+double permutationTestPvalueOld(unsigned int zLess,
                              unsigned int zEqual,
                              unsigned int nPermute) {
     return (1 + zLess + zEqual) / static_cast<double>(1 + nPermute);

### 2023-05-08 R/ff_function.R
Revert

@@ -172,13 +172,11 @@ fasano.franceschini.test <- function(S1,
             pval <- permutationTestPvalueSeeded(counts[1], counts[2], nPermute, seed)
         }
         names(pval) <- "p-value"
-        pval_old <- permutationTestPvalueOld(counts[1], counts[2], nPermute)
     }
 
     # Construct output
     result <- list(statistic = Dff,
                    p.value = pval,
-                   p.value.old = pval_old,
                    method = "Fasano-Franceschini Test",
                    data.name = dname)
     class(result) <- "htest"

### 2023-05-08 src/ff_internal_functions.cpp

@@ -160,13 +160,6 @@ double permutationTestPvalue(unsigned int zLess,
     return (zLess + (1 + zEqual) * dis(prng)) / static_cast<double>(1 + nPermute);
 }
 
-// [[Rcpp::export(permutationTestPvalueOld)]]
-double permutationTestPvalueOld(unsigned int zLess,
-                             unsigned int zEqual,
-                             unsigned int nPermute) {
-    return (1 + zLess + zEqual) / static_cast<double>(1 + nPermute);
-}
-
 // [[Rcpp::export(permutationTestPvalueSeeded)]]
 double permutationTestPvalueSeeded(unsigned int zLess,
                                    unsigned int zEqual,

### 2023-05-08 R/ff_function.R
Temp

@@ -148,6 +148,7 @@ fasano.franceschini.test <- function(S1,
 
     # Compute the p-value (if necessary)
     pval <- NULL
+    pval_old <- NULL
     if (nPermute > 0) {
         if (threads > 1) {
             # Run parallel version of permutation test
@@ -172,11 +173,14 @@ fasano.franceschini.test <- function(S1,
             pval <- permutationTestPvalueSeeded(counts[1], counts[2], nPermute, seed)
         }
         names(pval) <- "p-value"
+
+        pval_old <- permutationTestPvalueOld(counts[1], counts[2], nPermute)
     }
 
     # Construct output
     result <- list(statistic = Dff,
                    p.value = pval,
+                   p.value.old = pval_old,
                    method = "Fasano-Franceschini Test",
                    data.name = dname)
     class(result) <- "htest"

### 2023-05-08 src/ff_internal_functions.cpp

@@ -160,6 +160,13 @@ double permutationTestPvalue(unsigned int zLess,
     return (zLess + (1 + zEqual) * dis(prng)) / static_cast<double>(1 + nPermute);
 }
 
+// [[Rcpp::export(permutationTestPvalueOld)]]
+double permutationTestPvalueOld(unsigned int zLess,
+                                unsigned int zEqual,
+                                unsigned int nPermute) {
+    return (1 + zLess + zEqual) / static_cast<double>(1 + nPermute);
+}
+
 // [[Rcpp::export(permutationTestPvalueSeeded)]]
 double permutationTestPvalueSeeded(unsigned int zLess,
                                    unsigned int zEqual,

### 2023-05-08 R/RcppExports.R
Temp

@@ -9,6 +9,10 @@ permutationTestPvalue <- function(zLess, zEqual, nPermute) {
     .Call(`_fasano_franceschini_test_permutationTestPvalue`, zLess, zEqual, nPermute)
 }
 
+permutationTestPvalueOld <- function(zLess, zEqual, nPermute) {
+    .Call(`_fasano_franceschini_test_permutationTestPvalueOld`, zLess, zEqual, nPermute)
+}
+
 permutationTestPvalueSeeded <- function(zLess, zEqual, nPermute, seed) {
     .Call(`_fasano_franceschini_test_permutationTestPvalueSeeded`, zLess, zEqual, nPermute, seed)
 }

### 2023-05-08 src/RcppExports.cpp

@@ -36,6 +36,19 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
+// permutationTestPvalueOld
+double permutationTestPvalueOld(unsigned int zLess, unsigned int zEqual, unsigned int nPermute);
+RcppExport SEXP _fasano_franceschini_test_permutationTestPvalueOld(SEXP zLessSEXP, SEXP zEqualSEXP, SEXP nPermuteSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< unsigned int >::type zLess(zLessSEXP);
+    Rcpp::traits::input_parameter< unsigned int >::type zEqual(zEqualSEXP);
+    Rcpp::traits::input_parameter< unsigned int >::type nPermute(nPermuteSEXP);
+    rcpp_result_gen = Rcpp::wrap(permutationTestPvalueOld(zLess, zEqual, nPermute));
+    return rcpp_result_gen;
+END_RCPP
+}
 // permutationTestPvalueSeeded
 double permutationTestPvalueSeeded(unsigned int zLess, unsigned int zEqual, unsigned int nPermute, int seed);
 RcppExport SEXP _fasano_franceschini_test_permutationTestPvalueSeeded(SEXP zLessSEXP, SEXP zEqualSEXP, SEXP nPermuteSEXP, SEXP seedSEXP) {
@@ -114,6 +127,7 @@ END_RCPP
 static const R_CallMethodDef CallEntries[] = {
     {"_fasano_franceschini_test_ffTestStatistic", (DL_FUNC) &_fasano_franceschini_test_ffTestStatistic, 3},
     {"_fasano_franceschini_test_permutationTestPvalue", (DL_FUNC) &_fasano_franceschini_test_permutationTestPvalue, 3},
+    {"_fasano_franceschini_test_permutationTestPvalueOld", (DL_FUNC) &_fasano_franceschini_test_permutationTestPvalueOld, 3},
     {"_fasano_franceschini_test_permutationTestPvalueSeeded", (DL_FUNC) &_fasano_franceschini_test_permutationTestPvalueSeeded, 4},
     {"_fasano_franceschini_test_permutationTestSeeded", (DL_FUNC) &_fasano_franceschini_test_permutationTestSeeded, 6},
[…]

### 2023-05-08 R/RcppExports.R
2.2.1

@@ -9,10 +9,6 @@ permutationTestPvalue <- function(zLess, zEqual, nPermute) {
     .Call(`_fasano_franceschini_test_permutationTestPvalue`, zLess, zEqual, nPermute)
 }
 
-permutationTestPvalueOld <- function(zLess, zEqual, nPermute) {
-    .Call(`_fasano_franceschini_test_permutationTestPvalueOld`, zLess, zEqual, nPermute)
-}
-
 permutationTestPvalueSeeded <- function(zLess, zEqual, nPermute, seed) {
     .Call(`_fasano_franceschini_test_permutationTestPvalueSeeded`, zLess, zEqual, nPermute, seed)
 }

### 2023-05-08 R/ff_function.R

@@ -148,7 +148,6 @@ fasano.franceschini.test <- function(S1,
 
     # Compute the p-value (if necessary)
     pval <- NULL
-    pval_old <- NULL
     if (nPermute > 0) {
         if (threads > 1) {
             # Run parallel version of permutation test
@@ -173,14 +172,11 @@ fasano.franceschini.test <- function(S1,
             pval <- permutationTestPvalueSeeded(counts[1], counts[2], nPermute, seed)
         }
         names(pval) <- "p-value"
-
-        pval_old <- permutationTestPvalueOld(counts[1], counts[2], nPermute)
     }
 
     # Construct output
     result <- list(statistic = Dff,
                    p.value = pval,
-                   p.value.old = pval_old,
                    method = "Fasano-Franceschini Test",
                    data.name = dname)
     class(result) <- "htest"

### 2023-05-08 src/RcppExports.cpp

@@ -36,19 +36,6 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
-// permutationTestPvalueOld
-double permutationTestPvalueOld(unsigned int zLess, unsigned int zEqual, unsigned int nPermute);
-RcppExport SEXP _fasano_franceschini_test_permutationTestPvalueOld(SEXP zLessSEXP, SEXP zEqualSEXP, SEXP nPermuteSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject rcpp_result_gen;
-    Rcpp::RNGScope rcpp_rngScope_gen;
-    Rcpp::traits::input_parameter< unsigned int >::type zLess(zLessSEXP);
-    Rcpp::traits::input_parameter< unsigned int >::type zEqual(zEqualSEXP);
-    Rcpp::traits::input_parameter< unsigned int >::type nPermute(nPermuteSEXP);
-    rcpp_result_gen = Rcpp::wrap(permutationTestPvalueOld(zLess, zEqual, nPermute));
-    return rcpp_result_gen;
-END_RCPP
-}
 // permutationTestPvalueSeeded
 double permutationTestPvalueSeeded(unsigned int zLess, unsigned int zEqual, unsigned int nPermute, int seed);
 RcppExport SEXP _fasano_franceschini_test_permutationTestPvalueSeeded(SEXP zLessSEXP, SEXP zEqualSEXP, SEXP nPermuteSEXP, SEXP seedSEXP) {
@@ -127,7 +114,6 @@ END_RCPP
 static const R_CallMethodDef CallEntries[] = {
     {"_fasano_franceschini_test_ffTestStatistic", (DL_FUNC) &_fasano_franceschini_test_ffTestStatistic, 3},
     {"_fasano_franceschini_test_permutationTestPvalue", (DL_FUNC) &_fasano_franceschini_test_permutationTestPvalue, 3},
-    {"_fasano_franceschini_test_permutationTestPvalueOld", (DL_FUNC) &_fasano_franceschini_test_permutationTestPvalueOld, 3},
     {"_fasano_franceschini_test_permutationTestPvalueSeeded", (DL_FUNC) &_fasano_franceschini_test_permutationTestPvalueSeeded, 4},
     {"_fasano_franceschini_test_permutationTestSeeded", (DL_FUNC) &_fasano_franceschini_test_permutationTestSeeded, 6},
[…]

### 2023-05-08 src/ff_internal_functions.cpp

@@ -160,13 +160,6 @@ double permutationTestPvalue(unsigned int zLess,
     return (zLess + (1 + zEqual) * dis(prng)) / static_cast<double>(1 + nPermute);
 }
 
-// [[Rcpp::export(permutationTestPvalueOld)]]
-double permutationTestPvalueOld(unsigned int zLess,
-                                unsigned int zEqual,
-                                unsigned int nPermute) {
-    return (1 + zLess + zEqual) / static_cast<double>(1 + nPermute);
-}
-
 // [[Rcpp::export(permutationTestPvalueSeeded)]]
 double permutationTestPvalueSeeded(unsigned int zLess,
                                    unsigned int zEqual,

### 2024-02-07 README.Rmd
2.2.2

@@
 
-`r badge_custom(x = "arXiv", y = "abs/2106.10539", color = "yellow", url = "https://arxiv.org/abs/2106.10539")` 
+`r badge_custom(x = "R Journal", y = "10.32614/RJ-2023-067", color = "yellow", url = "https://doi.org/10.32614/RJ-2023-067")`
 `r badge_cran_release("fasano.franceschini.test", "orange")`
@@
 
-The `fasano.franceschini.test` package is an R implementation of the multidimensional Kolmogorov-Smirnov two-sample test as defined by Fasano and Franceschini (1987).
-
+The `fasano.franceschini.test` package is an R implementation of the multivariate Kolmogorov-Smirnov two-sample test as defined by Fasano and Franceschini (1987).
 ```
@@
 ```
+A manuscript accompanying this package was published in The R Journal.
+```
+Puritz, C., Ness-Cohn, E., Braun, R. (2023). fasano.franceschini.test: An Implementation
+of a Multivariate KS Test in R. The R Journal, 15(3):159-171. doi: 10.32614/RJ-2023-067.
+```
 
@@
 
-You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://cran.r-project.org/) with:
-
+You can install the released version of the `fasano.franceschini.test` package from CRAN with:
 ``` r
@@
 
-The development version of the `fasano.franceschini.test` package can be installed from Github with:
-
+The development version of the `fasano.franceschini.test` package can be installed from GitHub with:
 ``` r
-# install.packages("devtools")
+if (!require("devtools", quietly = TRUE))
+    install.packages("devtools")
 devtools::install_github("braunlab-nu/fasano.franceschini.test")
@@
 #### Underlying distributions are the same
-
 ```{r}
@@
 #### Underlying distributions are different
-
 ```{r}

### 2024-02-07 R/ff_function.R

@@ -1,6 +1,6 @@
 #' Fasano-Franceschini Test
 #'
-#' Performs a two-sample multidimensional Kolmogorov-Smirnov test as described
+#' Performs a two-sample multivariate Kolmogorov-Smirnov test as described
 #' by Fasano and Franceschini (1987). This test evaluates the null hypothesis
 #' that two i.i.d. random samples were drawn from the same underlying
 #' probability distribution. The data can be of any dimension and of any

### 2024-02-15 R/ff_function.R
Minor documentation updates

@@ -6,25 +6,27 @@
 #' probability distribution. The data can be of any dimension and of any
 #' type (continuous, discrete, or mixed).
 #'
-#' @param S1 \code{matrix} or \code{data.frame}.
-#' @param S2 \code{matrix} or \code{data.frame}.
-#' @param nPermute A nonnegative \code{integer} setting the number of permutations
-#' to use for performing the permutation test. Default is \code{100}. If set to
-#' \code{0}, only the test statistic is computed.
+#' @param S1 A \code{matrix} or \code{data.frame}. Each row represents one
+#' observation.
+#' @param S2 A \code{matrix} or \code{data.frame}. Each row represents one
+#' observation.
+#' @param nPermute A nonnegative \code{integer} setting the number of
+#' permutations to use for performing the permutation test. Default is \code{100}.
+#' If set to \code{0}, only the test statistic is computed.
 #' @param threads A positive \code{integer} or \code{"auto"} setting the number
 #' of threads to use during the permutation test. If set to \code{"auto"}, the
 #' number of threads is determined by \code{RcppParallel::defaultNumThreads()}.
 #' Default is \code{1}.
-#' @param seed An optional integer to seed the PRNG used for the permutation test.
-#' A seed must be passed to reproducibly compute p-values.
+#' @param seed An optional integer to seed the PRNG used for the permutation
+#' test. A seed must be passed to reproducibly compute p-values.
 #' @param verbose A \code{boolean} indicating whether to display a progress bar.
 #' Default is \code{TRUE}. Only available when \code{threads = 1}.
 #' @param method An optional \code{character} indicating which method to use to
 #' compute the test statistic. The two methods are \code{'r'} (range tree) and
 #' \code{'b'} (brute force). Both methods return the same results but may vary in
-#' computation speed. If this argument is not passed, the sample sizes and dimension
-#' of the data are used to infer which method is likely faster. See the Details
-#' section for more information.
+#' computation speed. If this argument is not passed, the sample sizes and
+#' dimension of the data are used to infer which method is likely faster. See the
+#' Details section for more information.
 #' @return A list of class \code{htest} containing the following components:
 #'   \item{statistic}{The value of the test statistic.}
 #'   \item{p.value}{The permutation test p-value.}

### 2024-02-15 README.Rmd
2.2.3

@@
 
-# set seed for reproducibility
+# Set seed to draw reproducible samples
 set.seed(0)
 
-# create 2D samples from the same underlying distribution
+# Create 2D samples from the same underlying distribution
 S1 <- data.frame(x = rnorm(n = 50, mean = 0, sd = 1),
@@
 
+# Set seed to generate reproducible p-values
 fasano.franceschini.test(S1, S2, seed = 0)
@@
 ```{r}
-# set seed for reproducibility
+# Set seed to draw reproducible samples
 set.seed(1)
 
-# create 3D samples from different underlying distributions
+# Create 3D samples from different underlying distributions
 S1 <- cbind(rgamma(n = 43, shape = 2),
@@
 
+# Set seed to generate reproducible p-values
 fasano.franceschini.test(S1, S2, seed = 1)
